---
title: "NLGF S305N+3 vs S305N+3"
author:
  - name: "Emir Turkes"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_file = file.path(
      "..", "..", "..", "results",
      unlist(strsplit(getwd(), "/"))[6], unlist(strsplit(getwd(), "/"))[7], "NLGF_S305N+3_S305N+3.html")
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [tau-mutant-omics-analysis](https://github.com/eturkes/tau-mutant-omics-analysis).*

The table of contents in the top left is clickable and can be used to quickly navigate the document.
To toggle the visibility of code, use the `CODE` toggles at the top right of chunks.
The toggle at the start of the document controls the visibility of all chunks.

```{r}
#    This file is part of tau-mutant-omics-analysis.
#    Copyright (C) 2023  Emir Turkes, Naoto Watamura, UK DRI at UCL
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# Load required packages, suppressing startup messages.
# -----------------------------------------------------
library(conflicted)
conflicts_prefer(
  base::intersect, DT::JS, edgeR::cpm, AnnotationDbi::select, matrixStats::rowMedians, dplyr::filter,
  MatrixGenerics::colMedians, .quiet = TRUE
)
packages <- c(
  "Seurat", "GSEABase", "GSVA", "biomaRt", "parallel", "ggplot2", "magrittr", "dplyr", "ggrepel", "scuttle", "edgeR",
  "DT", "scales", "RColorBrewer", "factoextra", "ComplexHeatmap", "reshape2", "pals", "dendextend", "GO.db", "tm",
  "igraph", "qgraph", "circlize", "parallelly", "multienrichjam"
)
invisible(suppressPackageStartupMessages(lapply(packages, FUN = library, character.only = TRUE)))
# -----------------------------------------------------

# Define global settings.
# -----------------------
knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# -----------------------

# Define functions.
# -----------------
source(file.path("..", "..", "utils.R"))

`%notin%` <- Negate(`%in%`)
# -----------------

# Useful variables.
# -----------------
analysis_no <- 1
analysis_name <- "NLGF_S305N+3_S305N+3"
data_name <- unlist(strsplit(getwd(), "/"))[7] # Name of dataset derived from path.

data_dir <- file.path("..", "..", "..", "data")
results_dir <- file.path("..", "..", "..", "results", "full-data", data_name)
cache_dir <- file.path("..", "..", "..", "cache", "full-data", data_name, paste0("0", analysis_no))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# -----------------
```

# Prep

```{r}
gsva1 <- readRDS(file.path(cache_dir, "..", "..", "NLGF_S305N+3", "02", "gsva.rds"))
gsva2 <- readRDS(file.path(cache_dir, "..", "..", "S305N+3", "02", "gsva.rds"))

seurat1 <- readRDS(file.path(cache_dir, "..", "..", "NLGF_S305N+3", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path(cache_dir, "..", "..", "S305N+3", "02", "seurat_cleaned.rds"))
```

```{r}
gsva1_sub <- gsva1[ , which(seurat1$seurat_clusters == 2)]
gsva2_sub <- gsva2[ , which(seurat2$seurat_clusters == 1)]

gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(c(rep("NLGF_S305N", ncol(gsva1_sub)), rep("S305N", ncol(gsva2_sub))))

design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_S305N-S305N, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in CA1 NLGF S305N+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in CA1 NLGF S305N+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[1]]), ]

color <- colorRampPalette(rev(brewer.pal(9, "RdBu")))(100)

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    color,
    cluster_rows = TRUE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    show_row_names = FALSE
  )
)
```

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$seurat_clusters == 2)])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$seurat_clusters == 1)])

seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(c(rep("NLGF_S305N", ncol(seurat1_sub)), rep("S305N", ncol(seurat2_sub))))

design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_S305N-S305N, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in CA1 NLGF S305N+3 Genes**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in CA1 NLGF S305N+3 Genes**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
