---
title: "Combined Samples - 05 NFT Signature"
author:
  - name: "Emir Turkes"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_file = file.path(
      "..", "..", "results", unlist(strsplit(getwd(), "/"))[6], "comb_05_NFT_signature.html"
  ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [tau-mutant-omics-analysis](https://github.com/eturkes/tau-mutant-omics-analysis).*

The table of contents in the top left is clickable and can be used to quickly navigate the document.
To toggle the visibility of code, use the `CODE` toggles at the top right of chunks.
The toggle at the start of the document controls the visibility of all chunks.

```{r}
#    This file is part of tau-mutant-omics-analysis.
#    Copyright (C) 2023  Emir Turkes, Naoto Watamura, UK DRI at UCL
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# Load required packages, suppressing startup messages.
# -----------------------------------------------------
library(conflicted)
conflicts_prefer(base::intersect, .quiet = TRUE)
packages <- c("Seurat", "biomaRt", "ggplot2", "dplyr", "ggrepel", "parallelly")
invisible(suppressPackageStartupMessages(lapply(packages, FUN = library, character.only = TRUE)))
# -----------------------------------------------------

# Define global settings.
# -----------------------
knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# -----------------------

# Define functions.
# -----------------
source(file.path("..", "utils.R"))
# -----------------

# Useful variables.
# -----------------
analysis_no <- 5
data_name <- unlist(strsplit(getwd(), "/"))[6] # Name of dataset derived from path.

data_dir <- file.path("..", "..", "data")
cache_dir <- file.path("..", "..", "cache", data_name, paste0("0", analysis_no)) # Store intermediate data objects here.
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# -----------------
```

# All

```{r}
rds <- file.path(cache_dir, "NFT_sig_seurat.rds")
if (file.exists(rds)) {
  NFT_sig_seurat <- readRDS(rds)
} else {

  NFT_sig_seurat <- readRDS(file.path(cache_dir, "..", "02", "seurat.rds"))

  hub_genes <- read.csv(
    file.path(data_dir, "gene-sets", "otero-garcia-2022-ssRNAseq", "hub-gene-data-mouse-ortho.csv"), header = FALSE
  )$V1

  DefaultAssay(NFT_sig_seurat) <- "RNA"
  NFT_sig_seurat[["SCT"]] <- NULL

  gc()
  mem <- as.numeric(unlist(strsplit(system("free -b", TRUE)[2], " "))[15])
  mem <- mem - 10 * 1024 ^ 3
  mem <- mem / availableCores()
  options(future.globals.maxSize = mem)
  plan("multisession")

  NFT_sig_seurat <- SCTransform(
    NFT_sig_seurat, residual.features = rownames(NFT_sig_seurat)[rownames(NFT_sig_seurat) %in% hub_genes],
    vst.flavor = "v2", verbose = FALSE
  )

  NFT_sig_seurat <- RunPCA(NFT_sig_seurat, npcs = 30, verbose = FALSE)
  add_df <- data.frame(Embeddings(NFT_sig_seurat)[ , 1:2])
  names(add_df) <- paste0("NFT_pca", seq(ncol(add_df)))
  NFT_sig_seurat$NFT_pca1 <- add_df$NFT_pca1
  NFT_sig_seurat$NFT_pca2 <- add_df$NFT_pca2

  NFT_sig_seurat <- RunUMAP(NFT_sig_seurat, dims = 1:30, reduction = "pca", verbose = FALSE)
  add_df <- data.frame(Embeddings(NFT_sig_seurat, "umap"))
  names(add_df) <- paste0("NFT_umap", seq(ncol(add_df)))
  NFT_sig_seurat$NFT_umap1 <- add_df$NFT_umap1
  NFT_sig_seurat$NFT_umap2 <- add_df$NFT_umap2

  set.seed(1)
  kmeans <- kmeans(data.frame(Embeddings(NFT_sig_seurat)[ , c(1, 2)]), centers = 2)
  NFT_sig_seurat$NFT_sig <- factor(kmeans$cluster)

  saveRDS(NFT_sig_seurat, rds)
}

print(NFT_sig_seurat[["pca"]], dims = c(1, 2), nfeatures = Inf)
NFT_sig_seurat$NFT_sig <- ifelse(NFT_sig_seurat$NFT_sig == 1, "Pos", "Neg")

red_dim_plot(NFT_sig_seurat, x = "umap1", y = "umap2", color = "NFT_sig")
red_dim_plot(NFT_sig_seurat, x = "NFT_umap1", y = "NFT_umap2", color = "NFT_sig")
red_dim_plot(NFT_sig_seurat, x = "NFT_umap1", y = "NFT_umap2", color = "sample")
red_dim_plot(NFT_sig_seurat, x = "NFT_pca1", y = "NFT_pca2", color = "NFT_sig")
red_dim_plot(NFT_sig_seurat, x = "NFT_pca1", y = "NFT_pca2", color = "sample")

table <- prop.table(table(NFT_sig_seurat$NFT_sig, NFT_sig_seurat$sample), margin = 2)
table

table <- as.data.frame(table)
colnames(table) <- c("NFT Signature", "Sample", "Proportion")
table_sub <- table[table$`NFT Signature` == "Pos", ]
table$Sample <- factor(table$Sample, levels = table_sub$Sample[order(table_sub$Proportion, decreasing = TRUE)])
ggplot(table, aes(x = Sample, y = Proportion, fill = `NFT Signature`)) + geom_col() + ggtitle("All Cells")
```

# ENT EX Neurons

```{r}
rds <- file.path(cache_dir, "NFT_sig_seurat_ENT.rds")
if (file.exists(rds)) {
  NFT_sig_seurat <- readRDS(rds)
} else {

  NFT_sig_seurat <- readRDS(file.path(cache_dir, "..", "02", "seurat.rds"))
  NFT_sig_seurat <- subset(NFT_sig_seurat, idents = "10")

  hub_genes <- read.csv(
    file.path(data_dir, "gene-sets", "otero-garcia-2022-ssRNAseq", "hub-gene-data-mouse-ortho.csv"), header = FALSE
  )$V1

  DefaultAssay(NFT_sig_seurat) <- "RNA"
  NFT_sig_seurat[["SCT"]] <- NULL

  gc()
  mem <- as.numeric(unlist(strsplit(system("free -b", TRUE)[2], " "))[15])
  mem <- mem - 10 * 1024 ^ 3
  mem <- mem / availableCores()
  options(future.globals.maxSize = mem)
  plan("multisession")

  NFT_sig_seurat <- SCTransform(
    NFT_sig_seurat, residual.features = rownames(NFT_sig_seurat)[rownames(NFT_sig_seurat) %in% hub_genes],
    vst.flavor = "v2", verbose = FALSE
  )

  NFT_sig_seurat <- RunPCA(NFT_sig_seurat, npcs = 30, verbose = FALSE)
  add_df <- data.frame(Embeddings(NFT_sig_seurat)[ , 1:2])
  names(add_df) <- paste0("NFT_pca", seq(ncol(add_df)))
  NFT_sig_seurat$NFT_pca1 <- add_df$NFT_pca1
  NFT_sig_seurat$NFT_pca2 <- add_df$NFT_pca2

  NFT_sig_seurat <- RunUMAP(NFT_sig_seurat, dims = 1:30, reduction = "pca", verbose = FALSE)
  add_df <- data.frame(Embeddings(NFT_sig_seurat, "umap"))
  names(add_df) <- paste0("NFT_umap", seq(ncol(add_df)))
  NFT_sig_seurat$NFT_umap1 <- add_df$NFT_umap1
  NFT_sig_seurat$NFT_umap2 <- add_df$NFT_umap2

  set.seed(1)
  kmeans <- kmeans(data.frame(Embeddings(NFT_sig_seurat)[ , c(1, 2)]), centers = 2)
  NFT_sig_seurat$NFT_sig <- factor(kmeans$cluster)

  saveRDS(NFT_sig_seurat, rds)
}

print(NFT_sig_seurat[["pca"]], dims = c(1, 2), nfeatures = Inf)
NFT_sig_seurat$NFT_sig <- ifelse(NFT_sig_seurat$NFT_sig == 1, "Pos", "Neg")

red_dim_plot(NFT_sig_seurat, x = "umap1", y = "umap2", color = "NFT_sig")
red_dim_plot(NFT_sig_seurat, x = "NFT_umap1", y = "NFT_umap2", color = "NFT_sig")
red_dim_plot(NFT_sig_seurat, x = "NFT_umap1", y = "NFT_umap2", color = "sample")
red_dim_plot(NFT_sig_seurat, x = "NFT_pca1", y = "NFT_pca2", color = "NFT_sig")
red_dim_plot(NFT_sig_seurat, x = "NFT_pca1", y = "NFT_pca2", color = "sample")

table <- prop.table(table(NFT_sig_seurat$NFT_sig, NFT_sig_seurat$sample), margin = 2)
table

table <- as.data.frame(table)
colnames(table) <- c("NFT Signature", "Sample", "Proportion")
table_sub <- table[table$`NFT Signature` == "Pos", ]
table$Sample <- factor(table$Sample, levels = table_sub$Sample[order(table_sub$Proportion, decreasing = TRUE)])
ggplot(table, aes(x = Sample, y = Proportion, fill = `NFT Signature`)) + geom_col() + ggtitle("ENT EX Neurons")
```

# CA3 EX Neurons

```{r}
rds <- file.path(cache_dir, "NFT_sig_seurat_CA3.rds")
if (file.exists(rds)) {
  NFT_sig_seurat <- readRDS(rds)
} else {

  NFT_sig_seurat <- readRDS(file.path(cache_dir, "..", "02", "seurat.rds"))
  NFT_sig_seurat <- subset(NFT_sig_seurat, idents = "13")

  hub_genes <- read.csv(
    file.path(data_dir, "gene-sets", "otero-garcia-2022-ssRNAseq", "hub-gene-data-mouse-ortho.csv"), header = FALSE
  )$V1

  DefaultAssay(NFT_sig_seurat) <- "RNA"
  NFT_sig_seurat[["SCT"]] <- NULL

  gc()
  mem <- as.numeric(unlist(strsplit(system("free -b", TRUE)[2], " "))[15])
  mem <- mem - 10 * 1024 ^ 3
  mem <- mem / availableCores()
  options(future.globals.maxSize = mem)
  plan("multisession")

  NFT_sig_seurat <- SCTransform(
    NFT_sig_seurat, residual.features = rownames(NFT_sig_seurat)[rownames(NFT_sig_seurat) %in% hub_genes],
    vst.flavor = "v2", verbose = FALSE
  )

  NFT_sig_seurat <- RunPCA(NFT_sig_seurat, npcs = 30, verbose = FALSE)
  add_df <- data.frame(Embeddings(NFT_sig_seurat)[ , 1:2])
  names(add_df) <- paste0("NFT_pca", seq(ncol(add_df)))
  NFT_sig_seurat$NFT_pca1 <- add_df$NFT_pca1
  NFT_sig_seurat$NFT_pca2 <- add_df$NFT_pca2

  NFT_sig_seurat <- RunUMAP(NFT_sig_seurat, dims = 1:30, reduction = "pca", verbose = FALSE)
  add_df <- data.frame(Embeddings(NFT_sig_seurat, "umap"))
  names(add_df) <- paste0("NFT_umap", seq(ncol(add_df)))
  NFT_sig_seurat$NFT_umap1 <- add_df$NFT_umap1
  NFT_sig_seurat$NFT_umap2 <- add_df$NFT_umap2

  set.seed(1)
  kmeans <- kmeans(data.frame(Embeddings(NFT_sig_seurat)[ , c(1, 2)]), centers = 2)
  NFT_sig_seurat$NFT_sig <- factor(kmeans$cluster)

  saveRDS(NFT_sig_seurat, rds)
}

print(NFT_sig_seurat[["pca"]], dims = c(1, 2), nfeatures = Inf)
NFT_sig_seurat$NFT_sig <- ifelse(NFT_sig_seurat$NFT_sig == 1, "Pos", "Neg")

red_dim_plot(NFT_sig_seurat, x = "umap1", y = "umap2", color = "NFT_sig")
red_dim_plot(NFT_sig_seurat, x = "NFT_umap1", y = "NFT_umap2", color = "NFT_sig")
red_dim_plot(NFT_sig_seurat, x = "NFT_umap1", y = "NFT_umap2", color = "sample")
red_dim_plot(NFT_sig_seurat, x = "NFT_pca1", y = "NFT_pca2", color = "NFT_sig")
red_dim_plot(NFT_sig_seurat, x = "NFT_pca1", y = "NFT_pca2", color = "sample")

table <- prop.table(table(NFT_sig_seurat$NFT_sig, NFT_sig_seurat$sample), margin = 2)
table

table <- as.data.frame(table)
colnames(table) <- c("NFT Signature", "Sample", "Proportion")
table_sub <- table[table$`NFT Signature` == "Pos", ]
table$Sample <- factor(table$Sample, levels = table_sub$Sample[order(table_sub$Proportion, decreasing = TRUE)])
ggplot(table, aes(x = Sample, y = Proportion, fill = `NFT Signature`)) + geom_col() + ggtitle("CA3 EX Neurons")
```

# CA1 EX Neurons

```{r}
rds <- file.path(cache_dir, "NFT_sig_seurat_CA1.rds")
if (file.exists(rds)) {
  NFT_sig_seurat <- readRDS(rds)
} else {

  NFT_sig_seurat <- readRDS(file.path(cache_dir, "..", "02", "seurat.rds"))
  NFT_sig_seurat <- subset(NFT_sig_seurat, idents = "5")

  hub_genes <- read.csv(
    file.path(data_dir, "gene-sets", "otero-garcia-2022-ssRNAseq", "hub-gene-data-mouse-ortho.csv"), header = FALSE
  )$V1

  DefaultAssay(NFT_sig_seurat) <- "RNA"
  NFT_sig_seurat[["SCT"]] <- NULL

  gc()
  mem <- as.numeric(unlist(strsplit(system("free -b", TRUE)[2], " "))[15])
  mem <- mem - 10 * 1024 ^ 3
  mem <- mem / availableCores()
  options(future.globals.maxSize = mem)
  plan("multisession")

  NFT_sig_seurat <- SCTransform(
    NFT_sig_seurat, residual.features = rownames(NFT_sig_seurat)[rownames(NFT_sig_seurat) %in% hub_genes],
    vst.flavor = "v2", verbose = FALSE
  )

  NFT_sig_seurat <- RunPCA(NFT_sig_seurat, npcs = 30, verbose = FALSE)
  add_df <- data.frame(Embeddings(NFT_sig_seurat)[ , 1:2])
  names(add_df) <- paste0("NFT_pca", seq(ncol(add_df)))
  NFT_sig_seurat$NFT_pca1 <- add_df$NFT_pca1
  NFT_sig_seurat$NFT_pca2 <- add_df$NFT_pca2

  NFT_sig_seurat <- RunUMAP(NFT_sig_seurat, dims = 1:30, reduction = "pca", verbose = FALSE)
  add_df <- data.frame(Embeddings(NFT_sig_seurat, "umap"))
  names(add_df) <- paste0("NFT_umap", seq(ncol(add_df)))
  NFT_sig_seurat$NFT_umap1 <- add_df$NFT_umap1
  NFT_sig_seurat$NFT_umap2 <- add_df$NFT_umap2

  set.seed(1)
  kmeans <- kmeans(data.frame(Embeddings(NFT_sig_seurat)[ , c(1, 2)]), centers = 2)
  NFT_sig_seurat$NFT_sig <- factor(kmeans$cluster)

  saveRDS(NFT_sig_seurat, rds)
}

print(NFT_sig_seurat[["pca"]], dims = c(1, 2), nfeatures = Inf)
NFT_sig_seurat$NFT_sig <- ifelse(NFT_sig_seurat$NFT_sig == 2, "Pos", "Neg")

red_dim_plot(NFT_sig_seurat, x = "umap1", y = "umap2", color = "NFT_sig")
red_dim_plot(NFT_sig_seurat, x = "NFT_umap1", y = "NFT_umap2", color = "NFT_sig")
red_dim_plot(NFT_sig_seurat, x = "NFT_umap1", y = "NFT_umap2", color = "sample")
red_dim_plot(NFT_sig_seurat, x = "NFT_pca1", y = "NFT_pca2", color = "NFT_sig")
red_dim_plot(NFT_sig_seurat, x = "NFT_pca1", y = "NFT_pca2", color = "sample")

table <- prop.table(table(NFT_sig_seurat$NFT_sig, NFT_sig_seurat$sample), margin = 2)
table

table <- as.data.frame(table)
colnames(table) <- c("NFT Signature", "Sample", "Proportion")
table_sub <- table[table$`NFT Signature` == "Pos", ]
table$Sample <- factor(table$Sample, levels = table_sub$Sample[order(table_sub$Proportion, decreasing = TRUE)])
ggplot(table, aes(x = Sample, y = Proportion, fill = `NFT Signature`)) + geom_col() + ggtitle("CA1 EX Neurons")
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
