---
title: "S305N+3 - 02 GeneFunnel"
author:
  - name: "Emir Turkes"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_file = file.path(
      "..", "..", "..", "results",
      unlist(strsplit(getwd(), "/"))[6], unlist(strsplit(getwd(), "/"))[7], "S305N+3_02_genefunnel.html")
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [tau-mutant-omics-analysis](https://github.com/eturkes/tau-mutant-omics-analysis).*

The table of contents in the top left is clickable and can be used to quickly navigate the document.
To toggle the visibility of code, use the `CODE` toggles at the top right of chunks.
The toggle at the start of the document controls the visibility of all chunks.

```{r}
#    This file is part of tau-mutant-omics-analysis.
#    Copyright (C) 2023  Emir Turkes, Naoto Watamura, UK DRI at UCL
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# Load required packages, suppressing startup messages.
# -----------------------------------------------------
library(conflicted)
conflicts_prefer(
  base::intersect, DT::JS, edgeR::cpm, AnnotationDbi::select, matrixStats::rowMedians, dplyr::filter,
  MatrixGenerics::colMedians, .quiet = TRUE
)
packages <- c(
  "Seurat", "GSEABase", "GSVA", "biomaRt", "parallel", "ggplot2", "magrittr", "dplyr", "ggrepel", "scuttle", "edgeR",
  "DT", "scales", "RColorBrewer", "factoextra", "ComplexHeatmap", "reshape2", "pals", "dendextend", "GO.db", "tm",
  "igraph", "qgraph", "circlize", "parallelly", "multienrichjam"
)
invisible(suppressPackageStartupMessages(lapply(packages, FUN = library, character.only = TRUE)))
# -----------------------------------------------------

# Define global settings.
# -----------------------
knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# -----------------------

# Define functions.
# -----------------
source(file.path("..", "..", "utils.R"))

`%notin%` <- Negate(`%in%`)
# -----------------

# Useful variables.
# -----------------
analysis_no <- 2
analysis_name <- "S305N+3"
data_name <- unlist(strsplit(getwd(), "/"))[7] # Name of dataset derived from path.

data_dir <- file.path("..", "..", "..", "data")
results_dir <- file.path("..", "..", "..", "results", "full-data", data_name)
cache_dir <- file.path("..", "..", "..", "cache", "full-data", data_name, paste0("0", analysis_no))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# -----------------
```

# Prep

```{r}
# Load and filter gene sets.
# --------------------------
rds <- file.path(cache_dir, "filtered_GO_comb_gene_sets.rds")
if (file.exists(rds)) {
  gene_sets <- readRDS(rds)
} else {

  gene_sets <- getGmt(file.path(data_dir, "gene-sets", "gprofiler_mmusculus.ENSG", "mmusculus.GO.comb.ENSG.gmt"))
  keep <- filterGeneSets(gene_sets, min.sz = 5, max.sz = 25)
  gene_sets <- gene_sets[names(gene_sets) %in% names(keep)]

  for (i in seq(length(gene_sets@.Data))) {
    go_id <- gene_sets[[i]]@setName
    suppressWarnings(gene_sets[[i]]@setName <- gene_sets[[i]]@shortDescription)
    suppressWarnings(gene_sets[[i]]@shortDescription <- go_id)
  }

  if (any(duplicated(names(gene_sets))) == TRUE) {
    remove <- which(duplicated(names(gene_sets)))
    gene_sets <- gene_sets[-remove]
  }

  remove <- grep("positive regulation of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }
  remove <- grep("negative regulation of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }
  remove <- grep("regulation of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }
  remove <- grep("downregulation of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }
  remove <- grep("inhibition of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }
  remove <- grep("termination of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }
  remove <- grep("activation of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }
  remove <- grep("maintenance of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }
  remove <- grep("upregulation of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }

  overlap <- computeGeneSetsOverlapMax(gene_sets, uniqGenes = unique(unlist(geneIds(gene_sets))))
  tmp <- rowSums(overlap)
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  gene_sets_sorted <- gene_sets[match(names(tmp), names(gene_sets))]

  overlap <- computeGeneSetsOverlapMax(gene_sets_sorted, uniqGenes = unique(unlist(geneIds(gene_sets_sorted))))
  overlap[upper.tri(overlap)] <- 0
  diag(overlap) <- 0
  keep <- apply(overlap, MARGIN = 1, FUN = max)
  keep <- keep[keep <= 0.25]
  gene_sets <- gene_sets[names(gene_sets) %in% names(keep)]

  saveRDS(gene_sets, rds)
}
# --------------------------

# Create a Seurat object containing only clusters of interest with downsampling.
# ------------------------------------------------------------------------------
rds <- file.path(cache_dir, "seurat_cleaned.rds")
rds2 <- file.path(cache_dir, "gene_anno.rds")
if (file.exists(rds) & file.exists(rds2)) {
  seurat <- readRDS(rds)
  gene_anno <- readRDS(rds2)
} else {

  seurat <- readRDS(file.path(cache_dir, "..", "01", "seurat_all.rds"))

  mart <- useEnsembl("ensembl", "mmusculus_gene_ensembl")
  attributes <- c("external_gene_name", "ensembl_gene_id")
  gene_anno <- getBM(attributes, filters = "ensembl_gene_id", values = unique(unlist(geneIds(gene_sets))), mart = mart)
  add <- which(gene_anno$external_gene_name == "")
  for (idx in add) {
    gene_anno[idx, 1] <- gene_anno[idx, 2]
  }

  dup <- gene_anno[duplicated(gene_anno$external_gene_name), ]
  if (nrow(dup) > 0) {
    for (i in seq(nrow(dup))) {
      for (j in seq(nrow(gene_anno))) {
        if (dup$ensembl_gene_id[i] == gene_anno$ensembl_gene_id[j]) {
          gene_anno$external_gene_name[j] <- paste0(gene_anno$external_gene_name[j], "-alt")
        }
      }
    }
    if (any(duplicated(gene_anno$external_gene_name))) {
      stop("Duplicates in gene_anno.")
    }
  }

  gene_anno_sub <- gene_anno[gene_anno$external_gene_name %in% rownames(seurat), ]
  seurat <- seurat[rownames(seurat) %in% gene_anno_sub$external_gene_name, ]
  gene_anno_sub <- gene_anno_sub[match(rownames(seurat), gene_anno_sub$external_gene_name), ]

  count_mat <- GetAssayData(seurat, layer = "counts")
  data_mat <- GetAssayData(seurat, layer = "data")

  rownames(count_mat) <- gene_anno_sub$ensembl_gene_id
  rownames(data_mat) <- gene_anno_sub$ensembl_gene_id
  seurat <- CreateSeuratObject(count_mat, data = data_mat, assay = "SCT", meta.data = seurat[[]])

  count_mat <- GetAssayData(seurat, layer = "counts")
  data_mat <- GetAssayData(seurat, layer = "data")

  add <- unique(gene_anno$ensembl_gene_id[which(gene_anno$ensembl_gene_id %notin% rownames(seurat))])
  null_mat <- matrix(0, nrow = length(add), ncol = ncol(count_mat))
  rownames(null_mat) <- add
  colnames(null_mat) <- colnames(seurat)

  set.seed(1)
  col <- sample(seq(ncol(null_mat)), size = length(add), replace = TRUE)
  for (i in seq_along(add)) {
    null_mat[i, col[i]] <- 1
  }

  count_mat <- rbind(count_mat, null_mat)
  data_mat <- rbind(data_mat, null_mat)

  seurat <- CreateSeuratObject(count_mat, data = data_mat, assay = "SCT", meta.data = seurat[[]])
  gene_anno <- gene_anno[match(rownames(seurat), gene_anno$ensembl_gene_id), ]

  rm(count_mat, data_mat, null_mat, gene_anno_sub, mart)
  saveRDS(seurat, rds)
  saveRDS(gene_anno, rds2)
}
# ------------------------------------------------------------------------------

hist(as.matrix(GetAssayData(seurat)), main = "SCTransform normalized log counts distribution")
```

```{r}
# Run GSVA on counts.
# -------------------
rds <- file.path(cache_dir, "gsva.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(
    as.matrix(GetAssayData(seurat, layer = "data")), gene_sets, method = "ssgsea",
    parallel.sz = availableCores(), ssgsea.norm = FALSE
  )
  saveRDS(gsva, rds)
}
# -------------------
```

```{r}
seurat[["GSVA"]] <- CreateAssay5Object(gsva)
DefaultAssay(seurat) <- "GSVA"
seurat@active.ident <- seurat$seurat_clusters

rds <- file.path(cache_dir, "markers.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, slot = "counts", only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
markers <- markers %>% distinct(gene, .keep_all = TRUE)
rownames(markers) <- markers$gene

# top_markers <- markers %>% group_by(cluster) %>% top_n(-1, p_val_adj)
# 
# markers <- markers %>% group_by(cluster) %>% filter(data.table::frank(p_val_adj, ties.method = "first") %in% 1:25)
# markers <- as.data.frame(markers)
# rownames(markers) <- markers$gene

sce <- as.SingleCellExperiment(seurat, assay = "GSVA")
sce <- suppressWarnings(
  aggregateAcrossCells(
    sce, colData(sce)[ , "seurat_clusters"], use_exprs_values = "counts", statistics = "median"
  )
)
# dge <- DGEList(counts(sce))
# dge <- calcNormFactors(dge)
# logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

top_genes <- c()
for (i in seq(ncol(sce))) {
  top_genes <- c(top_genes, rownames(sce)[order(counts(sce[ , i]), decreasing = TRUE)[1]])
}

umap <- as.matrix(data.frame(UMAP_1 = seurat$umap1, UMAP_2 = seurat$umap2))
seurat[["umap"]] <- CreateDimReducObject(embeddings = umap, key = "UMAP_", assay = "GSVA")
```

```{r}
markers$number <- seq(rownames(markers))

modules <- markers %>% group_by(cluster) %>%   summarise(named_vec = list(number)) %>%
    tibble::deframe()

module_names <- modules
gene_sets_sub_ids <- gene_sets[names(gene_sets) %in% rownames(markers)]
for (j in seq(length(gene_sets_sub_ids@.Data))) {
  suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
}
go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(markers)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]])
}
# for (l in seq_along(module_names)) {
#   module_names[[l]] <- paste0(rownames(mat)[module_names[[l]]])
# }
top_words_tmp <- data.frame(
  gene_sets = unlist(module_names), module = rep(seq_along(module_names), lapply(module_names, length))
)
top_words_list <- vector("list", length(unique(top_words_tmp$module)))
top_words <- c()
for (i in seq_along(top_words_list)) {
  print(i)
  top_words_tmp_sub <- top_words_tmp[top_words_tmp$module == i, ]
  top_words_list[[i]] <- vector("list", nrow(top_words_tmp_sub))
  for (j in seq_along(top_words_list[[i]])) {
    top_words_list[[i]][[j]] <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_sets[j]))
  }
  tmp <- table(strsplit(paste(unlist(top_words_list[[i]]), collapse = " "), "\\s+")[[1]])
  tmp <- names(tmp[tmp > 1])
  if (length(tmp) >= 3) {
    tmp <- tmp[1:3]
    tmp <- paste(paste0(tmp, "\n"), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp)-1)
  } else if (length(tmp > 0)) {
    add <- 3 - length(tmp)
    tmp2 <- unlist(strsplit(suppressWarnings(word_cloud(top_words_tmp_sub$gene_set)), "\\s+"))
    tmp2 <- tmp2[tmp2 %notin% tmp]
    tmp <- c(tmp, tmp2[1:add])
    tmp <- paste(paste0(tmp, "\n"), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp)-1)
  } else {
    tmp <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_set, 3))
  }
  top_words <- c(top_words, tmp)
}

top_words_all <- suppressWarnings(as.character(tapply(top_words_tmp$gene_sets, top_words_tmp$module, word_cloud)))
```

```{r, fig.height = 10}
mat <- counts(sce)
mat <- mat[rownames(mat) %in% rownames(markers), ]
mat <- mat[match(rownames(markers), rownames(mat)), ]

optimal_k <- as.numeric(unlist(lapply(seq_along(modules), function(i) {
    rep(names(modules)[i], length(modules[[i]]))
})))

module_colors <- cols25(length(unique(optimal_k)))

row_anno <-  HeatmapAnnotation(
  top_words = anno_block(
    gp = gpar(col = module_colors), labels = unique(optimal_k), labels_rot = 0, labels_gp = gpar(fontsize = 12)
  ),
  which = "row"
)

color <- colorRampPalette(rev(brewer.pal(9, "RdBu")))(100)

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    color,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = TRUE,
    show_row_names = FALSE,
    row_split = optimal_k,
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    left_annotation = row_anno,
    row_title = top_words,
    row_title_rot = 0,
    row_title_gp = gpar(fontsize = 12),
    column_names_rot = 0,
    column_names_gp = gpar(fontsize = 20)
  )
)
```

```{r}
new_mat <- matrix(data = NA, nrow = 0, ncol = ncol(mat))
for (i in seq_along(modules)) {
  if (length(modules[[i]]) == 1) {
    new_mat <- rbind(new_mat, mat[modules[[i]], ])
  } else {
    new_mat <- rbind(new_mat, colMedians(mat[modules[[i]], ]))
  }
}
```

```{r}
set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        new_mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    color,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = TRUE,
    show_row_names = FALSE,
    row_split = unique(optimal_k),
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    left_annotation = row_anno,
    row_title = top_words,
    row_title_rot = 0,
    row_title_gp = gpar(fontsize = 12),
    column_names_rot = 0,
    column_names_gp = gpar(fontsize = 20)
  )
)
```

```{r}
rds <- file.path(cache_dir, "cell_typing.rds")
if (file.exists(rds)) {
  seurat_orig <- readRDS(rds)
} else {
  seurat_orig <- readRDS(file.path(cache_dir, "..", "01", "seurat_all.rds"))
  ref <- readRDS(file.path(cache_dir, "..", "..", "..", "ref-datasets", "01", "allen_mouse_hip_ctx_10x.rds"))

  anchors <- FindTransferAnchors(
    ref, seurat_orig, normalization.method = "SCT", recompute.residuals = FALSE,
    reference.reduction = "pca", dims = 1:30, verbose = FALSE
  )
  seurat_orig <- MapQuery(
    anchors, seurat_orig, ref, refdata = "subclass_label", reference.reduction = "umap", verbose = FALSE
  )

  rm(ref)
  saveRDS(seurat_orig, rds)
}

red_dim_plot(seurat_orig, x = "umap1", y = "umap2", color = "predicted.id", type = "cat")

table <- table(seurat$seurat_clusters, seurat$predicted.id)
max <- colnames(table)[apply(table,1,which.max)]
colnames(new_mat) <- paste("Cluster", rownames(table), max, sep = " ")
```

```{r, fig.height = 7}
set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        new_mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    color,
    cluster_rows = FALSE,
    cluster_columns = TRUE,
    show_column_names = TRUE,
    show_row_names = FALSE,
    row_split = unique(optimal_k),
    rect_gp = gpar(col = "lightgray", lwd = 0.5),
    row_gap = unit(0, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "bottom",
    left_annotation = row_anno,
    row_title = top_words,
    row_title_rot = 0,
    column_names_rot = 45
  )
)
```

```{r}
new_mat_bak <- new_mat
rownames(new_mat) <- top_words
```

```{r, fig.height = 7}
set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        new_mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    color,
    cluster_rows = TRUE,
    cluster_columns = TRUE,
    show_column_names = TRUE,
    show_row_names = TRUE,
    rect_gp = gpar(col = "lightgray", lwd = 0.5),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "bottom",
    row_title_rot = 0,
    column_names_rot = 45,
    row_names_side = "left",
    row_dend_side = "right"
  )
)
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
