---
title: "Combined Samples - 02 Cell-typing"
author:
  - name: "Emir Turkes"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_file = file.path(
      "..", "..", "..", "results",
      unlist(strsplit(getwd(), "/"))[6], unlist(strsplit(getwd(), "/"))[7], "comb_02_cell_typing.html"
  )}))
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [tau-mutant-omics-analysis](https://github.com/eturkes/tau-mutant-omics-analysis).*

The table of contents in the top left is clickable and can be used to quickly navigate the document.
To toggle the visibility of code, use the `CODE` toggles at the top right of chunks.
The toggle at the start of the document controls the visibility of all chunks.

```{r}
#    This file is part of tau-mutant-omics-analysis.
#    Copyright (C) 2023  Emir Turkes, Naoto Watamura, UK DRI at UCL
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# Load required packages, suppressing startup messages.
# -----------------------------------------------------
library(conflicted)
conflicts_prefer(base::intersect, .quiet = TRUE)
packages <- c("Seurat", "ggplot2", "dplyr", "ggrepel")
invisible(suppressPackageStartupMessages(lapply(packages, FUN = library, character.only = TRUE)))
# -----------------------------------------------------

# Define global settings.
# -----------------------
knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# -----------------------

# Define functions.
# -----------------
source(file.path("..", "..", "utils.R"))
# -----------------

# Useful variables.
# -----------------
analysis_no <- 2
data_name <- unlist(strsplit(getwd(), "/"))[7] # Name of dataset derived from path.

cache_dir <- file.path("..", "..", "..", "cache", "low-coverage", data_name, paste0("0", analysis_no))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# -----------------
```

# Analysis

```{r, fig.width = 20, fig.height = 14}
rds <- file.path(cache_dir, "seurat.rds")
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {
  seurat <- readRDS(file.path(cache_dir, "..", "01", "processed_merged_seurat.rds"))
  ref <- readRDS(file.path(cache_dir, "..", "..", "ref-datasets", "01", "allen_mouse_hip_ctx_10x.rds"))

  anchors <- FindTransferAnchors(
    ref, seurat, normalization.method = "SCT", recompute.residuals = FALSE,
    reference.reduction = "pca", dims = 1:30, verbose = FALSE
  )
  seurat <- MapQuery(anchors, seurat, ref, refdata = "subclass_label", reference.reduction = "umap", verbose = FALSE)

  rm(ref)
  saveRDS(seurat, rds)
}

red_dim_plot(seurat, x = "umap1", y = "umap2", color = "predicted.id", type = "cat")
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
