---
title: "Combined Samples - 04 GeneFunnel NLGF MAPT KI vs. MAPT KI vs. NLGF S305N+3 vs. S305N+3 Microglia"
author:
  - name: "Emir Turkes"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_file = file.path(
      "..", "..",
      "..", "results",
      unlist(strsplit(getwd(), "/"))[6], unlist(strsplit(getwd(), "/"))[7],
      "comb_04_genefunnel_NLGF_MAPT_NLGF_S305N_S305N_microglia.html")
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [tau-mutant-omics-analysis](https://github.com/eturkes/tau-mutant-omics-analysis).*

The table of contents in the top left is clickable and can be used to quickly navigate the document.
To toggle the visibility of code, use the `CODE` toggles at the top right of chunks.
The toggle at the start of the document controls the visibility of all chunks.

```{r}
#    This file is part of tau-mutant-omics-analysis.
#    Copyright (C) 2023  Emir Turkes, Naoto Watamura, UK DRI at UCL
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# Load required packages, suppressing startup messages.
# -----------------------------------------------------
library(conflicted)
conflicts_prefer(base::intersect, DT::JS, edgeR::cpm, AnnotationDbi::select, matrixStats::rowMedians, .quiet = TRUE)
packages <- c(
  "Seurat", "GSEABase", "GSVA", "biomaRt", "parallel", "ggplot2", "magrittr", "dplyr", "ggrepel", "scuttle", "edgeR",
  "DT", "scales", "RColorBrewer", "factoextra", "ComplexHeatmap", "reshape2", "pals", "dendextend", "GO.db", "tm",
  "igraph", "qgraph", "circlize", "parallelly", "multienrichjam"
)
invisible(suppressPackageStartupMessages(lapply(packages, FUN = library, character.only = TRUE)))
# -----------------------------------------------------

# Define global settings.
# -----------------------
knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# -----------------------

# Define functions.
# -----------------
source(file.path("..", "..", "utils.R"))

`%notin%` <- Negate(`%in%`)
# -----------------

# Useful variables.
# -----------------
analysis_no <- 4
analysis_name <- "NLGF_MAPT_NLGF_S305N_S305N_microglia"
data_name <- unlist(strsplit(getwd(), "/"))[7] # Name of dataset derived from path.

data_dir <- file.path("..", "..", "..", "data")
results_dir <- file.path("..", "..", "..", "results", "low-coverage", data_name)
cache_dir <- file.path("..", "..", "..", "cache", "low-coverage", data_name, paste0("0", analysis_no), analysis_name)
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
```

# Prep

```{r}
# Load and filter gene sets.
# --------------------------
rds <- file.path(cache_dir, "filtered_GO_comb_gene_sets.rds")
if (file.exists(rds)) {
  gene_sets <- readRDS(rds)
} else {

  gene_sets <- getGmt(file.path(data_dir, "gene-sets", "gprofiler_mmusculus.ENSG", "mmusculus.GO.comb.ENSG.gmt"))
  keep <- filterGeneSets(gene_sets, min.sz = 5, max.sz = 25)
  gene_sets <- gene_sets[names(gene_sets) %in% names(keep)]

  for (i in seq(length(gene_sets@.Data))) {
    go_id <- gene_sets[[i]]@setName
    suppressWarnings(gene_sets[[i]]@setName <- gene_sets[[i]]@shortDescription)
    suppressWarnings(gene_sets[[i]]@shortDescription <- go_id)
  }

  if (any(duplicated(names(gene_sets))) == TRUE) {
    remove <- which(duplicated(names(gene_sets)))
    gene_sets <- gene_sets[-remove]
  }

  remove <- grep("positive regulation of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }
  remove <- grep("negative regulation of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }
  remove <- grep("regulation of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }
  remove <- grep("downregulation of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }
  remove <- grep("inhibition of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }
  remove <- grep("termination of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }
  remove <- grep("activation of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }
  remove <- grep("maintenance of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }
  remove <- grep("upregulation of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }

  overlap <- computeGeneSetsOverlapMax(gene_sets, uniqGenes = unique(unlist(geneIds(gene_sets))))
  tmp <- rowSums(overlap)
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  gene_sets_sorted <- gene_sets[match(names(tmp), names(gene_sets))]

  overlap <- computeGeneSetsOverlapMax(gene_sets_sorted, uniqGenes = unique(unlist(geneIds(gene_sets_sorted))))
  overlap[upper.tri(overlap)] <- 0
  diag(overlap) <- 0
  keep <- apply(overlap, MARGIN = 1, FUN = max)
  keep <- keep[keep <= 0.25]
  gene_sets <- gene_sets[names(gene_sets) %in% names(keep)]

  saveRDS(gene_sets, rds)
}
# --------------------------

# Create a Seurat object containing only clusters of interest with downsampling.
# ------------------------------------------------------------------------------
rds <- file.path(cache_dir, "seurat_cleaned.rds")
rds2 <- file.path(cache_dir, "gene_anno.rds")
if (file.exists(rds) & file.exists(rds2)) {
  seurat <- readRDS(rds)
  gene_anno <- readRDS(rds2)
} else {

  seurat <- readRDS(file.path(cache_dir, "..", "..", "02", "seurat.rds"))
  seurat <- CreateSeuratObject(GetAssayData(seurat, assay = "RNA", layer = "counts"), meta.data = seurat[[]])
  seurat@active.ident <- seurat$seurat_clusters

  seurat <- subset(seurat, idents = "11")

  sample1_seurat <- subset(seurat, subset = sample == "NLGF_S305N+3")
  sample2_seurat <- subset(seurat, subset = sample == "NLGF_MAPT_KI" )
  sample3_seurat <- subset(seurat, subset = sample == "S305N+3")
  sample4_seurat <- subset(seurat, subset = sample == "MAPT_KI")

  # Find the minimum number of cells from each of the samples.
  # This will be used to downsample each sample.
  # ----------------------------------------------------------
  downsample <- min(ncol(sample1_seurat), ncol(sample2_seurat), ncol(sample3_seurat), ncol(sample4_seurat))
  # ----------------------------------------------------------

  set.seed(1)
  sample1_seurat <- subset(sample1_seurat, cells = sample(Cells(sample1_seurat), size = downsample))
  set.seed(1)
  sample2_seurat <- subset(sample2_seurat, cells = sample(Cells(sample2_seurat), size = downsample))
  condition1_seurat <- merge(sample1_seurat, sample2_seurat, merge.dr = TRUE)

  set.seed(1)
  sample3_seurat <- subset(sample3_seurat, cells = sample(Cells(sample3_seurat), size = downsample))
  set.seed(1)
  sample4_seurat <- subset(sample4_seurat, cells = sample(Cells(sample4_seurat), size = downsample))
  condition2_seurat <- merge(sample3_seurat, sample4_seurat, merge.dr = TRUE)

  seurat <- merge(condition1_seurat, condition2_seurat, merge.dr = TRUE)
  seurat <- JoinLayers(seurat)
  rm(sample1_seurat, sample2_seurat, sample3_seurat, sample4_seurat, condition1_seurat, condition2_seurat)

  mart <- useEnsembl("ensembl", "mmusculus_gene_ensembl")
  attributes <- c("external_gene_name", "ensembl_gene_id")
  gene_anno <- getBM(attributes, filters = "ensembl_gene_id", values = unique(unlist(geneIds(gene_sets))), mart = mart)
  add <- which(gene_anno$external_gene_name == "")
  for (idx in add) {
    gene_anno[idx, 1] <- gene_anno[idx, 2]
  }

  dup <- gene_anno[duplicated(gene_anno$external_gene_name), ]
  if (nrow(dup) > 0) {
    for (i in seq(nrow(dup))) {
      for (j in seq(nrow(gene_anno))) {
        if (dup$ensembl_gene_id[i] == gene_anno$ensembl_gene_id[j]) {
          gene_anno$external_gene_name[j] <- paste0(gene_anno$external_gene_name[j], "-alt")
        }
      }
    }
    if (any(duplicated(gene_anno$external_gene_name))) {
      stop("Duplicates in gene_anno.")
    }
  }

  seurat <- seurat[rowSums(as.matrix(GetAssayData(seurat, layer = "counts")) > 0) > 0, ]
  gene_anno_sub <- gene_anno[gene_anno$external_gene_name %in% rownames(seurat), ]
  seurat <- seurat[rownames(seurat) %in% gene_anno_sub$external_gene_name, ]
  gene_anno_sub <- gene_anno_sub[match(rownames(seurat), gene_anno_sub$external_gene_name), ]
  new_mat <- GetAssayData(seurat, layer = "counts")
  rownames(new_mat) <- gene_anno_sub$ensembl_gene_id
  seurat <- CreateSeuratObject(new_mat, meta.data = seurat[[]])

  add <- unique(gene_anno$ensembl_gene_id[which(gene_anno$ensembl_gene_id %notin% rownames(seurat))])
  new_mat <- GetAssayData(seurat, layer = "counts")
  null_mat <- matrix(0, nrow = length(add), ncol = ncol(new_mat))
  rownames(null_mat) <- add
  colnames(null_mat) <- colnames(seurat)

  set.seed(1)
  col <- sample(seq(ncol(null_mat)), size = length(add), replace = TRUE)
  for (i in seq_along(add)) {
    null_mat[i, col[i]] <- 1
  }
  new_mat <- rbind(new_mat, null_mat)
  seurat <- CreateSeuratObject(new_mat, meta.data = seurat[[]])
  gene_anno <- gene_anno[match(rownames(seurat), gene_anno$ensembl_gene_id), ]

  seurat$condition <- paste(
    "Microglia", c(rep("NLGF", times = ncol(seurat) / 2), rep("WT", times = ncol(seurat) / 2)), sep = "_"
  )
  seurat$condition <- factor(seurat$condition, levels = unique(seurat$condition))
  seurat$sample <- factor(seurat$sample, levels = c("NLGF_S305N+3", "NLGF_MAPT_KI", "S305N+3", "MAPT_KI"))

  tmp <- SCTransform(
    GetAssayData(seurat, assay = "RNA", layer = "counts"), seurat[[]], vst.flavor = "v2", min_cells = 0, verbose = FALSE
  )
  seurat[["SCT"]] <- CreateAssay5Object(tmp$umi_corrected, log1p(tmp$umi_corrected))
  DefaultAssay(seurat) <- "SCT"
  rm(tmp)

  saveRDS(seurat, rds)
  saveRDS(gene_anno, rds2)
}
# ------------------------------------------------------------------------------

hist(as.matrix(GetAssayData(seurat)), main = "SCTransform normalized log counts distribution")

# sce <- as.SingleCellExperiment(seurat)
# logcounts(sce) <- as.matrix(counts(sce))
# sce_filt <- as.SingleCellExperiment(seurat[rowSums(as.matrix(GetAssayData(seurat, slot = "counts")) > 0) > 10, ])
# logcounts(sce_filt) <- as.matrix(counts(sce_filt))
# 
# hist(logcounts(sce_filt), main = "SCTransform normalized log counts distribution after filtering zero counts")

sce <- as.SingleCellExperiment(seurat, assay = "SCT")
keep <- filterByExpr(counts(sce), group = sce$condition, min.total.count = 1, min.count = 1, min.prop = 0.01)
sce_filt <- sce[keep, ]

dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

dge <- DGEList(counts(sce_filt))
dge <- calcNormFactors(dge)
logcounts(sce_filt, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

hist(logcounts(sce), main = "logCPM distribution")
hist(logcounts(sce_filt), main = "Filtered logCPM distribution")

# Create new Seurat object with gene symbols.
# -------------------------------------------
new_mat <- GetAssayData(seurat)
rownames(new_mat) <- gene_anno$external_gene_name
seurat_symbols <- CreateSeuratObject(new_mat, meta.data = seurat[[]])
# -------------------------------------------

# Run GSVA on pseudo-bulk counts.
# -------------------------------
rds <- file.path(cache_dir, "gsva.rds")
if (file.exists(rds)) {
  gsva <- readRDS(rds)
} else {
  gsva <- gsva(
    as.matrix(logcounts(sce)), gene_sets, method = "ssgsea",
    parallel.sz = availableCores(), ssgsea.norm = FALSE, verbose = FALSE
  )
  # gsva_filt <- gsva(
  #   as.matrix(logcounts(sce_filt)), gene_sets, method = "ssgsea", parallel.sz = availableCores(),
  #   min.sz = 5, ssgsea.norm = FALSE, verbose = FALSE
  # )
  # gsva <- gsva_orig[rownames(gsva_orig) %in% rownames(gsva_filt), ]

  colnames(gsva) <- sce$Sample.ID.Custom
  saveRDS(gsva, rds)
}
# -------------------------------

# gsva1 <- gsva[ , which(sce$condition == unique(sce$condition)[1])]
# gsva2 <- gsva[ , which(sce$condition == unique(sce$condition)[2])]
# gsva1 <- gsva1[rowMedians(gsva1) >= quantile(gsva)[2], ]
# gsva2 <- gsva2[rowMedians(gsva2) >= quantile(gsva)[2], ]
# gsva <- gsva[rownames(gsva) %in% c(rownames(gsva1), rownames(gsva2)), ]
# gene_sets <- gene_sets[which(names(gene_sets) %in% rownames(gsva))]

# keep <- (range(gsva)[1] + range(gsva)[2]) / 2.4
# gsva1 <- gsva[ , which(sce$condition == unique(sce$condition)[1])]
# gsva2 <- gsva[ , which(sce$condition == unique(sce$condition)[2])]
# gsva1 <- gsva1[rowMedians(gsva1) >= keep, ]
# gsva2 <- gsva2[rowMedians(gsva2) >= keep, ]
# gsva <- gsva[rownames(gsva) %in% c(rownames(gsva1), rownames(gsva2)), ]
# gene_sets <- gene_sets[which(names(gene_sets) %in% rownames(gsva))]

overlap <- lapply(geneIds(gene_sets), function(x, y) na.omit(fastmatch::fmatch(x, y)), rownames(sce_filt))
prop <- lengths(overlap) / lengths(geneIds(gene_sets))
keep <- which(prop >= 0.25)
gsva <- gsva[rownames(gsva) %in% names(keep), ]
# gene_sets <- gene_sets[keep]

hist(gsva, main = "GSVA distribution")
```

# Analysis

```{r}
# Fit model on pseudo-bulk GSVA results.
# --------------------------------------
design <- model.matrix(~ 0 + sce$condition)
colnames(design) <- make.names(unique(sce$condition))
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(Microglia_NLGF-Microglia_WT, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)
tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list_pseudobulk <- list(all, up, down)
names(gsva_list_pseudobulk) <- c(
  "All DE", paste0("Upregulated in ", unique(sce$condition)[1]),
  paste0("Downregulated in ", unique(sce$condition)[1])
)
# --------------------------------------

# Fit a strict model on pseudo-bulk GSVA results.
# -----------------------------------------------
design <- model.matrix(~ 0 + sce$sample)
colnames(design) <- make.names(unique(sce$sample))
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_S305N.3-NLGF_MAPT_KI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)
tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list_pseudobulk_strict <- list(all, up, down)
names(gsva_list_pseudobulk_strict) <- c(
  "All DE", paste0("Upregulated in ", levels(sce$sample)[1]),
  paste0("Downregulated in ", levels(sce$sample)[1])
)
# -----------------------------------------------

# Fit a stricter model on pseudo-bulk GSVA results.
# -------------------------------------------------
design <- model.matrix(~ 0 + sce$sample)
colnames(design) <- make.names(unique(sce$sample))
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(S305N.3-MAPT_KI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)
tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list_pseudobulk_stricter <- list(all, up, down)
names(gsva_list_pseudobulk_stricter) <- c(
  "All DE", paste0("Upregulated in ", levels(sce$sample)[3]),
  paste0("Downregulated in ", levels(sce$sample)[3])
)
# -------------------------------------------------

# Fit the strictest model on pseudo-bulk GSVA results.
# ----------------------------------------------------
up <- intersect(intersect(rownames(gsva_list_pseudobulk[[1]]), rownames(gsva_list_pseudobulk_strict[[1]])), rownames(gsva_list_pseudobulk_stricter[[1]]))
down <- intersect(intersect(rownames(gsva_list_pseudobulk[[2]]), rownames(gsva_list_pseudobulk_strict[[2]])), rownames(gsva_list_pseudobulk_stricter[[2]]))
all <- c(up, down)
gsva_list_pseudobulk_strictest <- list(all, up, down)
# ----------------------------------------------------
```

**Upregulated in Microglia NLGF MAPT KI GSE Initial**

```{r}
ncol <- ncol(gsva_list_pseudobulk[[2]])
datatable_download_exp(gsva_list_pseudobulk[[2]][(ncol-3):ncol])
```

**Upregulated in Microglia NLGF MAPT KI GSE Strict**

```{r}
ncol <- ncol(gsva_list_pseudobulk_strict[[2]])
datatable_download_exp(gsva_list_pseudobulk_strict[[2]][(ncol-3):ncol])
```

**Upregulated in Microglia NLGF MAPT KI GSE Stricter**

```{r}
ncol <- ncol(gsva_list_pseudobulk_stricter[[2]])
datatable_download_exp(gsva_list_pseudobulk_stricter[[2]][(ncol-3):ncol])
```

**Upregulated in Microglia NLGF MAPT KI GSE Strictest**

```{r}
# ncol <- ncol(gsva_list_pseudobulk_strictest[[2]])
# datatable_download_exp(gsva_list_pseudobulk_strictest[[2]][(ncol-3):ncol])
```

**Downregulated in Microglia NLGF MAPT KI GSE Initial**

```{r}
ncol <- ncol(gsva_list_pseudobulk[[3]])
datatable_download_exp(gsva_list_pseudobulk[[3]][(ncol-3):ncol])
```

**Downregulated in Microglia NLGF MAPT KI GSE Strict**

```{r}
ncol <- ncol(gsva_list_pseudobulk_strict[[3]])
datatable_download_exp(gsva_list_pseudobulk_strict[[3]][(ncol-3):ncol])
```

**Downregulated in Microglia NLGF MAPT KI GSE Stricter**

```{r}
ncol <- ncol(gsva_list_pseudobulk_stricter[[3]])
datatable_download_exp(gsva_list_pseudobulk_stricter[[3]][(ncol-3):ncol])
```

**Downregulated in Microglia NLGF MAPT KI GSE Strictest**

```{r}
# ncol <- ncol(gsva_list_pseudobulk_strictest[[3]])
# datatable_download_exp(gsva_list_pseudobulk_strictest[[3]][(ncol-3):ncol])
```

```{r}
# Fit model on pseudo-bulk counts.
# --------------------------------
design <- model.matrix(~ 0 + sce_filt$condition)
colnames(design) <- make.names(unique(sce_filt$condition))
fit <- lmFit(logcounts(sce_filt), design)
contrast_mat <- makeContrasts(Microglia_NLGF-Microglia_WT, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)
tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
markers_list_pseudobulk <- list(all, up, down)
names(markers_list_pseudobulk) <- c(
  "All DE", paste0("Upregulated in ", unique(sce_filt$condition)[1]),
  paste0("Downregulated in ", unique(sce_filt$condition)[1])
)
# --------------------------------------

# Fit a strict model on pseudo-bulk GSVA results.
# -----------------------------------------------
design <- model.matrix(~ 0 + sce$sample)
colnames(design) <- make.names(unique(sce$sample))
fit <- lmFit(logcounts(sce_filt), design)
contrast_mat <- makeContrasts(NLGF_S305N.3-NLGF_MAPT_KI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)
tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
markers_list_pseudobulk_strict <- list(all, up, down)
names(markers_list_pseudobulk_strict) <- c(
  "All DE", paste0("Upregulated in ", unique(sce_filt$condition)[1]),
  paste0("Downregulated in ", unique(sce_filt$condition)[2])
)
# -----------------------------------------------

# Fit a stricter model on pseudo-bulk GSVA results.
# -------------------------------------------------
design <- model.matrix(~ 0 + sce$sample)
colnames(design) <- make.names(unique(sce$sample))
fit <- lmFit(logcounts(sce_filt), design)
contrast_mat <- makeContrasts(S305N.3-MAPT_KI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)
tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
markers_list_pseudobulk_stricter <- list(all, up, down)
names(markers_list_pseudobulk_stricter) <- c(
  "All DE", paste0("Upregulated in ", levels(sce_filt$sample)[3]),
  paste0("Downregulated in ", levels(sce_filt$sample)[3])
)
# -------------------------------------------------

# Fit the strictest model on pseudo-bulk GSVA results.
# ----------------------------------------------------
up <- intersect(intersect(rownames(markers_list_pseudobulk[[1]]), rownames(markers_list_pseudobulk_strict[[1]])), rownames(markers_list_pseudobulk_stricter[[1]]))
down <- intersect(intersect(rownames(markers_list_pseudobulk[[2]]), rownames(markers_list_pseudobulk_strict[[2]])), rownames(markers_list_pseudobulk_stricter[[2]]))
all <- c(up, down)
markers_list_pseudobulk_strictest <- list(all, up, down)
# ----------------------------------------------------

# Convert ENSEMBL IDs back to gene symbols.
# -----------------------------------------
genes <- gene_anno[gene_anno$ensembl_gene_id %in% rownames(markers_list_pseudobulk[[2]]), ]
genes <- genes[order(match(genes$ensembl_gene_id, rownames(markers_list_pseudobulk[[2]]))), ]
up <- markers_list_pseudobulk[[2]]
rownames(up) <- genes$external_gene_name

genes <- gene_anno[gene_anno$ensembl_gene_id %in% rownames(markers_list_pseudobulk[[3]]), ]
genes <- genes[order(match(genes$ensembl_gene_id, rownames(markers_list_pseudobulk[[3]]))), ]
down <- markers_list_pseudobulk[[3]]
rownames(down) <- genes$external_gene_name

all <- rbind(up, down)
all <- all[order(-all$B), ]
markers_list_symbols_pseudobulk <- list(all, up, down)

genes <- gene_anno[gene_anno$ensembl_gene_id %in% rownames(markers_list_pseudobulk_strict[[2]]), ]
genes <- genes[order(match(genes$ensembl_gene_id, rownames(markers_list_pseudobulk_strict[[2]]))), ]
up <- markers_list_pseudobulk_strict[[2]]
rownames(up) <- genes$external_gene_name

genes <- gene_anno[gene_anno$ensembl_gene_id %in% rownames(markers_list_pseudobulk_strict[[3]]), ]
genes <- genes[order(match(genes$ensembl_gene_id, rownames(markers_list_pseudobulk_strict[[3]]))), ]
down <- markers_list_pseudobulk_strict[[3]]
rownames(down) <- genes$external_gene_name

all <- rbind(up, down)
all <- all[order(-all$B), ]
markers_list_symbols_pseudobulk_strict <- list(all, up, down)

genes <- gene_anno[gene_anno$ensembl_gene_id %in% rownames(markers_list_pseudobulk_stricter[[2]]), ]
genes <- genes[
  order(match(genes$ensembl_gene_id, rownames(markers_list_pseudobulk_stricter[[2]]))),
]
up <- markers_list_pseudobulk_stricter[[2]]
rownames(up) <- genes$external_gene_name

genes <- gene_anno[gene_anno$ensembl_gene_id %in% rownames(markers_list_pseudobulk_stricter[[3]]), ]
genes <- genes[
  order(match(genes$ensembl_gene_id, rownames(markers_list_pseudobulk_stricter[[3]]))),
]
down <- markers_list_pseudobulk_stricter[[3]]
rownames(down) <- genes$external_gene_name

all <- rbind(up, down)
all <- all[order(-all$B), ]
markers_list_symbols_pseudobulk_stricter <- list(all, up, down)

genes <- gene_anno[
  gene_anno$ensembl_gene_id %in% markers_list_pseudobulk_strictest[[2]],
]
genes <- genes[
  order(match(genes$ensembl_gene_id, markers_list_pseudobulk_strictest[[2]])),
]
up <- markers_list_pseudobulk_strictest[[2]]
up <- genes$external_gene_name

genes <- gene_anno[
  gene_anno$ensembl_gene_id %in% markers_list_pseudobulk_strictest[[3]],
]
genes <- genes[
  order(match(genes$ensembl_gene_id, markers_list_pseudobulk_strictest[[3]])),
]
down <- markers_list_pseudobulk_strictest[[3]]
down <- genes$external_gene_name

all <- c(up, down)
markers_list_symbols_pseudobulk_strictest <- list(all, up, down)
# -----------------------------------------
```

**Upregulated in Microglia NLGF MAPT KI DEG Initial**

```{r}
ncol <- ncol(markers_list_symbols_pseudobulk[[2]])
datatable_download_exp(markers_list_symbols_pseudobulk[[2]][(ncol-3):ncol])
```

**Upregulated in Microglia NLGF MAPT KI DEG Strict**

```{r}
ncol <- ncol(markers_list_symbols_pseudobulk_strict[[2]])
datatable_download_exp(markers_list_symbols_pseudobulk_strict[[2]][(ncol-3):ncol])
```

**Upregulated in Microglia NLGF MAPT KI DEG Stricter**

```{r}
ncol <- ncol(markers_list_symbols_pseudobulk_stricter[[2]])
datatable_download_exp(markers_list_symbols_pseudobulk_stricter[[2]][(ncol-3):ncol])
```

**Upregulated in Microglia NLGF MAPT KI DEG Strictest**

```{r}
# ncol <- ncol(markers_list_symbols_pseudobulk_strictest[[2]])
# datatable_download_exp(markers_list_symbols_pseudobulk_strictest[[2]][(ncol-3):ncol])
```

**Downregulated in Microglia NLGF MAPT KI DEG Initial**

```{r}
ncol <- ncol(markers_list_symbols_pseudobulk[[3]])
datatable_download_exp(markers_list_symbols_pseudobulk[[3]][(ncol-3):ncol])
```

**Downregulated in Microglia NLGF MAPT KI DEG Strict**

```{r}
ncol <- ncol(markers_list_symbols_pseudobulk_strict[[3]])
datatable_download_exp(markers_list_symbols_pseudobulk_strict[[3]][(ncol-3):ncol])
```

**Downregulated in Microglia NLGF MAPT KI DEG Stricter**

```{r}
ncol <- ncol(markers_list_symbols_pseudobulk_stricter[[3]])
datatable_download_exp(markers_list_symbols_pseudobulk_stricter[[3]][(ncol-3):ncol])
```

**Downregulated in Microglia NLGF MAPT KI DEG Strictest**

```{r}
# ncol <- ncol(markers_list_symbols_pseudobulk_strictest[[3]])
# datatable_download_exp(markers_list_symbols_pseudobulk_strictest[[3]][(ncol-3):ncol])
```

# Gene Set Heatmap

```{r, fig.width = 10, fig.height = 20}
anno_color <- hue_pal()(length(unique(sce$sample)))
color <- colorRampPalette(rev(brewer.pal(9, "RdBu")))(100) # For main gene set heatmap.

gsva_list_pseudobulk_all <- list(
  c(
    rownames(gsva_list_pseudobulk[[1]]), rownames(gsva_list_pseudobulk_strict[[1]]),
    rownames(gsva_list_pseudobulk_stricter[[1]]), rownames(gsva_list_pseudobulk_strictest[[1]])
  ),
  c(
    rownames(gsva_list_pseudobulk[[2]]), rownames(gsva_list_pseudobulk_strict[[2]]),
    rownames(gsva_list_pseudobulk_stricter[[2]]), rownames(gsva_list_pseudobulk_strictest[[2]])
  ),
  c(
    rownames(gsva_list_pseudobulk[[3]]), rownames(gsva_list_pseudobulk_strict[[3]]),
    rownames(gsva_list_pseudobulk_stricter[[3]]), rownames(gsva_list_pseudobulk_strictest[[3]])
  )
)

markers_list_pseudobulk_all <- list(
  c(
    rownames(markers_list_pseudobulk[[1]]),
    rownames(markers_list_pseudobulk_strict[[1]]),
    rownames(markers_list_pseudobulk_stricter[[1]]),
    markers_list_pseudobulk_strictest[[1]]
  ),
  c(
    rownames(markers_list_pseudobulk[[2]]),
    rownames(markers_list_pseudobulk_strict[[2]]),
    rownames(markers_list_pseudobulk_stricter[[2]]),
    markers_list_pseudobulk_strictest[[2]]
  ),
  c(
    rownames(markers_list_pseudobulk[[3]]),
    rownames(markers_list_pseudobulk_strict[[3]]),
    rownames(markers_list_pseudobulk_stricter[[3]]),
    markers_list_pseudobulk_strictest[[3]]
  )
)
markers_list_symbols_pseudobulk_all <- list(
  c(
    rownames(markers_list_symbols_pseudobulk[[1]]),
    rownames(markers_list_symbols_pseudobulk_strict[[1]]),
    rownames(markers_list_symbols_pseudobulk_stricter[[1]]),
    markers_list_symbols_pseudobulk_strictest[[1]]
  ),
  c(
    rownames(markers_list_symbols_pseudobulk[[2]]),
    rownames(markers_list_symbols_pseudobulk_strict[[2]]),
    rownames(markers_list_symbols_pseudobulk_stricter[[2]]),
    markers_list_symbols_pseudobulk_strictest[[2]]
  ),
  c(
    rownames(markers_list_symbols_pseudobulk[[3]]),
    rownames(markers_list_symbols_pseudobulk_strict[[3]]),
    rownames(markers_list_symbols_pseudobulk_stricter[[3]]),
    markers_list_symbols_pseudobulk_strictest[[3]]
  )
)

mat <- gsva
mat <- mat[rownames(mat) %in% gsva_list_pseudobulk_all[[1]], ]
mat_up <- mat[rownames(mat) %in% gsva_list_pseudobulk_all[[2]], ]
mat_down <- mat[rownames(mat) %in% gsva_list_pseudobulk_all[[3]], ]
gene_sets_sub_up <- gene_sets[names(gene_sets) %in% rownames(mat_up)]
up_vector <- vector()
for (j in seq(length(gene_sets_sub_up@.Data))) {
  if (
    length(which(gene_sets_sub_up[[j]]@geneIds %in% markers_list_pseudobulk_all[[2]])) >=
      ceiling(length(gene_sets_sub_up[[j]]@geneIds) / length(gene_sets_sub_up[[j]]@geneIds))
    ) {
    up_vector <- append(up_vector, names(gene_sets_sub_up)[j])
  }
}
gene_sets_sub_down <- gene_sets[names(gene_sets) %in% rownames(mat_down)]
down_vector <- c()
for (j in seq(length(gene_sets_sub_down@.Data))) {
  if (
    length(which(gene_sets_sub_down[[j]]@geneIds %in% markers_list_pseudobulk_all[[3]])) >=
      ceiling(length(gene_sets_sub_down[[j]]@geneIds) / length(gene_sets_sub_down[[j]]@geneIds))
  ) {
    down_vector <- append(down_vector, names(gene_sets_sub_down)[j])
  }
}
all_vector <- append(up_vector, down_vector)
# mat <- mat[rownames(mat) %in% all_vector, ]
gene_sets_sub <- gene_sets[names(gene_sets) %in% rownames(mat)]
gene_sets_sub_ids <- gene_sets_sub
for (j in seq(length(gene_sets_sub_ids@.Data))) {
  suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
}

discard <- computeGeneSetsOverlapMax(
  gene_sets_sub_ids, unique(unlist(geneIds(gene_sets_sub_ids)))
)
discard <- colnames(discard)[colSums(discard) <= 1]
discard <- which(names(gene_sets_sub_ids) %in% discard)
mat <- mat[-discard, ]
gene_sets_sub <- gene_sets_sub[-discard, ]
gene_sets_sub_ids <- gene_sets_sub_ids[-discard, ]

gene_set_overlap <- computeGeneSetsOverlapMax(
  gene_sets_sub_ids, unique(unlist(geneIds(gene_sets_sub_ids)))
)
clustering <- hclust(get_dist(gene_set_overlap, "spearman"), "ward.D2")

anno <- HeatmapAnnotation(
  Sample = sce$sample,
  cell_type = anno_block(
    gp = gpar(fill = "lightgray", col = "lightgray"),
    labels = c(unique(sce$condition)[[1]], unique(sce$condition)[[2]]),
    labels_gp = gpar(col = "black", fontsize = 8),
    height = unit(6, "mm")
  ),
  col = list(
    Sample = c(
      "NLGF_S305N+3" = anno_color[1], "NLGF_MAPT_KI" = anno_color[2],
      "S305N+3" = anno_color[3], "MAPT_KI" = anno_color[4]
    )
  ),
  simple_anno_size = unit(2, "mm"),
  annotation_legend_param = list(Sample = list(nrow = 1))
)

modules_up <- rownames(mat)[rownames(mat) %in% gsva_list_pseudobulk_all[[2]]]
modules_down <- rownames(mat)[rownames(mat) %in% gsva_list_pseudobulk_all[[3]]]

module_gene_sets_sub <- gene_sets[names(gene_sets) %in% modules_up]
frequent_genes_tmp <- as.data.frame(table(unlist(geneIds(module_gene_sets_sub))))
frequent_genes_tmp <- frequent_genes_tmp[
  frequent_genes_tmp$Var1 %in% gene_anno$ensembl_gene_id,
]
rownames(frequent_genes_tmp) <- NULL
frequent_genes_up <- frequent_genes_tmp

module_gene_sets_sub <- gene_sets[names(gene_sets) %in% modules_down]
frequent_genes_tmp <- as.data.frame(table(unlist(geneIds(module_gene_sets_sub))))
frequent_genes_tmp <- frequent_genes_tmp[
  frequent_genes_tmp$Var1 %in% gene_anno$ensembl_gene_id,
]
rownames(frequent_genes_tmp) <- NULL
frequent_genes_down <- frequent_genes_tmp
frequent_genes <- rbind(frequent_genes_up, frequent_genes_down)

node_data <- melt(geneIds(gene_sets_sub))
node_data <- node_data[node_data$value %in% frequent_genes$Var1, ]
names(node_data)[1] <- "ensembl_gene_id"
names(node_data)[2] <- "gene_set"
node_data$direction <- ifelse(
  node_data$gene_set %in% gsva_list_pseudobulk_all[[2]], "up", "down"
)

unique_direction <- as.data.frame.matrix(table(node_data[ , c(1, 3)]), )
unique_direction$down <- unique_direction$down * -1
module_membership_norm <- unique_direction
module_membership_norm <- module_membership_norm[
  abs(rowSums(module_membership_norm)) >= 2,
]

hub_gene_all_GO <- table(unlist(geneIds(gene_sets)))
hub_gene_all_GO <- hub_gene_all_GO[
  names(hub_gene_all_GO) %in% rownames(module_membership_norm)
]

module_membership_thresh <- median(
  (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
    abs(rowSums(module_membership_norm))
)
# module_membership_thresh <- summary(
#   as.numeric(
#     (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
#       abs(rowSums(module_membership_norm))
#   )
# )[2][[1]]
module_membership_thresh <- summary(
  as.numeric(
    (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
      abs(rowSums(module_membership_norm))
  )
)[5][[1]]
# module_membership_thresh <- mean(
#   (abs(rowSums(module_membership_norm)) / hub_gene_all_GO) *
#     abs(rowSums(module_membership_norm))
# )

modules_with_hubs <- vector("list", 26)
modules_with_hubs[[1]] <- 0
hub_gene_list <- list(gene_anno)
for (l in 2:26) {
  set.seed(1)
  heatmap <- Heatmap(
    mat, cluster_rows = clustering, cluster_columns = FALSE, show_column_names = FALSE,
    show_row_names = FALSE, show_heatmap_legend = FALSE, row_split = l,
  )
  modules <- suppressWarnings(row_order(heatmap))
  hub_gene_list <- vector("list", length(modules))
  for (m in seq_along(modules)) {
    module_sets <- rownames(mat)[modules[[m]]]
    modules_up <- module_sets[module_sets %in% gsva_list_pseudobulk_all[[2]]]
    modules_down <- module_sets[module_sets %in% gsva_list_pseudobulk_all[[3]]]
    module_hub_genes <- vector()
    
    if (length(modules_up) == 0) {
    } else {
      module_gene_sets_sub <- gene_sets[names(gene_sets) %in% modules_up]
      frequent_genes <- as.data.frame(table(unlist(geneIds(module_gene_sets_sub))))
      hub_gene_all_GO_sub <- hub_gene_all_GO[
        names(hub_gene_all_GO) %in% frequent_genes$Var1
      ]
      frequent_genes <- frequent_genes[frequent_genes$Var1 %in% names(hub_gene_all_GO), ]
      frequent_genes$Freq <-
        (frequent_genes$Freq / hub_gene_all_GO_sub) * frequent_genes$Freq
      # frequent_genes <- frequent_genes[
      #   frequent_genes$Freq >= module_membership_thresh,
      # ]
      module_hub_genes_tmp <- gene_anno[
        gene_anno$ensembl_gene_id %in% frequent_genes$Var1,
      ]
      module_hub_genes_tmp <- module_hub_genes_tmp[
        module_hub_genes_tmp$ensembl_gene_id %in% markers_list_pseudobulk_all[[2]],
      ]
      module_hub_genes <- module_hub_genes_tmp$ensembl_gene_id
    }
    
    if (length(modules_down) == 0) {
    } else {
      module_gene_sets_sub <- gene_sets[names(gene_sets) %in% modules_down]
      frequent_genes <- as.data.frame(table(unlist(geneIds(module_gene_sets_sub))))
      hub_gene_all_GO_sub <- hub_gene_all_GO[
        names(hub_gene_all_GO) %in% frequent_genes$Var1
      ]
      frequent_genes <- frequent_genes[frequent_genes$Var1 %in% names(hub_gene_all_GO), ]
      frequent_genes$Freq <-
        (frequent_genes$Freq / hub_gene_all_GO_sub) * frequent_genes$Freq
      # frequent_genes <- frequent_genes[
      #   frequent_genes$Freq >= module_membership_thresh,
      # ]
      module_hub_genes_tmp <- gene_anno[
        gene_anno$ensembl_gene_id %in% frequent_genes$Var1,
      ]
      module_hub_genes_tmp <- module_hub_genes_tmp[
        module_hub_genes_tmp$ensembl_gene_id %in% markers_list_pseudobulk_all[[3]],
      ]
      module_hub_genes <- append(module_hub_genes, module_hub_genes_tmp$ensembl_gene_id)
    }
    
    hub_gene_list[[m]] <- module_hub_genes
  }
  modules_with_hubs[[l]] <- length(which(lengths(hub_gene_list) != 0)) / l
}
optimal_k <- 27 - which.max(rev(modules_with_hubs))
module_colors <- cols25(optimal_k)
clustering <- color_branches(clustering, optimal_k, col = module_colors)
set.seed(1)
heatmap <- Heatmap(
  mat, cluster_rows = clustering, cluster_columns = FALSE, show_column_names = FALSE,
  show_row_names = FALSE, show_heatmap_legend = FALSE, row_split = optimal_k
)

modules <- suppressWarnings(row_order(heatmap))
module_names <- modules
go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(mat)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]])
}
# for (l in seq_along(module_names)) {
#   module_names[[l]] <- paste0(rownames(mat)[module_names[[l]]])
# }
top_words_tmp <- data.frame(
  gene_sets = unlist(module_names), module = rep(seq_along(module_names), lapply(module_names, length))
)
top_words_list <- vector("list", length(unique(top_words_tmp$module)))
top_words <- c()
for (i in seq_along(top_words_list)) {
  top_words_tmp_sub <- top_words_tmp[top_words_tmp$module == i, ]
  top_words_list[[i]] <- vector("list", nrow(top_words_tmp_sub))
  for (j in seq_along(top_words_list[[i]])) {
    top_words_list[[i]][[j]] <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_sets[j]))
  }
  tmp <- table(strsplit(paste(unlist(top_words_list[[i]]), collapse = " "), "\\s+")[[1]])
  tmp <- names(tmp[tmp > 1])
  if (length(tmp) >= 3) {
    tmp <- tmp[1:3]
    tmp <- paste(paste0(tmp, "\n"), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp)-1)
  } else if (length(tmp > 0)) {
    add <- 3 - length(tmp)
    tmp2 <- unlist(strsplit(suppressWarnings(word_cloud(top_words_tmp_sub$gene_set)), "\\s+"))
    tmp2 <- tmp2[tmp2 %notin% tmp]
    tmp <- c(tmp, tmp2[1:add])
    tmp <- paste(paste0(tmp, "\n"), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp)-1)
  } else {
    tmp <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_set, 3))
  }
  top_words <- c(top_words, tmp)
}

top_words_all <- suppressWarnings(as.character(tapply(top_words_tmp$gene_sets, top_words_tmp$module, word_cloud)))
# top_words_old <- suppressWarnings(
#   as.character(tapply(top_words_tmp$gene_sets, top_words_tmp$module, word_cloud, width = 3))
# )
row_anno <-  HeatmapAnnotation(
  top_words = anno_block(
    gp = gpar(col = module_colors), labels = seq(optimal_k), labels_rot = 0, labels_gp = gpar(fontsize = 20)
  ),
  which = "row"
)

row_colors <- rownames(mat)
row_colors <- data.frame(
  row_colors = row_colors,
  DE_strict = row_colors %in% rownames(gsva_list_pseudobulk_strict[[1]]),
  DE_stricter = row_colors %in% rownames(gsva_list_pseudobulk_stricter[[1]]),
  DE_strictest = row_colors %in% gsva_list_pseudobulk_strictest[[1]]
) %>%
  mutate(
    type = case_when(
      DE_strictest == TRUE ~ "purple",
      DE_stricter == TRUE ~ "red",
      DE_strict == TRUE ~ "blue",
      TRUE ~ "purple"
    )
  )
row_colors <- row_colors$type

# rownames(mat)[
#   rownames(mat) =="oxidoreductase activity, acting on paired donors, with incorporation or reduction of molecular oxygen, reduced pteridine as one donor, and incorporation of one atom of oxygen"
# ] <- "oxidoreductase activity, acting on paired donors"

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    color,
    cluster_rows = clustering,
    cluster_columns = FALSE,
    row_names_max_width = max_text_width(rownames(mat), gpar(fontsize = 8)),
    show_column_names = FALSE,
    column_split = sce$condition,
    top_annotation = anno,
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    column_title = " ",
    row_split = optimal_k,
    row_title = top_words,
    row_title_rot = 0,
    row_title_gp = gpar(fontsize = 12),
    row_names_gp = gpar(fontsize = 8, col = unlist(row_colors)),
    left_annotation = row_anno,
    row_gap = unit(1, "mm"),
    column_gap = unit(1, "mm"),
    heatmap_legend_param = list(
      title = "scaled ssGSEA score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
