---
title: "Combined Samples - 01 Prep"
author:
  - name: "Emir Turkes"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_file = file.path(
      "..", "..", "..", "results",
      unlist(strsplit(getwd(), "/"))[6], unlist(strsplit(getwd(), "/"))[7], "comb_01_prep.html"
  )}))
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [tau-mutant-omics-analysis](https://github.com/eturkes/tau-mutant-omics-analysis).*

The table of contents in the top left is clickable and can be used to quickly navigate the document.
To toggle the visibility of code, use the `CODE` toggles at the top right of chunks.
The toggle at the start of the document controls the visibility of all chunks.

```{r}
#    This file is part of tau-mutant-omics-analysis.
#    Copyright (C) 2023  Emir Turkes, Naoto Watamura, UK DRI at UCL
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# Load required packages, suppressing startup messages.
# -----------------------------------------------------
library(conflicted)
conflicts_prefer(base::intersect, .quiet = TRUE)
packages <- c("Seurat", "SingleCellExperiment", "scuttle", "parallelly", "ggplot2", "dplyr", "ggrepel", "viridis")
invisible(suppressPackageStartupMessages(lapply(packages, FUN = library, character.only = TRUE)))
# -----------------------------------------------------

# Define global settings.
# -----------------------
knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# -----------------------

# Define functions.
# -----------------
source(file.path("..", "..", "utils.R"))
# -----------------

# Useful variables.
# -----------------
analysis_no <- 1
data_name <- unlist(strsplit(getwd(), "/"))[7] # Name of dataset derived from path.

cache_dir <- file.path("..", "..", "..", "cache", "low-coverage", data_name, paste0("0", analysis_no))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

protocol <- c("mouse", "droplet", "single-nuc", "umis") # See `cluster_pipeline` in `utils.R`.
vars_to_regress <- NULL # See `cluster_pipeline` in `utils.R`.
parallel_override <- 0 # See `parallel_plan` in `utils.R`.
# -----------------

# Metadata to plot after dimensionality reduction and clustering.
# Values in list can include "no_legend and/or "no_label" to exclude those.
# -------------------------------------------------------------------------
metadata_to_plot <- vector("list", length = 3)
names(metadata_to_plot) <- c("seurat_clusters", "sample", "doublets")
metadata_to_plot[[2]] <- "no_label"
metadata_to_plot[[3]] <- "no_label"
# -------------------------------------------------------------------------
```

# Merging

```{r}
rds <- file.path(cache_dir, "merged_seurat.rds")
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {

  file_list <- list.files(file.path(cache_dir, "..", ".."), recursive = TRUE)
  file_list <- file_list[grep("seurat_doubletfinder.rds", file_list)]
  files <- vector("list", length(file_list))
  for (i in seq_along(files)) {
    files[[i]] <- readRDS(file.path(cache_dir, "..", "..", file_list[i]))
  }

  genes <- Reduce(
    intersect,
    list(
      rownames(files[[1]]), rownames(files[[2]]), rownames(files[[3]]),
      rownames(files[[4]]), rownames(files[[5]]), rownames(files[[6]])
    )
  )
  for (i in seq_along(files)) {
    DefaultAssay(files[[i]]) <- "RNA"
    files[[i]][["SCT"]] <- NULL
  }

  seurat <- merge(files[[1]], c(files[[2]], files[[3]], files[[4]], files[[5]], files[[6]]))
  seurat@meta.data[grep("SCT", colnames(seurat@meta.data))] <- NULL
  seurat@meta.data[grep("pANN", colnames(seurat@meta.data))] <- NULL
  seurat@meta.data[grep("DF", colnames(seurat@meta.data))] <- NULL

  seurat$sample <- c(
    rep(sub("\\/.*", "", file_list[1]), ncol(files[[1]])),
    rep(sub("\\/.*", "", file_list[2]), ncol(files[[2]])),
    rep(sub("\\/.*", "", file_list[3]), ncol(files[[3]])),
    rep(sub("\\/.*", "", file_list[4]), ncol(files[[4]])),
    rep(sub("\\/.*", "", file_list[5]), ncol(files[[5]])),
    rep(sub("\\/.*", "", file_list[6]), ncol(files[[6]]))
  )

  ribo <- grep("^RP[SL][[:digit:]]", rownames(seurat), ignore.case = TRUE)
  sce <- as.SingleCellExperiment(seurat)
  sce <- addPerCellQC(sce, list(ribo = ribo))
  seurat$subsets_ribo_percent <- sce$subsets_ribo_percent
  rm(sce)

  saveRDS(seurat, file = rds)
}
```

# Cluster Pipeline

```{r, fig.width = 20, fig.height = 14}
sub_name <- "processed_merged"
seurat <- cluster_pipeline(
  seurat, cache_dir = cache_dir, sub_name = sub_name, protocol = protocol, vars_to_regress = vars_to_regress,
  parallel_override = parallel_override, cc = FALSE, res_divider = 1000000, conserve_memory = TRUE
)
seurat

for (i in 1:length(metadata_to_plot)) {
  if ("no_legend" %in% metadata_to_plot[[i]] && "no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i], type = "cat") +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i], type = "cat") +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(red_dim_plot(seurat, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont") + ggtitle("nFeature_SCT"))
  print(red_dim_plot(seurat, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont") + ggtitle("nCount_SCT"))
} else if (protocol[4] == "reads") {
  print(red_dim_plot(seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont") + ggtitle("nFeature_RNA"))
  print(red_dim_plot(seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont") + ggtitle("nCount_RNA"))
}

red_dim_plot(seurat, x = "umap1", y = "umap2", color = "subsets_mito_percent", type = "cont") +
  ggtitle("subsets_mito_percent")
red_dim_plot(seurat, x = "umap1", y = "umap2", color = "subsets_ribo_percent", type = "cont") +
  ggtitle("subsets_ribo_percent")
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
