---
title: "Comb - Batch 02 - 01 GeneFunnel"
author:
  - name: "Emir Turkes"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_file = file.path("..", "results", "comb_batch02_01genefunnel.html")
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 35px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [tau-mutant-omics-analysis](https://github.com/eturkes/tau-mutant-omics-analysis).*

The table of contents in the top left is clickable and can be used to quickly navigate the document.
To toggle the visibility of code, use the `CODE` toggles at the top right of chunks.
The toggle at the start of the document controls the visibility of all chunks.

```{r}
#    This file is part of tau-mutant-omics-analysis.
#    Copyright (C) 2023  Emir Turkes, Naoto Watamura, UK DRI at UCL
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# Load required packages, suppressing startup messages.
# -----------------------------------------------------
library(conflicted)
conflicts_prefer(DT::JS, .quiet = TRUE)
packages <- c("Seurat", "limma", "DT", "RColorBrewer", "ComplexHeatmap")
invisible(suppressPackageStartupMessages(lapply(packages, FUN = library, character.only = TRUE)))
# -----------------------------------------------------

# Define global settings.
# -----------------------
knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# -----------------------

# Define functions.
# -----------------
source("utils.R")
# -----------------

# Useful variables.
# -----------------
genotype <- "comb"
batch <- 2
step <- 1

cache_dir <- file.path("..", "cache", genotype, paste0("0", batch), paste0("0", step))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

colour <- colorRampPalette(rev(brewer.pal(9, "RdBu")))(100)
# -----------------

# Useful data.
# ------------
gene_anno <- readRDS(file.path("..", "cache", "MAPTKI", "02", "02", "gene_anno.rds"))
# ------------
```

# Merging

```{r}
rds <- file.path(cache_dir, "seurat_merged.rds")
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {

  file_list <- list.files(file.path(cache_dir, "..", "..", ".."), recursive = TRUE)
  file_list <- file_list[grep("seurat_cleaned.rds", x = file_list)]
  file_list <- c(file_list[1], file_list[6], file_list[5], file_list[2], file_list[4], file_list[3])
  files <- vector("list", length = length(file_list))
  for (i in seq_along(files)) {
    files[[i]] <- readRDS(file.path(cache_dir, "..", "..", "..", file_list[i]))
  }

  for (i in seq_along(files)) {
    data_mat <- GetAssayData(files[[i]], layer = "data")
    data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
    rownames(data_mat) <- gene_anno$external_gene_name
    files[i] <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = files[[i]][[]]) 
  }
  rm(data_mat)

  seurat <- Reduce(function(...) merge(...), x = files)
  seurat@meta.data[grep("SCT", colnames(seurat@meta.data))] <- NULL
  seurat <- JoinLayers(seurat)

  seurat$sample <- c(
    rep(sub("\\/.*", "", file_list[1]), times = ncol(files[[1]])),
    rep(sub("\\/.*", "", file_list[2]), times = ncol(files[[2]])),
    rep(sub("\\/.*", "", file_list[3]), times = ncol(files[[3]])),
    rep(sub("\\/.*", "", file_list[4]), times = ncol(files[[4]])),
    rep(sub("\\/.*", "", file_list[5]), times = ncol(files[[5]])),
    rep(sub("\\/.*", "", file_list[6]), times = ncol(files[[6]]))
  )

  file_list <- list.files(file.path(cache_dir, "..", "..", ".."), recursive = TRUE)
  file_list <- file_list[grep("gsva.rds", x = file_list)]
  file_list <- c(file_list[1], file_list[6], file_list[5], file_list[2], file_list[4], file_list[3])
  files <- vector("list", length = length(file_list))
  for (i in seq_along(files)) {
    files[[i]] <- readRDS(file.path(cache_dir, "..", "..", "..", file_list[i]))
  }

  gsva <- Reduce(function(...) cbind(...), x = files)
  rm(files)
  colnames(gsva) <- colnames(seurat)

  seurat[["GSVA"]] <- CreateAssay5Object(gsva)
  rm(gsva)

  saveRDS(seurat, file = rds)
}
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
