---
title: "Comb - Batch 02 - 01 GeneFunnel"
author:
  - name: "Emir Turkes"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_file = file.path("..", "results", "comb_batch02_01genefunnel.html")
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 35px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [tau-mutant-omics-analysis](https://github.com/eturkes/tau-mutant-omics-analysis).*

The table of contents in the top left is clickable and can be used to quickly navigate the document.
To toggle the visibility of code, use the `CODE` toggles at the top right of chunks.
The toggle at the start of the document controls the visibility of all chunks.

```{r}
#    This file is part of tau-mutant-omics-analysis.
#    Copyright (C) 2023-2024  Emir Turkes, Naoto Watamura, UK DRI at UCL
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# Load required packages, suppressing startup messages.
# -----------------------------------------------------
library(conflicted)
conflicts_prefer(AnnotationDbi::select, edgeR::cpm, DT::JS, .quiet = TRUE)
packages <- c(
  "GSEABase", "Seurat", "ggplot2", "dplyr", "ggrepel", "scuttle", "GO.db", "tm", "pals", "ComplexHeatmap",
  "RColorBrewer", "edgeR", "DT"
)
invisible(suppressPackageStartupMessages(lapply(packages, FUN = library, character.only = TRUE)))
# -----------------------------------------------------

# Define global settings.
# -----------------------
knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# -----------------------

# Define functions.
# -----------------
source("utils.R")

`%notin%` <- Negate(`%in%`)
# -----------------

# Useful variables.
# -----------------
genotype <- "comb"
batch <- 2
step <- 1

cache_dir <- file.path("..", "cache", genotype, paste0("0", batch), paste0("0", step))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

colour <- colorRampPalette(rev(brewer.pal(9, "RdBu")))(100)
# -----------------

# Useful data.
# ------------
gene_anno <- readRDS(file.path("..", "cache", "MAPTKI", "02", "02", "gene_anno.rds"))
gene_sets <- readRDS(file.path("..", "cache", "filtered_gene_sets.rds"))
# ------------
```

# Merging

```{r}
rds <- file.path(cache_dir, "seurat_merged.rds")
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {

  file_list <- list.files(file.path(cache_dir, "..", "..", ".."), recursive = TRUE)
  file_list <- file_list[grep("seurat_cleaned.rds", x = file_list)]
  file_list <- c(file_list[1], file_list[6], file_list[5], file_list[2], file_list[4], file_list[3])
  files <- vector("list", length = length(file_list))
  for (i in seq_along(files)) {
    files[[i]] <- readRDS(file.path(cache_dir, "..", "..", "..", file_list[i]))
  }

  for (i in seq_along(files)) {
    counts_mat <- GetAssayData(files[[i]], layer = "counts")
    counts_mat <- counts_mat[match(gene_anno$ensembl_gene_id, rownames(counts_mat)), ]
    rownames(counts_mat) <- gene_anno$external_gene_name

    data_mat <- GetAssayData(files[[i]], layer = "data")
    data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
    rownames(data_mat) <- gene_anno$external_gene_name

    files[i] <- CreateSeuratObject(counts = counts_mat, data = data_mat, assay = "SCT", meta.data = files[[i]][[]])
  }

  seurat <- Reduce(function(...) merge(...), x = files)
  seurat@meta.data[grep("SCT", colnames(seurat@meta.data))] <- NULL
  seurat@meta.data[grep("orig.ident", colnames(seurat@meta.data))] <- NULL
  seurat@meta.data[grep("pca", colnames(seurat@meta.data))] <- NULL
  seurat@meta.data[grep("umap", colnames(seurat@meta.data))] <- NULL
  seurat@meta.data[grep("seurat_clusters", colnames(seurat@meta.data))] <- NULL
  seurat@meta.data[grep("allen_labels", colnames(seurat@meta.data))] <- NULL
  seurat <- JoinLayers(seurat)

  seurat$sample <- c(
    rep(sub("\\/.*", "", file_list[1]), times = ncol(files[[1]])),
    rep(sub("\\/.*", "", file_list[2]), times = ncol(files[[2]])),
    rep(sub("\\/.*", "", file_list[3]), times = ncol(files[[3]])),
    rep(sub("\\/.*", "", file_list[4]), times = ncol(files[[4]])),
    rep(sub("\\/.*", "", file_list[5]), times = ncol(files[[5]])),
    rep(sub("\\/.*", "", file_list[6]), times = ncol(files[[6]]))
  )
  seurat$sample <- factor(seurat$sample, levels = unique(seurat$sample))
  seurat@active.ident <- seurat$sample

  file_list <- list.files(file.path(cache_dir, "..", "..", ".."), recursive = TRUE)
  file_list <- file_list[grep("gsva.rds", x = file_list)]
  file_list <- c(file_list[1], file_list[6], file_list[5], file_list[2], file_list[4], file_list[3])
  files <- vector("list", length = length(file_list))
  for (i in seq_along(files)) {
    files[[i]] <- readRDS(file.path(cache_dir, "..", "..", "..", file_list[i]))
  }

  gsva <- Reduce(function(...) cbind(...), x = files)
  colnames(gsva) <- colnames(seurat)
  min <- abs(min(gsva))
  gsva <- gsva + min
  seurat[["GSVA"]] <- CreateAssay5Object(counts = gsva, data = log1p(gsva))

  rm(counts_mat, data_mat, files, gsva)
  saveRDS(seurat, file = rds)
}
```

# Analysis

```{r}
rds <- file.path(cache_dir, "seurat_analysed.rds")
if (file.exists(rds)) {
  seurat <- readRDS(rds)
} else {

  seurat <- ScaleData(seurat, verbose = FALSE)

  seurat <- RunPCA(seurat, features = rownames(seurat), reduction.name = "SCT_pca", verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "SCT_pca"))
  seurat$SCT_pca1 <- add_df[ , 1]
  seurat$SCT_pca2 <- add_df[ , 2]

  seurat <- RunUMAP(seurat, dims = 1:30, reduction = "SCT_pca", reduction.name = "SCT_umap", verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "SCT_umap"))
  seurat$SCT_umap1 <- add_df[ , 1]
  seurat$SCT_umap2 <- add_df[ , 2]

  DefaultAssay(seurat) <- "GSVA"
  seurat <- ScaleData(seurat, verbose = FALSE)

  seurat <- RunPCA(seurat, features = rownames(seurat), reduction.name = "GSVA_pca", verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "GSVA_pca"))
  seurat$GSVA_pca1 <- add_df[ , 1]
  seurat$GSVA_pca2 <- add_df[ , 2]

  seurat <- RunUMAP(seurat, dims = 1:30, reduction = "GSVA_pca", reduction.name = "GSVA_umap", verbose = FALSE)
  add_df <- data.frame(Embeddings(seurat, reduction = "GSVA_umap"))
  seurat$GSVA_umap1 <- add_df[ , 1]
  seurat$GSVA_umap2 <- add_df[ , 2]

  saveRDS(seurat, file = rds)
}

red_dim_plot(seurat, x = "SCT_umap1", y = "SCT_umap2", color = "sample")
red_dim_plot(seurat, x = "GSVA_umap1", y = "GSVA_umap2", color = "sample", type = "cat")
```

# GSE

## All Cell-types

```{r}
sce <- as.SingleCellExperiment(seurat, assay = "GSVA")
sce <- suppressWarnings(
  aggregateAcrossCells(sce, colData(sce)[ , "sample"], use_exprs_values = "counts", statistics = "sum")
)
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

rds <- file.path(cache_dir, "markers_gsva_all.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
markers <- markers %>% distinct(gene, .keep_all = TRUE)
rownames(markers) <- markers$gene
datatable_download_exp(markers)
markers$number <- seq(rownames(markers))

modules <- markers %>% group_by(cluster) %>% summarise(named_vec = list(number)) %>% tibble::deframe()
module_names <- modules
gene_sets_sub_ids <- gene_sets[names(gene_sets) %in% rownames(markers)]
gene_sets_sub_ids <- gene_sets_sub_ids[match(rownames(markers), names(gene_sets_sub_ids)), ]

for (j in seq(length(gene_sets_sub_ids@.Data))) {
  suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
}
go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(markers)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]])
}
top_words_tmp <- data.frame(
  gene_sets = unlist(module_names), module = rep(seq_along(module_names), lapply(module_names, length))
)
top_words_list <- vector("list", length = length(unique(top_words_tmp$module)))
top_words <- c()

for (i in seq_along(top_words_list)) {
  top_words_tmp_sub <- top_words_tmp[top_words_tmp$module == i, ]
  top_words_list[[i]] <- vector("list", nrow(top_words_tmp_sub))
  for (j in seq_along(top_words_list[[i]])) {
    top_words_list[[i]][[j]] <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_sets[j]))
  }
  tmp <- strsplit(paste(unlist(top_words_list[[i]]), collapse = " "), "\\s+")[[1]]
  remove <- which(tmp == "")
  if (length(remove) > 0) {
    tmp <- tmp[-which(tmp == "")]
  }
  tmp <- table(tmp)
  for (j in seq_along(tmp)) {
    log2FC <- grep(paste0("\\b", names(tmp)[j], "\\b"), unlist(top_words_list[[i]]))
    tmp[j] <- sum(markers$avg_log2FC[log2FC])
  }
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  dup <- which(duplicated(tmp))
  if (length(dup) > 0) {
    tmp <- tmp[-dup]
  }
  tmp <- names(tmp)
  if (length(tmp) >= 3) {
    tmp <- tmp[1:3]
    tmp <- paste(paste0(tmp, " "), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else if (length(tmp > 0)) {
    add <- 3 - length(tmp)
    tmp2 <- unlist(strsplit(suppressWarnings(word_cloud(top_words_tmp_sub$gene_set)), "\\s+"))
    tmp2 <- tmp2[tmp2 %notin% tmp]
    tmp <- c(tmp, tmp2[1:add])
    if (any(is.na(tmp))) {
      tmp <- tmp[-which(is.na(tmp))]
    }
    tmp <- paste(paste0(tmp, " "), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else {
    tmp <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_set, 3))
    tmp <- gsub("\n", " ", tmp)
  }
  top_words <- c(top_words, tmp)
}

optimal_k <- unlist(lapply(seq_along(modules), function(i) {rep(names(modules)[i], length(modules[[i]]))}))
optimal_k <- factor(optimal_k, levels = unique(optimal_k))
module_colours <- cols25(length(unique(optimal_k)))

row_anno <-  HeatmapAnnotation(
  top_words = anno_block(
    gp = gpar(col = module_colours), labels = seq(length(unique(optimal_k))),
    labels_rot = 0, labels_gp = gpar(fontsize = 12)
  ),
  which = "row"
)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    row_split = optimal_k,
    left_annotation = row_anno,
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title = top_words,
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled ssGSEA Score", title_position = "lefttop-rot")
  ),
  column_title = "All Cell-types"
)
```

### PCA

```{r}
red_dim_plot(seurat, x = "GSVA_pca1", y = "GSVA_pca2", color = "sample", type = "cat")

markers <- as.data.frame(Loadings(seurat, reduction = "GSVA_pca"))
markers <- markers[ , 1:2]
q1 <- quantile(markers[ , 1])[4]
q2 <- quantile(markers[ , 2])[2]
markers <- markers[markers[ , 1] > q1, ]
markers <- markers[markers[ , 2] < q2, ]
markers <- abs(markers)

markers$total <- rowSums(markers)
markers <- markers[order(markers$total, decreasing = TRUE), ]
markers$gene <- rownames(markers)
datatable_download(markers)
markers$cluster <- rep("0", times = nrow(markers))
markers$number <- seq(rownames(markers))
```

```{r, fig.width = 10}
FeaturePlot(
  seurat, features = rownames(markers)[1:4], reduction = "GSVA_pca",
  order = TRUE, cols = c("lightgrey", "red")
)
```

```{r}
modules <- markers %>% group_by(cluster) %>% summarise(named_vec = list(number)) %>% tibble::deframe()
module_names <- modules
gene_sets_sub_ids <- gene_sets[names(gene_sets) %in% rownames(markers)]
gene_sets_sub_ids <- gene_sets_sub_ids[match(rownames(markers), names(gene_sets_sub_ids)), ]

for (j in seq(length(gene_sets_sub_ids@.Data))) {
  suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
}
go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(markers)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]])
}
top_words_tmp <- data.frame(
  gene_sets = unlist(module_names), module = rep(seq_along(module_names), lapply(module_names, length))
)
top_words_list <- vector("list", length = length(unique(top_words_tmp$module)))
top_words <- c()

for (i in seq_along(top_words_list)) {
  top_words_tmp_sub <- top_words_tmp[top_words_tmp$module == i, ]
  top_words_list[[i]] <- vector("list", nrow(top_words_tmp_sub))
  for (j in seq_along(top_words_list[[i]])) {
    top_words_list[[i]][[j]] <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_sets[j]))
  }
  tmp <- strsplit(paste(unlist(top_words_list[[i]]), collapse = " "), "\\s+")[[1]]
  remove <- which(tmp == "")
  if (length(remove) > 0) {
    tmp <- tmp[-which(tmp == "")]
  }
  tmp <- table(tmp)
  for (j in seq_along(tmp)) {
    total <- grep(paste0("\\b", names(tmp)[j], "\\b"), unlist(top_words_list[[i]]))
    tmp[j] <- sum(markers$total[total])
  }
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  dup <- which(duplicated(tmp))
  if (length(dup) > 0) {
    tmp <- tmp[-dup]
  }
  tmp <- names(tmp)
  if (length(tmp) >= 10) {
    tmp <- tmp[1:10]
    tmp <- paste(paste0(tmp, "\n"), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else if (length(tmp > 0)) {
    add <- 10 - length(tmp)
    tmp2 <- unlist(strsplit(suppressWarnings(word_cloud(top_words_tmp_sub$gene_set)), "\\s+"))
    tmp2 <- tmp2[tmp2 %notin% tmp]
    tmp <- c(tmp, tmp2[1:add])
    if (any(is.na(tmp))) {
      tmp <- tmp[-which(is.na(tmp))]
    }
    tmp <- paste(paste0(tmp, "\n"), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else {
    tmp <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_set, 10))
  }
  top_words <- c(top_words, tmp)
}

optimal_k <- unlist(lapply(seq_along(modules), function(i) {rep(names(modules)[i], length(modules[[i]]))}))
optimal_k <- factor(optimal_k, levels = unique(optimal_k))
module_colours <- cols25(length(unique(optimal_k)))

row_anno <-  HeatmapAnnotation(
  top_words = anno_block(
    gp = gpar(col = module_colours), labels = seq(length(unique(optimal_k))),
    labels_rot = 0, labels_gp = gpar(fontsize = 12)
  ),
  which = "row"
)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    row_split = optimal_k,
    left_annotation = row_anno,
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title = top_words,
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled ssGSEA Score", title_position = "lefttop-rot")
  ),
  column_title = "All Cell-types PCA"
)
```

## Astrocyte

```{r}
seurat_sub <- seurat[ , which(seurat$annotations_short == "Astrocyte")]

sce <- as.SingleCellExperiment(seurat_sub, assay = "GSVA")
sce <- suppressWarnings(
  aggregateAcrossCells(sce, colData(sce)[ , "sample"], use_exprs_values = "counts", statistics = "sum")
)
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

rds <- file.path(cache_dir, "markers_gsva_astro.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat_sub, only.pos = TRUE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
markers <- markers %>% distinct(gene, .keep_all = TRUE)
rownames(markers) <- markers$gene
datatable_download_exp(markers)
markers$number <- seq(rownames(markers))

modules <- markers %>% group_by(cluster) %>% summarise(named_vec = list(number)) %>% tibble::deframe()
module_names <- modules
gene_sets_sub_ids <- gene_sets[names(gene_sets) %in% rownames(markers)]
gene_sets_sub_ids <- gene_sets_sub_ids[match(rownames(markers), names(gene_sets_sub_ids)), ]

for (j in seq(length(gene_sets_sub_ids@.Data))) {
  suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
}
go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(markers)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]])
}
top_words_tmp <- data.frame(
  gene_sets = unlist(module_names), module = rep(seq_along(module_names), lapply(module_names, length))
)
top_words_list <- vector("list", length = length(unique(top_words_tmp$module)))
top_words <- c()

for (i in seq_along(top_words_list)) {
  top_words_tmp_sub <- top_words_tmp[top_words_tmp$module == i, ]
  top_words_list[[i]] <- vector("list", nrow(top_words_tmp_sub))
  for (j in seq_along(top_words_list[[i]])) {
    top_words_list[[i]][[j]] <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_sets[j]))
  }
  tmp <- strsplit(paste(unlist(top_words_list[[i]]), collapse = " "), "\\s+")[[1]]
  remove <- which(tmp == "")
  if (length(remove) > 0) {
    tmp <- tmp[-which(tmp == "")]
  }
  tmp <- table(tmp)
  for (j in seq_along(tmp)) {
    log2FC <- grep(paste0("\\b", names(tmp)[j], "\\b"), unlist(top_words_list[[i]]))
    tmp[j] <- sum(markers$avg_log2FC[log2FC])
  }
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  dup <- which(duplicated(tmp))
  if (length(dup) > 0) {
    tmp <- tmp[-dup]
  }
  tmp <- names(tmp)
  if (length(tmp) >= 3) {
    tmp <- tmp[1:3]
    tmp <- paste(paste0(tmp, " "), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else if (length(tmp > 0)) {
    add <- 3 - length(tmp)
    tmp2 <- unlist(strsplit(suppressWarnings(word_cloud(top_words_tmp_sub$gene_set)), "\\s+"))
    tmp2 <- tmp2[tmp2 %notin% tmp]
    tmp <- c(tmp, tmp2[1:add])
    if (any(is.na(tmp))) {
      tmp <- tmp[-which(is.na(tmp))]
    }
    tmp <- paste(paste0(tmp, " "), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else {
    tmp <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_set, 3))
    tmp <- gsub("\n", " ", tmp)
  }
  top_words <- c(top_words, tmp)
}

optimal_k <- unlist(lapply(seq_along(modules), function(i) {rep(names(modules)[i], length(modules[[i]]))}))
optimal_k <- factor(optimal_k, levels = unique(optimal_k))
module_colours <- cols25(length(unique(optimal_k)))

row_anno <-  HeatmapAnnotation(
  top_words = anno_block(
    gp = gpar(col = module_colours), labels = seq(length(unique(optimal_k))),
    labels_rot = 0, labels_gp = gpar(fontsize = 12)
  ),
  which = "row"
)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    row_split = optimal_k,
    left_annotation = row_anno,
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title = top_words,
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled ssGSEA Score", title_position = "lefttop-rot")
  ),
  column_title = "Astrocyte"
)
```

### PCA

```{r}
seurat_sub <- RunPCA(seurat_sub, features = rownames(seurat_sub), reduction.name = "GSVA_pca", verbose = FALSE)
add_df <- data.frame(Embeddings(seurat_sub, reduction = "GSVA_pca"))
seurat_sub$GSVA_pca1 <- add_df[ , 1]
seurat_sub$GSVA_pca2 <- add_df[ , 2]

red_dim_plot(seurat_sub, x = "GSVA_pca1", y = "GSVA_pca2", color = "sample", type = "cat")

markers <- as.data.frame(Loadings(seurat_sub, reduction = "GSVA_pca"))
markers <- markers[ , 1:2]
q1 <- quantile(markers[ , 1])[2]
q2 <- quantile(markers[ , 2])[4]
markers <- markers[markers[ , 1] < q1, ]
markers <- markers[markers[ , 2] > q2, ]
markers <- abs(markers)

markers$total <- rowSums(markers)
markers <- markers[order(markers$total, decreasing = TRUE), ]
markers$gene <- rownames(markers)
datatable_download(markers)
markers$cluster <- rep("0", times = nrow(markers))
markers$number <- seq(rownames(markers))
```

```{r, fig.width = 20}
FeaturePlot(
  seurat_sub, features = rownames(markers)[1:4], reduction = "GSVA_pca",
  order = TRUE, cols = c("lightgrey", "red")
)
```

```{r}
modules <- markers %>% group_by(cluster) %>% summarise(named_vec = list(number)) %>% tibble::deframe()
module_names <- modules
gene_sets_sub_ids <- gene_sets[names(gene_sets) %in% rownames(markers)]
gene_sets_sub_ids <- gene_sets_sub_ids[match(rownames(markers), names(gene_sets_sub_ids)), ]

for (j in seq(length(gene_sets_sub_ids@.Data))) {
  suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
}
go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(markers)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]])
}
top_words_tmp <- data.frame(
  gene_sets = unlist(module_names), module = rep(seq_along(module_names), lapply(module_names, length))
)
top_words_list <- vector("list", length = length(unique(top_words_tmp$module)))
top_words <- c()

for (i in seq_along(top_words_list)) {
  top_words_tmp_sub <- top_words_tmp[top_words_tmp$module == i, ]
  top_words_list[[i]] <- vector("list", nrow(top_words_tmp_sub))
  for (j in seq_along(top_words_list[[i]])) {
    top_words_list[[i]][[j]] <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_sets[j]))
  }
  tmp <- strsplit(paste(unlist(top_words_list[[i]]), collapse = " "), "\\s+")[[1]]
  remove <- which(tmp == "")
  if (length(remove) > 0) {
    tmp <- tmp[-which(tmp == "")]
  }
  tmp <- table(tmp)
  for (j in seq_along(tmp)) {
    total <- grep(paste0("\\b", names(tmp)[j], "\\b"), unlist(top_words_list[[i]]))
    tmp[j] <- sum(markers$total[total])
  }
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  dup <- which(duplicated(tmp))
  if (length(dup) > 0) {
    tmp <- tmp[-dup]
  }
  tmp <- names(tmp)
  if (length(tmp) >= 10) {
    tmp <- tmp[1:10]
    tmp <- paste(paste0(tmp, "\n"), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else if (length(tmp > 0)) {
    add <- 10 - length(tmp)
    tmp2 <- unlist(strsplit(suppressWarnings(word_cloud(top_words_tmp_sub$gene_set)), "\\s+"))
    tmp2 <- tmp2[tmp2 %notin% tmp]
    tmp <- c(tmp, tmp2[1:add])
    if (any(is.na(tmp))) {
      tmp <- tmp[-which(is.na(tmp))]
    }
    tmp <- paste(paste0(tmp, "\n"), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else {
    tmp <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_set, 10))
  }
  top_words <- c(top_words, tmp)
}

optimal_k <- unlist(lapply(seq_along(modules), function(i) {rep(names(modules)[i], length(modules[[i]]))}))
optimal_k <- factor(optimal_k, levels = unique(optimal_k))
module_colours <- cols25(length(unique(optimal_k)))

row_anno <-  HeatmapAnnotation(
  top_words = anno_block(
    gp = gpar(col = module_colours), labels = seq(length(unique(optimal_k))),
    labels_rot = 0, labels_gp = gpar(fontsize = 12)
  ),
  which = "row"
)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    row_split = optimal_k,
    left_annotation = row_anno,
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title = top_words,
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled ssGSEA Score", title_position = "lefttop-rot")
  ),
  column_title = "Astrocyte PCA"
)
```

## Cortical Glutamatergic

```{r}
seurat_sub <- seurat[ , which(seurat$annotations_short == "Cortical Glutamatergic")]

sce <- as.SingleCellExperiment(seurat_sub, assay = "GSVA")
sce <- suppressWarnings(
  aggregateAcrossCells(sce, colData(sce)[ , "sample"], use_exprs_values = "counts", statistics = "sum")
)
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

rds <- file.path(cache_dir, "markers_gsva_glut.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat_sub, only.pos = TRUE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
markers <- markers %>% distinct(gene, .keep_all = TRUE)
rownames(markers) <- markers$gene
datatable_download_exp(markers)
markers$number <- seq(rownames(markers))

modules <- markers %>% group_by(cluster) %>% summarise(named_vec = list(number)) %>% tibble::deframe()
module_names <- modules
gene_sets_sub_ids <- gene_sets[names(gene_sets) %in% rownames(markers)]
gene_sets_sub_ids <- gene_sets_sub_ids[match(rownames(markers), names(gene_sets_sub_ids)), ]

for (j in seq(length(gene_sets_sub_ids@.Data))) {
  suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
}
go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(markers)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]])
}
top_words_tmp <- data.frame(
  gene_sets = unlist(module_names), module = rep(seq_along(module_names), lapply(module_names, length))
)
top_words_list <- vector("list", length = length(unique(top_words_tmp$module)))
top_words <- c()

for (i in seq_along(top_words_list)) {
  top_words_tmp_sub <- top_words_tmp[top_words_tmp$module == i, ]
  top_words_list[[i]] <- vector("list", nrow(top_words_tmp_sub))
  for (j in seq_along(top_words_list[[i]])) {
    top_words_list[[i]][[j]] <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_sets[j]))
  }
  tmp <- strsplit(paste(unlist(top_words_list[[i]]), collapse = " "), "\\s+")[[1]]
  remove <- which(tmp == "")
  if (length(remove) > 0) {
    tmp <- tmp[-which(tmp == "")]
  }
  tmp <- table(tmp)
  for (j in seq_along(tmp)) {
    log2FC <- grep(paste0("\\b", names(tmp)[j], "\\b"), unlist(top_words_list[[i]]))
    tmp[j] <- sum(markers$avg_log2FC[log2FC])
  }
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  dup <- which(duplicated(tmp))
  if (length(dup) > 0) {
    tmp <- tmp[-dup]
  }
  tmp <- names(tmp)
  if (length(tmp) >= 3) {
    tmp <- tmp[1:3]
    tmp <- paste(paste0(tmp, " "), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else if (length(tmp > 0)) {
    add <- 3 - length(tmp)
    tmp2 <- unlist(strsplit(suppressWarnings(word_cloud(top_words_tmp_sub$gene_set)), "\\s+"))
    tmp2 <- tmp2[tmp2 %notin% tmp]
    tmp <- c(tmp, tmp2[1:add])
    if (any(is.na(tmp))) {
      tmp <- tmp[-which(is.na(tmp))]
    }
    tmp <- paste(paste0(tmp, " "), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else {
    tmp <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_set, 3))
    tmp <- gsub("\n", " ", tmp)
  }
  top_words <- c(top_words, tmp)
}

optimal_k <- unlist(lapply(seq_along(modules), function(i) {rep(names(modules)[i], length(modules[[i]]))}))
optimal_k <- factor(optimal_k, levels = unique(optimal_k))
module_colours <- cols25(length(unique(optimal_k)))

row_anno <-  HeatmapAnnotation(
  top_words = anno_block(
    gp = gpar(col = module_colours), labels = seq(length(unique(optimal_k))),
    labels_rot = 0, labels_gp = gpar(fontsize = 12)
  ),
  which = "row"
)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    row_split = optimal_k,
    left_annotation = row_anno,
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title = top_words,
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled ssGSEA Score", title_position = "lefttop-rot")
  ),
  column_title = "Cortical Glutamatergic"
)
```

### PCA

```{r}
seurat_sub <- RunPCA(seurat_sub, features = rownames(seurat_sub), reduction.name = "GSVA_pca", verbose = FALSE)
add_df <- data.frame(Embeddings(seurat_sub, reduction = "GSVA_pca"))
seurat_sub$GSVA_pca1 <- add_df[ , 1]
seurat_sub$GSVA_pca2 <- add_df[ , 2]

red_dim_plot(seurat_sub, x = "GSVA_pca1", y = "GSVA_pca2", color = "sample", type = "cat")

markers <- as.data.frame(Loadings(seurat_sub, reduction = "GSVA_pca"))
markers <- markers[ , 1:2]
q1 <- quantile(markers[ , 1])[4]
markers <- markers[markers[ , 1] > q1, ]
markers <- abs(markers)

markers$total <- rowSums(markers)
markers <- markers[order(markers$total, decreasing = TRUE), ]
markers$gene <- rownames(markers)
datatable_download(markers)
markers$cluster <- rep("0", times = nrow(markers))
markers$number <- seq(rownames(markers))
```

```{r}
FeaturePlot(
  seurat_sub, features = rownames(markers)[1:4], reduction = "GSVA_pca",
  order = TRUE, cols = c("lightgrey", "red")
)
```

```{r}
modules <- markers %>% group_by(cluster) %>% summarise(named_vec = list(number)) %>% tibble::deframe()
module_names <- modules
gene_sets_sub_ids <- gene_sets[names(gene_sets) %in% rownames(markers)]
gene_sets_sub_ids <- gene_sets_sub_ids[match(rownames(markers), names(gene_sets_sub_ids)), ]

for (j in seq(length(gene_sets_sub_ids@.Data))) {
  suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
}
go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(markers)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]])
}
top_words_tmp <- data.frame(
  gene_sets = unlist(module_names), module = rep(seq_along(module_names), lapply(module_names, length))
)
top_words_list <- vector("list", length = length(unique(top_words_tmp$module)))
top_words <- c()

for (i in seq_along(top_words_list)) {
  top_words_tmp_sub <- top_words_tmp[top_words_tmp$module == i, ]
  top_words_list[[i]] <- vector("list", nrow(top_words_tmp_sub))
  for (j in seq_along(top_words_list[[i]])) {
    top_words_list[[i]][[j]] <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_sets[j]))
  }
  tmp <- strsplit(paste(unlist(top_words_list[[i]]), collapse = " "), "\\s+")[[1]]
  remove <- which(tmp == "")
  if (length(remove) > 0) {
    tmp <- tmp[-which(tmp == "")]
  }
  tmp <- table(tmp)
  for (j in seq_along(tmp)) {
    total <- grep(paste0("\\b", names(tmp)[j], "\\b"), unlist(top_words_list[[i]]))
    tmp[j] <- sum(markers$total[total])
  }
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  dup <- which(duplicated(tmp))
  if (length(dup) > 0) {
    tmp <- tmp[-dup]
  }
  tmp <- names(tmp)
  if (length(tmp) >= 10) {
    tmp <- tmp[1:10]
    tmp <- paste(paste0(tmp, "\n"), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else if (length(tmp > 0)) {
    add <- 10 - length(tmp)
    tmp2 <- unlist(strsplit(suppressWarnings(word_cloud(top_words_tmp_sub$gene_set)), "\\s+"))
    tmp2 <- tmp2[tmp2 %notin% tmp]
    tmp <- c(tmp, tmp2[1:add])
    if (any(is.na(tmp))) {
      tmp <- tmp[-which(is.na(tmp))]
    }
    tmp <- paste(paste0(tmp, "\n"), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else {
    tmp <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_set, 10))
  }
  top_words <- c(top_words, tmp)
}

optimal_k <- unlist(lapply(seq_along(modules), function(i) {rep(names(modules)[i], length(modules[[i]]))}))
optimal_k <- factor(optimal_k, levels = unique(optimal_k))
module_colours <- cols25(length(unique(optimal_k)))

row_anno <-  HeatmapAnnotation(
  top_words = anno_block(
    gp = gpar(col = module_colours), labels = seq(length(unique(optimal_k))),
    labels_rot = 0, labels_gp = gpar(fontsize = 12)
  ),
  which = "row"
)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    row_split = optimal_k,
    left_annotation = row_anno,
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title = top_words,
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled ssGSEA Score", title_position = "lefttop-rot")
  ),
  column_title = "Cortical Glutamatergic PCA"
)
```

## Dentate Gyrus Glutamatergic

```{r}
seurat_sub <- seurat[ , which(seurat$annotations_short == "Dentate Gyrus Glutamatergic")]

sce <- as.SingleCellExperiment(seurat_sub, assay = "GSVA")
sce <- suppressWarnings(
  aggregateAcrossCells(sce, colData(sce)[ , "sample"], use_exprs_values = "counts", statistics = "sum")
)
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

rds <- file.path(cache_dir, "markers_gsva_dg.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat_sub, only.pos = TRUE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
markers <- markers %>% distinct(gene, .keep_all = TRUE)
rownames(markers) <- markers$gene
datatable_download_exp(markers)
markers$number <- seq(rownames(markers))

modules <- markers %>% group_by(cluster) %>% summarise(named_vec = list(number)) %>% tibble::deframe()
module_names <- modules
gene_sets_sub_ids <- gene_sets[names(gene_sets) %in% rownames(markers)]
gene_sets_sub_ids <- gene_sets_sub_ids[match(rownames(markers), names(gene_sets_sub_ids)), ]

for (j in seq(length(gene_sets_sub_ids@.Data))) {
  suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
}
go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(markers)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]])
}
top_words_tmp <- data.frame(
  gene_sets = unlist(module_names), module = rep(seq_along(module_names), lapply(module_names, length))
)
top_words_list <- vector("list", length = length(unique(top_words_tmp$module)))
top_words <- c()

for (i in seq_along(top_words_list)) {
  top_words_tmp_sub <- top_words_tmp[top_words_tmp$module == i, ]
  top_words_list[[i]] <- vector("list", nrow(top_words_tmp_sub))
  for (j in seq_along(top_words_list[[i]])) {
    top_words_list[[i]][[j]] <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_sets[j]))
  }
  tmp <- strsplit(paste(unlist(top_words_list[[i]]), collapse = " "), "\\s+")[[1]]
  remove <- which(tmp == "")
  if (length(remove) > 0) {
    tmp <- tmp[-which(tmp == "")]
  }
  tmp <- table(tmp)
  for (j in seq_along(tmp)) {
    log2FC <- grep(paste0("\\b", names(tmp)[j], "\\b"), unlist(top_words_list[[i]]))
    tmp[j] <- sum(markers$avg_log2FC[log2FC])
  }
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  dup <- which(duplicated(tmp))
  if (length(dup) > 0) {
    tmp <- tmp[-dup]
  }
  tmp <- names(tmp)
  if (length(tmp) >= 3) {
    tmp <- tmp[1:3]
    tmp <- paste(paste0(tmp, " "), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else if (length(tmp > 0)) {
    add <- 3 - length(tmp)
    tmp2 <- unlist(strsplit(suppressWarnings(word_cloud(top_words_tmp_sub$gene_set)), "\\s+"))
    tmp2 <- tmp2[tmp2 %notin% tmp]
    tmp <- c(tmp, tmp2[1:add])
    if (any(is.na(tmp))) {
      tmp <- tmp[-which(is.na(tmp))]
    }
    tmp <- paste(paste0(tmp, " "), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else {
    tmp <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_set, 3))
    tmp <- gsub("\n", " ", tmp)
  }
  top_words <- c(top_words, tmp)
}

optimal_k <- unlist(lapply(seq_along(modules), function(i) {rep(names(modules)[i], length(modules[[i]]))}))
optimal_k <- factor(optimal_k, levels = unique(optimal_k))
module_colours <- cols25(length(unique(optimal_k)))

row_anno <-  HeatmapAnnotation(
  top_words = anno_block(
    gp = gpar(col = module_colours), labels = seq(length(unique(optimal_k))),
    labels_rot = 0, labels_gp = gpar(fontsize = 12)
  ),
  which = "row"
)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    row_split = optimal_k,
    left_annotation = row_anno,
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title = top_words,
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled ssGSEA Score", title_position = "lefttop-rot")
  ),
  column_title = "Dentate Gyrus Glutamatergic"
)
```

### PCA

```{r}
seurat_sub <- RunPCA(seurat_sub, features = rownames(seurat_sub), reduction.name = "GSVA_pca", verbose = FALSE)
add_df <- data.frame(Embeddings(seurat_sub, reduction = "GSVA_pca"))
seurat_sub$GSVA_pca1 <- add_df[ , 1]
seurat_sub$GSVA_pca2 <- add_df[ , 2]

red_dim_plot(seurat_sub, x = "GSVA_pca1", y = "GSVA_pca2", color = "sample", type = "cat")
```

## Endothelial

```{r}
seurat_sub <- seurat[ , which(seurat$annotations_short == "Endothelial")]

sce <- as.SingleCellExperiment(seurat_sub, assay = "GSVA")
sce <- suppressWarnings(
  aggregateAcrossCells(sce, colData(sce)[ , "sample"], use_exprs_values = "counts", statistics = "sum")
)
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

rds <- file.path(cache_dir, "markers_gsva_endo.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat_sub, only.pos = TRUE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
markers <- markers %>% distinct(gene, .keep_all = TRUE)
rownames(markers) <- markers$gene
datatable_download_exp(markers)
markers$number <- seq(rownames(markers))

modules <- markers %>% group_by(cluster) %>% summarise(named_vec = list(number)) %>% tibble::deframe()
module_names <- modules
gene_sets_sub_ids <- gene_sets[names(gene_sets) %in% rownames(markers)]
gene_sets_sub_ids <- gene_sets_sub_ids[match(rownames(markers), names(gene_sets_sub_ids)), ]

for (j in seq(length(gene_sets_sub_ids@.Data))) {
  suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
}
go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(markers)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]])
}
top_words_tmp <- data.frame(
  gene_sets = unlist(module_names), module = rep(seq_along(module_names), lapply(module_names, length))
)
top_words_list <- vector("list", length = length(unique(top_words_tmp$module)))
top_words <- c()

for (i in seq_along(top_words_list)) {
  top_words_tmp_sub <- top_words_tmp[top_words_tmp$module == i, ]
  top_words_list[[i]] <- vector("list", nrow(top_words_tmp_sub))
  for (j in seq_along(top_words_list[[i]])) {
    top_words_list[[i]][[j]] <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_sets[j]))
  }
  tmp <- strsplit(paste(unlist(top_words_list[[i]]), collapse = " "), "\\s+")[[1]]
  remove <- which(tmp == "")
  if (length(remove) > 0) {
    tmp <- tmp[-which(tmp == "")]
  }
  tmp <- table(tmp)
  for (j in seq_along(tmp)) {
    log2FC <- grep(paste0("\\b", names(tmp)[j], "\\b"), unlist(top_words_list[[i]]))
    tmp[j] <- sum(markers$avg_log2FC[log2FC])
  }
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  dup <- which(duplicated(tmp))
  if (length(dup) > 0) {
    tmp <- tmp[-dup]
  }
  tmp <- names(tmp)
  if (length(tmp) >= 3) {
    tmp <- tmp[1:3]
    tmp <- paste(paste0(tmp, " "), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else if (length(tmp > 0)) {
    add <- 3 - length(tmp)
    tmp2 <- unlist(strsplit(suppressWarnings(word_cloud(top_words_tmp_sub$gene_set)), "\\s+"))
    tmp2 <- tmp2[tmp2 %notin% tmp]
    tmp <- c(tmp, tmp2[1:add])
    if (any(is.na(tmp))) {
      tmp <- tmp[-which(is.na(tmp))]
    }
    tmp <- paste(paste0(tmp, " "), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else {
    tmp <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_set, 3))
    tmp <- gsub("\n", " ", tmp)
  }
  top_words <- c(top_words, tmp)
}

optimal_k <- unlist(lapply(seq_along(modules), function(i) {rep(names(modules)[i], length(modules[[i]]))}))
optimal_k <- factor(optimal_k, levels = unique(optimal_k))
module_colours <- cols25(length(unique(optimal_k)))

row_anno <-  HeatmapAnnotation(
  top_words = anno_block(
    gp = gpar(col = module_colours), labels = seq(length(unique(optimal_k))),
    labels_rot = 0, labels_gp = gpar(fontsize = 12)
  ),
  which = "row"
)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    row_split = optimal_k,
    left_annotation = row_anno,
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title = top_words,
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled ssGSEA Score", title_position = "lefttop-rot")
  ),
  column_title = "Endothelial"
)
```

### PCA

```{r}
seurat_sub <- RunPCA(seurat_sub, features = rownames(seurat_sub), reduction.name = "GSVA_pca", verbose = FALSE)
add_df <- data.frame(Embeddings(seurat_sub, reduction = "GSVA_pca"))
seurat_sub$GSVA_pca1 <- add_df[ , 1]
seurat_sub$GSVA_pca2 <- add_df[ , 2]

red_dim_plot(seurat_sub, x = "GSVA_pca1", y = "GSVA_pca2", color = "sample", type = "cat")
```

## GABAergic

```{r}
seurat_sub <- seurat[ , which(seurat$annotations_short == "GABAergic")]

sce <- as.SingleCellExperiment(seurat_sub, assay = "GSVA")
sce <- suppressWarnings(
  aggregateAcrossCells(sce, colData(sce)[ , "sample"], use_exprs_values = "counts", statistics = "sum")
)
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

rds <- file.path(cache_dir, "markers_gsva_gaba.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat_sub, only.pos = TRUE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
markers <- markers %>% distinct(gene, .keep_all = TRUE)
rownames(markers) <- markers$gene
datatable_download_exp(markers)
markers$number <- seq(rownames(markers))

modules <- markers %>% group_by(cluster) %>% summarise(named_vec = list(number)) %>% tibble::deframe()
module_names <- modules
gene_sets_sub_ids <- gene_sets[names(gene_sets) %in% rownames(markers)]
gene_sets_sub_ids <- gene_sets_sub_ids[match(rownames(markers), names(gene_sets_sub_ids)), ]

for (j in seq(length(gene_sets_sub_ids@.Data))) {
  suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
}
go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(markers)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]])
}
top_words_tmp <- data.frame(
  gene_sets = unlist(module_names), module = rep(seq_along(module_names), lapply(module_names, length))
)
top_words_list <- vector("list", length = length(unique(top_words_tmp$module)))
top_words <- c()

for (i in seq_along(top_words_list)) {
  top_words_tmp_sub <- top_words_tmp[top_words_tmp$module == i, ]
  top_words_list[[i]] <- vector("list", nrow(top_words_tmp_sub))
  for (j in seq_along(top_words_list[[i]])) {
    top_words_list[[i]][[j]] <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_sets[j]))
  }
  tmp <- strsplit(paste(unlist(top_words_list[[i]]), collapse = " "), "\\s+")[[1]]
  remove <- which(tmp == "")
  if (length(remove) > 0) {
    tmp <- tmp[-which(tmp == "")]
  }
  tmp <- table(tmp)
  for (j in seq_along(tmp)) {
    log2FC <- grep(paste0("\\b", names(tmp)[j], "\\b"), unlist(top_words_list[[i]]))
    tmp[j] <- sum(markers$avg_log2FC[log2FC])
  }
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  dup <- which(duplicated(tmp))
  if (length(dup) > 0) {
    tmp <- tmp[-dup]
  }
  tmp <- names(tmp)
  if (length(tmp) >= 3) {
    tmp <- tmp[1:3]
    tmp <- paste(paste0(tmp, " "), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else if (length(tmp > 0)) {
    add <- 3 - length(tmp)
    tmp2 <- unlist(strsplit(suppressWarnings(word_cloud(top_words_tmp_sub$gene_set)), "\\s+"))
    tmp2 <- tmp2[tmp2 %notin% tmp]
    tmp <- c(tmp, tmp2[1:add])
    if (any(is.na(tmp))) {
      tmp <- tmp[-which(is.na(tmp))]
    }
    tmp <- paste(paste0(tmp, " "), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else {
    tmp <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_set, 3))
    tmp <- gsub("\n", " ", tmp)
  }
  top_words <- c(top_words, tmp)
}

optimal_k <- unlist(lapply(seq_along(modules), function(i) {rep(names(modules)[i], length(modules[[i]]))}))
optimal_k <- factor(optimal_k, levels = unique(optimal_k))
module_colours <- cols25(length(unique(optimal_k)))

row_anno <-  HeatmapAnnotation(
  top_words = anno_block(
    gp = gpar(col = module_colours), labels = seq(length(unique(optimal_k))),
    labels_rot = 0, labels_gp = gpar(fontsize = 12)
  ),
  which = "row"
)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    row_split = optimal_k,
    left_annotation = row_anno,
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title = top_words,
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled ssGSEA Score", title_position = "lefttop-rot")
  ),
  column_title = "GABAergic"
)
```

### PCA

```{r}
seurat_sub <- RunPCA(seurat_sub, features = rownames(seurat_sub), reduction.name = "GSVA_pca", verbose = FALSE)
add_df <- data.frame(Embeddings(seurat_sub, reduction = "GSVA_pca"))
seurat_sub$GSVA_pca1 <- add_df[ , 1]
seurat_sub$GSVA_pca2 <- add_df[ , 2]

red_dim_plot(seurat_sub, x = "GSVA_pca1", y = "GSVA_pca2", color = "sample", type = "cat")

markers <- as.data.frame(Loadings(seurat, reduction = "GSVA_pca"))
markers <- markers[ , 1:2]
q1 <- quantile(markers[ , 1])[4]
q2 <- quantile(markers[ , 2])[2]
markers <- markers[markers[ , 1] > q1, ]
markers <- markers[markers[ , 2] < q2, ]
markers <- abs(markers)

markers$total <- rowSums(markers)
markers <- markers[order(markers$total, decreasing = TRUE), ]
markers$gene <- rownames(markers)
datatable_download(markers)
markers$cluster <- rep("0", times = nrow(markers))
markers$number <- seq(rownames(markers))
```

```{r, fig.width = 10}
FeaturePlot(
  seurat_sub, features = rownames(markers)[1:4], reduction = "GSVA_pca",
  order = TRUE, cols = c("lightgrey", "red")
)
```

```{r}
modules <- markers %>% group_by(cluster) %>% summarise(named_vec = list(number)) %>% tibble::deframe()
module_names <- modules
gene_sets_sub_ids <- gene_sets[names(gene_sets) %in% rownames(markers)]
gene_sets_sub_ids <- gene_sets_sub_ids[match(rownames(markers), names(gene_sets_sub_ids)), ]

for (j in seq(length(gene_sets_sub_ids@.Data))) {
  suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
}
go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(markers)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]])
}
top_words_tmp <- data.frame(
  gene_sets = unlist(module_names), module = rep(seq_along(module_names), lapply(module_names, length))
)
top_words_list <- vector("list", length = length(unique(top_words_tmp$module)))
top_words <- c()

for (i in seq_along(top_words_list)) {
  top_words_tmp_sub <- top_words_tmp[top_words_tmp$module == i, ]
  top_words_list[[i]] <- vector("list", nrow(top_words_tmp_sub))
  for (j in seq_along(top_words_list[[i]])) {
    top_words_list[[i]][[j]] <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_sets[j]))
  }
  tmp <- strsplit(paste(unlist(top_words_list[[i]]), collapse = " "), "\\s+")[[1]]
  remove <- which(tmp == "")
  if (length(remove) > 0) {
    tmp <- tmp[-which(tmp == "")]
  }
  tmp <- table(tmp)
  for (j in seq_along(tmp)) {
    total <- grep(paste0("\\b", names(tmp)[j], "\\b"), unlist(top_words_list[[i]]))
    tmp[j] <- sum(markers$total[total])
  }
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  dup <- which(duplicated(tmp))
  if (length(dup) > 0) {
    tmp <- tmp[-dup]
  }
  tmp <- names(tmp)
  if (length(tmp) >= 10) {
    tmp <- tmp[1:10]
    tmp <- paste(paste0(tmp, "\n"), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else if (length(tmp > 0)) {
    add <- 10 - length(tmp)
    tmp2 <- unlist(strsplit(suppressWarnings(word_cloud(top_words_tmp_sub$gene_set)), "\\s+"))
    tmp2 <- tmp2[tmp2 %notin% tmp]
    tmp <- c(tmp, tmp2[1:add])
    if (any(is.na(tmp))) {
      tmp <- tmp[-which(is.na(tmp))]
    }
    tmp <- paste(paste0(tmp, "\n"), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else {
    tmp <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_set, 10))
  }
  top_words <- c(top_words, tmp)
}

optimal_k <- unlist(lapply(seq_along(modules), function(i) {rep(names(modules)[i], length(modules[[i]]))}))
optimal_k <- factor(optimal_k, levels = unique(optimal_k))
module_colours <- cols25(length(unique(optimal_k)))

row_anno <-  HeatmapAnnotation(
  top_words = anno_block(
    gp = gpar(col = module_colours), labels = seq(length(unique(optimal_k))),
    labels_rot = 0, labels_gp = gpar(fontsize = 12)
  ),
  which = "row"
)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    row_split = optimal_k,
    left_annotation = row_anno,
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title = top_words,
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled ssGSEA Score", title_position = "lefttop-rot")
  ),
  column_title = "GABAergic PCA"
)
```

## Hippocampal Glutamatergic

```{r}
seurat_sub <- seurat[ , which(seurat$annotations_short == "Hippocampal Glutamatergic")]

sce <- as.SingleCellExperiment(seurat_sub, assay = "GSVA")
sce <- suppressWarnings(
  aggregateAcrossCells(sce, colData(sce)[ , "sample"], use_exprs_values = "counts", statistics = "sum")
)
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

rds <- file.path(cache_dir, "markers_gsva_hip.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat_sub, only.pos = TRUE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
markers <- markers %>% distinct(gene, .keep_all = TRUE)
rownames(markers) <- markers$gene
datatable_download_exp(markers)
markers$number <- seq(rownames(markers))

modules <- markers %>% group_by(cluster) %>% summarise(named_vec = list(number)) %>% tibble::deframe()
module_names <- modules
gene_sets_sub_ids <- gene_sets[names(gene_sets) %in% rownames(markers)]
gene_sets_sub_ids <- gene_sets_sub_ids[match(rownames(markers), names(gene_sets_sub_ids)), ]

for (j in seq(length(gene_sets_sub_ids@.Data))) {
  suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
}
go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(markers)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]])
}
top_words_tmp <- data.frame(
  gene_sets = unlist(module_names), module = rep(seq_along(module_names), lapply(module_names, length))
)
top_words_list <- vector("list", length = length(unique(top_words_tmp$module)))
top_words <- c()

for (i in seq_along(top_words_list)) {
  top_words_tmp_sub <- top_words_tmp[top_words_tmp$module == i, ]
  top_words_list[[i]] <- vector("list", nrow(top_words_tmp_sub))
  for (j in seq_along(top_words_list[[i]])) {
    top_words_list[[i]][[j]] <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_sets[j]))
  }
  tmp <- strsplit(paste(unlist(top_words_list[[i]]), collapse = " "), "\\s+")[[1]]
  remove <- which(tmp == "")
  if (length(remove) > 0) {
    tmp <- tmp[-which(tmp == "")]
  }
  tmp <- table(tmp)
  for (j in seq_along(tmp)) {
    log2FC <- grep(paste0("\\b", names(tmp)[j], "\\b"), unlist(top_words_list[[i]]))
    tmp[j] <- sum(markers$avg_log2FC[log2FC])
  }
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  dup <- which(duplicated(tmp))
  if (length(dup) > 0) {
    tmp <- tmp[-dup]
  }
  tmp <- names(tmp)
  if (length(tmp) >= 3) {
    tmp <- tmp[1:3]
    tmp <- paste(paste0(tmp, " "), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else if (length(tmp > 0)) {
    add <- 3 - length(tmp)
    tmp2 <- unlist(strsplit(suppressWarnings(word_cloud(top_words_tmp_sub$gene_set)), "\\s+"))
    tmp2 <- tmp2[tmp2 %notin% tmp]
    tmp <- c(tmp, tmp2[1:add])
    if (any(is.na(tmp))) {
      tmp <- tmp[-which(is.na(tmp))]
    }
    tmp <- paste(paste0(tmp, " "), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else {
    tmp <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_set, 3))
    tmp <- gsub("\n", " ", tmp)
  }
  top_words <- c(top_words, tmp)
}

optimal_k <- unlist(lapply(seq_along(modules), function(i) {rep(names(modules)[i], length(modules[[i]]))}))
optimal_k <- factor(optimal_k, levels = unique(optimal_k))
module_colours <- cols25(length(unique(optimal_k)))

row_anno <-  HeatmapAnnotation(
  top_words = anno_block(
    gp = gpar(col = module_colours), labels = seq(length(unique(optimal_k))),
    labels_rot = 0, labels_gp = gpar(fontsize = 12)
  ),
  which = "row"
)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    row_split = optimal_k,
    left_annotation = row_anno,
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title = top_words,
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled ssGSEA Score", title_position = "lefttop-rot")
  ),
  column_title = "Hippocampal Glutamatergic"
)
```

### PCA

```{r}
seurat_sub <- RunPCA(seurat_sub, features = rownames(seurat_sub), reduction.name = "GSVA_pca", verbose = FALSE)
add_df <- data.frame(Embeddings(seurat_sub, reduction = "GSVA_pca"))
seurat_sub$GSVA_pca1 <- add_df[ , 1]
seurat_sub$GSVA_pca2 <- add_df[ , 2]

red_dim_plot(seurat_sub, x = "GSVA_pca1", y = "GSVA_pca2", color = "sample", type = "cat")

markers <- as.data.frame(Loadings(seurat, reduction = "GSVA_pca"))
markers <- markers[ , 1:2]
q1 <- quantile(markers[ , 1])[4]
q2 <- quantile(markers[ , 2])[2]
markers <- markers[markers[ , 1] > q1, ]
markers <- markers[markers[ , 2] < q2, ]
markers <- abs(markers)

markers$total <- rowSums(markers)
markers <- markers[order(markers$total, decreasing = TRUE), ]
markers$gene <- rownames(markers)
datatable_download(markers)
markers$cluster <- rep("0", times = nrow(markers))
markers$number <- seq(rownames(markers))
```

```{r, fig.width = 12.5}
FeaturePlot(
  seurat_sub, features = rownames(markers)[1:4], reduction = "GSVA_pca",
  order = TRUE, cols = c("lightgrey", "red")
)
```

```{r}
modules <- markers %>% group_by(cluster) %>% summarise(named_vec = list(number)) %>% tibble::deframe()
module_names <- modules
gene_sets_sub_ids <- gene_sets[names(gene_sets) %in% rownames(markers)]
gene_sets_sub_ids <- gene_sets_sub_ids[match(rownames(markers), names(gene_sets_sub_ids)), ]

for (j in seq(length(gene_sets_sub_ids@.Data))) {
  suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
}
go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(markers)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]])
}
top_words_tmp <- data.frame(
  gene_sets = unlist(module_names), module = rep(seq_along(module_names), lapply(module_names, length))
)
top_words_list <- vector("list", length = length(unique(top_words_tmp$module)))
top_words <- c()

for (i in seq_along(top_words_list)) {
  top_words_tmp_sub <- top_words_tmp[top_words_tmp$module == i, ]
  top_words_list[[i]] <- vector("list", nrow(top_words_tmp_sub))
  for (j in seq_along(top_words_list[[i]])) {
    top_words_list[[i]][[j]] <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_sets[j]))
  }
  tmp <- strsplit(paste(unlist(top_words_list[[i]]), collapse = " "), "\\s+")[[1]]
  remove <- which(tmp == "")
  if (length(remove) > 0) {
    tmp <- tmp[-which(tmp == "")]
  }
  tmp <- table(tmp)
  for (j in seq_along(tmp)) {
    total <- grep(paste0("\\b", names(tmp)[j], "\\b"), unlist(top_words_list[[i]]))
    tmp[j] <- sum(markers$total[total])
  }
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  dup <- which(duplicated(tmp))
  if (length(dup) > 0) {
    tmp <- tmp[-dup]
  }
  tmp <- names(tmp)
  if (length(tmp) >= 10) {
    tmp <- tmp[1:10]
    tmp <- paste(paste0(tmp, "\n"), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else if (length(tmp > 0)) {
    add <- 10 - length(tmp)
    tmp2 <- unlist(strsplit(suppressWarnings(word_cloud(top_words_tmp_sub$gene_set)), "\\s+"))
    tmp2 <- tmp2[tmp2 %notin% tmp]
    tmp <- c(tmp, tmp2[1:add])
    if (any(is.na(tmp))) {
      tmp <- tmp[-which(is.na(tmp))]
    }
    tmp <- paste(paste0(tmp, "\n"), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else {
    tmp <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_set, 10))
  }
  top_words <- c(top_words, tmp)
}

optimal_k <- unlist(lapply(seq_along(modules), function(i) {rep(names(modules)[i], length(modules[[i]]))}))
optimal_k <- factor(optimal_k, levels = unique(optimal_k))
module_colours <- cols25(length(unique(optimal_k)))

row_anno <-  HeatmapAnnotation(
  top_words = anno_block(
    gp = gpar(col = module_colours), labels = seq(length(unique(optimal_k))),
    labels_rot = 0, labels_gp = gpar(fontsize = 12)
  ),
  which = "row"
)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    row_split = optimal_k,
    left_annotation = row_anno,
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title = top_words,
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled ssGSEA Score", title_position = "lefttop-rot")
  ),
  column_title = "Hippocampal Glutamatergic PCA"
)
```

## Microglia

```{r}
seurat_sub <- seurat[ , which(seurat$annotations_short == "Microglia")]

sce <- as.SingleCellExperiment(seurat_sub, assay = "GSVA")
sce <- suppressWarnings(
  aggregateAcrossCells(sce, colData(sce)[ , "sample"], use_exprs_values = "counts", statistics = "sum")
)
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

rds <- file.path(cache_dir, "markers_gsva_micro.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat_sub, only.pos = TRUE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
markers <- markers %>% distinct(gene, .keep_all = TRUE)
rownames(markers) <- markers$gene
datatable_download_exp(markers)
markers$number <- seq(rownames(markers))

modules <- markers %>% group_by(cluster) %>% summarise(named_vec = list(number)) %>% tibble::deframe()
module_names <- modules
gene_sets_sub_ids <- gene_sets[names(gene_sets) %in% rownames(markers)]
gene_sets_sub_ids <- gene_sets_sub_ids[match(rownames(markers), names(gene_sets_sub_ids)), ]

for (j in seq(length(gene_sets_sub_ids@.Data))) {
  suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
}
go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(markers)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]])
}
top_words_tmp <- data.frame(
  gene_sets = unlist(module_names), module = rep(seq_along(module_names), lapply(module_names, length))
)
top_words_list <- vector("list", length = length(unique(top_words_tmp$module)))
top_words <- c()

for (i in seq_along(top_words_list)) {
  top_words_tmp_sub <- top_words_tmp[top_words_tmp$module == i, ]
  top_words_list[[i]] <- vector("list", nrow(top_words_tmp_sub))
  for (j in seq_along(top_words_list[[i]])) {
    top_words_list[[i]][[j]] <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_sets[j]))
  }
  tmp <- strsplit(paste(unlist(top_words_list[[i]]), collapse = " "), "\\s+")[[1]]
  remove <- which(tmp == "")
  if (length(remove) > 0) {
    tmp <- tmp[-which(tmp == "")]
  }
  tmp <- table(tmp)
  for (j in seq_along(tmp)) {
    log2FC <- grep(paste0("\\b", names(tmp)[j], "\\b"), unlist(top_words_list[[i]]))
    tmp[j] <- sum(markers$avg_log2FC[log2FC])
  }
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  dup <- which(duplicated(tmp))
  if (length(dup) > 0) {
    tmp <- tmp[-dup]
  }
  tmp <- names(tmp)
  if (length(tmp) >= 3) {
    tmp <- tmp[1:3]
    tmp <- paste(paste0(tmp, " "), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else if (length(tmp > 0)) {
    add <- 3 - length(tmp)
    tmp2 <- unlist(strsplit(suppressWarnings(word_cloud(top_words_tmp_sub$gene_set)), "\\s+"))
    tmp2 <- tmp2[tmp2 %notin% tmp]
    tmp <- c(tmp, tmp2[1:add])
    if (any(is.na(tmp))) {
      tmp <- tmp[-which(is.na(tmp))]
    }
    tmp <- paste(paste0(tmp, " "), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else {
    tmp <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_set, 3))
    tmp <- gsub("\n", " ", tmp)
  }
  top_words <- c(top_words, tmp)
}

optimal_k <- unlist(lapply(seq_along(modules), function(i) {rep(names(modules)[i], length(modules[[i]]))}))
optimal_k <- factor(optimal_k, levels = unique(optimal_k))
module_colours <- cols25(length(unique(optimal_k)))

row_anno <-  HeatmapAnnotation(
  top_words = anno_block(
    gp = gpar(col = module_colours), labels = seq(length(unique(optimal_k))),
    labels_rot = 0, labels_gp = gpar(fontsize = 12)
  ),
  which = "row"
)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    row_split = optimal_k,
    left_annotation = row_anno,
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title = top_words,
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled ssGSEA Score", title_position = "lefttop-rot")
  ),
  column_title = "Microglia"
)
```

### PCA

```{r}
seurat_sub <- RunPCA(seurat_sub, features = rownames(seurat_sub), reduction.name = "GSVA_pca", verbose = FALSE)
add_df <- data.frame(Embeddings(seurat_sub, reduction = "GSVA_pca"))
seurat_sub$GSVA_pca1 <- add_df[ , 1]
seurat_sub$GSVA_pca2 <- add_df[ , 2]

red_dim_plot(seurat_sub, x = "GSVA_pca1", y = "GSVA_pca2", color = "sample", type = "cat")

markers <- as.data.frame(Loadings(seurat, reduction = "GSVA_pca"))
markers <- markers[ , 1:2]
q1 <- quantile(markers[ , 1])[4]
markers <- markers[markers[ , 1] > q1, ]
markers <- abs(markers)

markers$total <- rowSums(markers)
markers <- markers[order(markers$total, decreasing = TRUE), ]
markers$gene <- rownames(markers)
datatable_download(markers)
markers$cluster <- rep("0", times = nrow(markers))
markers$number <- seq(rownames(markers))
```

```{r}
FeaturePlot(
  seurat_sub, features = rownames(markers)[1:4], reduction = "GSVA_pca",
  order = TRUE, cols = c("lightgrey", "red")
)
```

```{r}
modules <- markers %>% group_by(cluster) %>% summarise(named_vec = list(number)) %>% tibble::deframe()
module_names <- modules
gene_sets_sub_ids <- gene_sets[names(gene_sets) %in% rownames(markers)]
gene_sets_sub_ids <- gene_sets_sub_ids[match(rownames(markers), names(gene_sets_sub_ids)), ]

for (j in seq(length(gene_sets_sub_ids@.Data))) {
  suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
}
go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(markers)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]])
}
top_words_tmp <- data.frame(
  gene_sets = unlist(module_names), module = rep(seq_along(module_names), lapply(module_names, length))
)
top_words_list <- vector("list", length = length(unique(top_words_tmp$module)))
top_words <- c()

for (i in seq_along(top_words_list)) {
  top_words_tmp_sub <- top_words_tmp[top_words_tmp$module == i, ]
  top_words_list[[i]] <- vector("list", nrow(top_words_tmp_sub))
  for (j in seq_along(top_words_list[[i]])) {
    top_words_list[[i]][[j]] <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_sets[j]))
  }
  tmp <- strsplit(paste(unlist(top_words_list[[i]]), collapse = " "), "\\s+")[[1]]
  remove <- which(tmp == "")
  if (length(remove) > 0) {
    tmp <- tmp[-which(tmp == "")]
  }
  tmp <- table(tmp)
  for (j in seq_along(tmp)) {
    total <- grep(paste0("\\b", names(tmp)[j], "\\b"), unlist(top_words_list[[i]]))
    tmp[j] <- sum(markers$total[total])
  }
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  dup <- which(duplicated(tmp))
  if (length(dup) > 0) {
    tmp <- tmp[-dup]
  }
  tmp <- names(tmp)
  if (length(tmp) >= 10) {
    tmp <- tmp[1:10]
    tmp <- paste(paste0(tmp, "\n"), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else if (length(tmp > 0)) {
    add <- 10 - length(tmp)
    tmp2 <- unlist(strsplit(suppressWarnings(word_cloud(top_words_tmp_sub$gene_set)), "\\s+"))
    tmp2 <- tmp2[tmp2 %notin% tmp]
    tmp <- c(tmp, tmp2[1:add])
    if (any(is.na(tmp))) {
      tmp <- tmp[-which(is.na(tmp))]
    }
    tmp <- paste(paste0(tmp, "\n"), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else {
    tmp <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_set, 10))
  }
  top_words <- c(top_words, tmp)
}

optimal_k <- unlist(lapply(seq_along(modules), function(i) {rep(names(modules)[i], length(modules[[i]]))}))
optimal_k <- factor(optimal_k, levels = unique(optimal_k))
module_colours <- cols25(length(unique(optimal_k)))

row_anno <-  HeatmapAnnotation(
  top_words = anno_block(
    gp = gpar(col = module_colours), labels = seq(length(unique(optimal_k))),
    labels_rot = 0, labels_gp = gpar(fontsize = 12)
  ),
  which = "row"
)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    row_split = optimal_k,
    left_annotation = row_anno,
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title = top_words,
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled ssGSEA Score", title_position = "lefttop-rot")
  ),
  column_title = "Microglia PCA"
)
```

## Oligodendrocyte

```{r}
seurat_sub <- seurat[ , which(seurat$annotations_short == "Oligodendrocyte")]

sce <- as.SingleCellExperiment(seurat_sub, assay = "GSVA")
sce <- suppressWarnings(
  aggregateAcrossCells(sce, colData(sce)[ , "sample"], use_exprs_values = "counts", statistics = "sum")
)
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

rds <- file.path(cache_dir, "markers_gsva_oligo.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat_sub, only.pos = TRUE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
markers <- markers %>% distinct(gene, .keep_all = TRUE)
rownames(markers) <- markers$gene
datatable_download_exp(markers)
markers$number <- seq(rownames(markers))

modules <- markers %>% group_by(cluster) %>% summarise(named_vec = list(number)) %>% tibble::deframe()
module_names <- modules
gene_sets_sub_ids <- gene_sets[names(gene_sets) %in% rownames(markers)]
gene_sets_sub_ids <- gene_sets_sub_ids[match(rownames(markers), names(gene_sets_sub_ids)), ]

for (j in seq(length(gene_sets_sub_ids@.Data))) {
  suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
}
go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(markers)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]])
}
top_words_tmp <- data.frame(
  gene_sets = unlist(module_names), module = rep(seq_along(module_names), lapply(module_names, length))
)
top_words_list <- vector("list", length = length(unique(top_words_tmp$module)))
top_words <- c()

for (i in seq_along(top_words_list)) {
  top_words_tmp_sub <- top_words_tmp[top_words_tmp$module == i, ]
  top_words_list[[i]] <- vector("list", nrow(top_words_tmp_sub))
  for (j in seq_along(top_words_list[[i]])) {
    top_words_list[[i]][[j]] <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_sets[j]))
  }
  tmp <- strsplit(paste(unlist(top_words_list[[i]]), collapse = " "), "\\s+")[[1]]
  remove <- which(tmp == "")
  if (length(remove) > 0) {
    tmp <- tmp[-which(tmp == "")]
  }
  tmp <- table(tmp)
  for (j in seq_along(tmp)) {
    log2FC <- grep(paste0("\\b", names(tmp)[j], "\\b"), unlist(top_words_list[[i]]))
    tmp[j] <- sum(markers$avg_log2FC[log2FC])
  }
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  dup <- which(duplicated(tmp))
  if (length(dup) > 0) {
    tmp <- tmp[-dup]
  }
  tmp <- names(tmp)
  if (length(tmp) >= 3) {
    tmp <- tmp[1:3]
    tmp <- paste(paste0(tmp, " "), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else if (length(tmp > 0)) {
    add <- 3 - length(tmp)
    tmp2 <- unlist(strsplit(suppressWarnings(word_cloud(top_words_tmp_sub$gene_set)), "\\s+"))
    tmp2 <- tmp2[tmp2 %notin% tmp]
    tmp <- c(tmp, tmp2[1:add])
    if (any(is.na(tmp))) {
      tmp <- tmp[-which(is.na(tmp))]
    }
    tmp <- paste(paste0(tmp, " "), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else {
    tmp <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_set, 3))
    tmp <- gsub("\n", " ", tmp)
  }
  top_words <- c(top_words, tmp)
}

optimal_k <- unlist(lapply(seq_along(modules), function(i) {rep(names(modules)[i], length(modules[[i]]))}))
optimal_k <- factor(optimal_k, levels = unique(optimal_k))
module_colours <- cols25(length(unique(optimal_k)))

row_anno <-  HeatmapAnnotation(
  top_words = anno_block(
    gp = gpar(col = module_colours), labels = seq(length(unique(optimal_k))),
    labels_rot = 0, labels_gp = gpar(fontsize = 12)
  ),
  which = "row"
)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    row_split = optimal_k,
    left_annotation = row_anno,
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title = top_words,
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled ssGSEA Score", title_position = "lefttop-rot")
  ),
  column_title = "Oligodendrocyte"
)
```

### PCA

```{r}
seurat_sub <- RunPCA(seurat_sub, features = rownames(seurat_sub), reduction.name = "GSVA_pca", verbose = FALSE)
add_df <- data.frame(Embeddings(seurat_sub, reduction = "GSVA_pca"))
seurat_sub$GSVA_pca1 <- add_df[ , 1]
seurat_sub$GSVA_pca2 <- add_df[ , 2]

red_dim_plot(seurat_sub, x = "GSVA_pca1", y = "GSVA_pca2", color = "sample", type = "cat")

markers <- as.data.frame(Loadings(seurat, reduction = "GSVA_pca"))
markers <- markers[ , 1:2]
q1 <- quantile(markers[ , 1])[4]
q2 <- quantile(markers[ , 2])[2]
markers <- markers[markers[ , 1] > q1, ]
markers <- markers[markers[ , 2] < q2, ]
markers <- abs(markers)

markers$total <- rowSums(markers)
markers <- markers[order(markers$total, decreasing = TRUE), ]
markers$gene <- rownames(markers)
datatable_download(markers)
markers$cluster <- rep("0", times = nrow(markers))
markers$number <- seq(rownames(markers))
```

```{r, fig.width = 10}
FeaturePlot(
  seurat_sub, features = rownames(markers)[1:4], reduction = "GSVA_pca",
  order = TRUE, cols = c("lightgrey", "red")
)
```

```{r}
modules <- markers %>% group_by(cluster) %>% summarise(named_vec = list(number)) %>% tibble::deframe()
module_names <- modules
gene_sets_sub_ids <- gene_sets[names(gene_sets) %in% rownames(markers)]
gene_sets_sub_ids <- gene_sets_sub_ids[match(rownames(markers), names(gene_sets_sub_ids)), ]

for (j in seq(length(gene_sets_sub_ids@.Data))) {
  suppressWarnings(gene_sets_sub_ids[[j]]@setName <- gene_sets_sub_ids[[j]]@shortDescription)
}
go_anno <- select(GO.db, names(gene_sets_sub_ids), c("GOID", "TERM", "DEFINITION"), "GOID")
for (l in seq_along(module_names)) {
  module_names[[l]] <- paste0(rownames(markers)[module_names[[l]]], " ", go_anno$DEFINITION[module_names[[l]]])
}
top_words_tmp <- data.frame(
  gene_sets = unlist(module_names), module = rep(seq_along(module_names), lapply(module_names, length))
)
top_words_list <- vector("list", length = length(unique(top_words_tmp$module)))
top_words <- c()

for (i in seq_along(top_words_list)) {
  top_words_tmp_sub <- top_words_tmp[top_words_tmp$module == i, ]
  top_words_list[[i]] <- vector("list", nrow(top_words_tmp_sub))
  for (j in seq_along(top_words_list[[i]])) {
    top_words_list[[i]][[j]] <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_sets[j]))
  }
  tmp <- strsplit(paste(unlist(top_words_list[[i]]), collapse = " "), "\\s+")[[1]]
  remove <- which(tmp == "")
  if (length(remove) > 0) {
    tmp <- tmp[-which(tmp == "")]
  }
  tmp <- table(tmp)
  for (j in seq_along(tmp)) {
    total <- grep(paste0("\\b", names(tmp)[j], "\\b"), unlist(top_words_list[[i]]))
    tmp[j] <- sum(markers$total[total])
  }
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  dup <- which(duplicated(tmp))
  if (length(dup) > 0) {
    tmp <- tmp[-dup]
  }
  tmp <- names(tmp)
  if (length(tmp) >= 10) {
    tmp <- tmp[1:10]
    tmp <- paste(paste0(tmp, "\n"), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else if (length(tmp > 0)) {
    add <- 10 - length(tmp)
    tmp2 <- unlist(strsplit(suppressWarnings(word_cloud(top_words_tmp_sub$gene_set)), "\\s+"))
    tmp2 <- tmp2[tmp2 %notin% tmp]
    tmp <- c(tmp, tmp2[1:add])
    if (any(is.na(tmp))) {
      tmp <- tmp[-which(is.na(tmp))]
    }
    tmp <- paste(paste0(tmp, "\n"), collapse = "")
    tmp <- substr(tmp, 1, nchar(tmp) - 1)
  } else {
    tmp <- suppressWarnings(word_cloud(top_words_tmp_sub$gene_set, 10))
  }
  top_words <- c(top_words, tmp)
}

optimal_k <- unlist(lapply(seq_along(modules), function(i) {rep(names(modules)[i], length(modules[[i]]))}))
optimal_k <- factor(optimal_k, levels = unique(optimal_k))
module_colours <- cols25(length(unique(optimal_k)))

row_anno <-  HeatmapAnnotation(
  top_words = anno_block(
    gp = gpar(col = module_colours), labels = seq(length(unique(optimal_k))),
    labels_rot = 0, labels_gp = gpar(fontsize = 12)
  ),
  which = "row"
)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_row_names = FALSE,
    row_split = optimal_k,
    left_annotation = row_anno,
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title = top_words,
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled ssGSEA Score", title_position = "lefttop-rot")
  ),
  column_title = "Oligodendrocyte PCA"
)
```

# DGE

```{r}
DefaultAssay(seurat) <- "SCT"
```

## All Cell-types

```{r}
sce <- as.SingleCellExperiment(seurat, assay = "SCT")
sce <- suppressWarnings(
  aggregateAcrossCells(sce, colData(sce)[ , "sample"], use_exprs_values = "counts", statistics = "sum")
)
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

rds <- file.path(cache_dir, "markers_sct_all.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
markers <- markers %>% distinct(gene, .keep_all = TRUE)
rownames(markers) <- markers$gene
datatable_download_exp(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 3)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    row_split = rep(seq(6), each = 3),
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled Normalised Counts", title_position = "lefttop-rot")
  ),
  column_title = "All Cell-types"
)
```

## Astrocyte

```{r}
seurat_sub <- seurat[ , which(seurat$annotations_short == "Astrocyte")]

sce <- as.SingleCellExperiment(seurat_sub, assay = "SCT")
sce <- suppressWarnings(
  aggregateAcrossCells(sce, colData(sce)[ , "sample"], use_exprs_values = "counts", statistics = "sum")
)
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

rds <- file.path(cache_dir, "markers_sct_astro.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat_sub, only.pos = TRUE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
markers <- markers %>% distinct(gene, .keep_all = TRUE)
rownames(markers) <- markers$gene
datatable_download_exp(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 3)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    row_split = rep(seq(6), each = 3),
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled Normalised Counts", title_position = "lefttop-rot")
  ),
  column_title = "Astrocyte"
)
```

## Cortical Glutamatergic

```{r}
seurat_sub <- seurat[ , which(seurat$annotations_short == "Cortical Glutamatergic")]

sce <- as.SingleCellExperiment(seurat_sub, assay = "SCT")
sce <- suppressWarnings(
  aggregateAcrossCells(sce, colData(sce)[ , "sample"], use_exprs_values = "counts", statistics = "sum")
)
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

rds <- file.path(cache_dir, "markers_sct_glut.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat_sub, only.pos = TRUE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
markers <- markers %>% distinct(gene, .keep_all = TRUE)
rownames(markers) <- markers$gene
datatable_download_exp(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 3)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    row_split = rep(seq(6), each = 3),
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled Normalised Counts", title_position = "lefttop-rot")
  ),
  column_title = "Cortical Glutamatergic"
)
```

## Dentate Gyrus Glutamatergic

```{r}
seurat_sub <- seurat[ , which(seurat$annotations_short == "Dentate Gyrus Glutamatergic")]

sce <- as.SingleCellExperiment(seurat_sub, assay = "SCT")
sce <- suppressWarnings(
  aggregateAcrossCells(sce, colData(sce)[ , "sample"], use_exprs_values = "counts", statistics = "sum")
)
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

rds <- file.path(cache_dir, "markers_sct_dg.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat_sub, only.pos = TRUE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
markers <- markers %>% distinct(gene, .keep_all = TRUE)
rownames(markers) <- markers$gene
datatable_download_exp(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 3)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    row_split = rep(seq(3), each = 3),
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled Normalised Counts", title_position = "lefttop-rot")
  ),
  column_title = "Dentate Gyrus Glutamatergic"
)
```

## GABAergic

```{r}
seurat_sub <- seurat[ , which(seurat$annotations_short == "GABAergic")]

sce <- as.SingleCellExperiment(seurat_sub, assay = "SCT")
sce <- suppressWarnings(
  aggregateAcrossCells(sce, colData(sce)[ , "sample"], use_exprs_values = "counts", statistics = "sum")
)
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

rds <- file.path(cache_dir, "markers_sct_gaba.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat_sub, only.pos = TRUE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
markers <- markers %>% distinct(gene, .keep_all = TRUE)
rownames(markers) <- markers$gene
datatable_download_exp(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 3)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    row_split = rep(seq(6), each = 3),
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled Normalised Counts", title_position = "lefttop-rot")
  ),
  column_title = "GABAergic"
)
```

## Hippocampal Glutamatergic

```{r}
seurat_sub <- seurat[ , which(seurat$annotations_short == "Hippocampal Glutamatergic")]

sce <- as.SingleCellExperiment(seurat_sub, assay = "SCT")
sce <- suppressWarnings(
  aggregateAcrossCells(sce, colData(sce)[ , "sample"], use_exprs_values = "counts", statistics = "sum")
)
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

rds <- file.path(cache_dir, "markers_sct_hip.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat_sub, only.pos = TRUE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
markers <- markers %>% distinct(gene, .keep_all = TRUE)
rownames(markers) <- markers$gene
datatable_download_exp(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 3)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    row_split = rep(seq(6), each = 3),
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled Normalised Counts", title_position = "lefttop-rot")
  ),
  column_title = "Hippocampal Glutamatergic"
)
```

## Microglia

```{r}
seurat_sub <- seurat[ , which(seurat$annotations_short == "Microglia")]

sce <- as.SingleCellExperiment(seurat_sub, assay = "SCT")
sce <- suppressWarnings(
  aggregateAcrossCells(sce, colData(sce)[ , "sample"], use_exprs_values = "counts", statistics = "sum")
)
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

rds <- file.path(cache_dir, "markers_sct_micro.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat_sub, only.pos = TRUE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
markers <- markers %>% distinct(gene, .keep_all = TRUE)
rownames(markers) <- markers$gene
datatable_download_exp(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 3)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    row_split = rep(seq(6), each = 3),
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled Normalised Counts", title_position = "lefttop-rot")
  ),
  column_title = "Microglia"
)
```

## Oligodendrocyte

```{r}
seurat_sub <- seurat[ , which(seurat$annotations_short == "Oligodendrocyte")]

sce <- as.SingleCellExperiment(seurat_sub, assay = "SCT")
sce <- suppressWarnings(
  aggregateAcrossCells(sce, colData(sce)[ , "sample"], use_exprs_values = "counts", statistics = "sum")
)
dge <- DGEList(counts(sce))
dge <- calcNormFactors(dge)
logcounts(sce, withDimnames = FALSE) <- cpm(dge, log = TRUE, prior.count = 3)

rds <- file.path(cache_dir, "markers_sct_oligo.rds")
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat_sub, only.pos = TRUE)
  saveRDS(markers, rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
markers <- markers %>% distinct(gene, .keep_all = TRUE)
rownames(markers) <- markers$gene
datatable_download_exp(markers)
markers <- markers %>% group_by(cluster) %>% slice_head(n = 3)

mat <- logcounts(sce)
mat <- mat[rownames(mat) %in% markers$gene, ]
mat <- mat[match(markers$gene, rownames(mat)), ]

set.seed(1)
draw(
  Heatmap(
    t(
      apply(
        mat, 1,
        function (x) ((2 * (x - min(x)) / (max(x) - min(x))) - 1)
      )
    ),
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    row_split = rep(seq(6), each = 3),
    rect_gp = gpar(col = "lightgray", lwd = 0.1),
    row_gap = unit(2, "mm"),
    column_gap = unit(2, "mm"),
    column_names_side = "top",
    row_title_rot = 0,
    column_names_rot = 45,
    heatmap_legend_param = list(title = "Scaled Normalised Counts", title_position = "lefttop-rot")
  ),
  column_title = "Oligodendrocyte"
)
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
