---
title: "Comb - Batch 02 - 01 GeneFunnel"
author:
  - name: "Emir Turkes"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_file = file.path("..", "results", "comb_batch02_01genefunnel.html")
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 35px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [tau-mutant-omics-analysis](https://github.com/eturkes/tau-mutant-omics-analysis).*

The table of contents in the top left is clickable and can be used to quickly navigate the document.
To toggle the visibility of code, use the `CODE` toggles at the top right of chunks.
The toggle at the start of the document controls the visibility of all chunks.

```{r}
#    This file is part of tau-mutant-omics-analysis.
#    Copyright (C) 2023  Emir Turkes, Naoto Watamura, UK DRI at UCL
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# Load required packages, suppressing startup messages.
# -----------------------------------------------------
library(conflicted)
conflicts_prefer(DT::JS, .quiet = TRUE)
packages <- c("Seurat", "limma", "DT", "RColorBrewer", "ComplexHeatmap")
invisible(suppressPackageStartupMessages(lapply(packages, FUN = library, character.only = TRUE)))
# -----------------------------------------------------

# Define global settings.
# -----------------------
knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# -----------------------

# Define functions.
# -----------------
source("utils.R")
# -----------------

# Useful variables.
# -----------------
genotype <- "comb"
batch <- 2
step <- 1

cache_dir <- file.path("..", "cache", genotype, paste0("0", batch), paste0("0", step))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

colour <- colorRampPalette(rev(brewer.pal(9, "RdBu")))(100)
# -----------------

# Useful data.
# ------------
gene_anno <- readRDS(file.path("..", "cache", "MAPTKI", "02", "02", "gene_anno.rds"))
# ------------
```

# NLGF MAPT KI vs. MAPT KI

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "MAPTKI", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "MAPTKI", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Astrocyte

```{r}
ctype <- "Astrocyte"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Astrocyte")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Astrocyte")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("MAPTKI", ncol(gsva1_sub)), rep("NLGF_MAPTKI", ncol(gsva2_sub))), levels = c("MAPTKI", "NLGF_MAPTKI")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_MAPTKI-MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Astrocyte NLGF MAPT KI GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Astrocyte NLGF MAPT KI GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Astrocyte")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Astrocyte")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("MAPTKI", ncol(seurat1_sub)), rep("NLGF_MAPTKI", ncol(seurat2_sub))), levels = c("MAPTKI", "NLGF_MAPTKI")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_MAPTKI-MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Astrocyte NLGF MAPT KI DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Astrocyte NLGF MAPT KI DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF S305N+3 vs. S305N+3

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "S305N", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_S305N", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "S305N", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_S305N", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Astrocyte

```{r}
ctype <- "Astrocyte"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Astrocyte")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Astrocyte")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("S305N", ncol(gsva1_sub)), rep("NLGF_S305N", ncol(gsva2_sub))), levels = c("S305N", "NLGF_S305N")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_S305N-S305N, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Astrocyte NLGF S305N+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Astrocyte NLGF S305N+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Astrocyte")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Astrocyte")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("S305N", ncol(seurat1_sub)), rep("NLGF_S305N", ncol(seurat2_sub))), levels = c("S305N", "NLGF_S305N")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_S305N-S305N, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Astrocyte NLGF S305N+3 DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Astrocyte NLGF S305N+3 DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF P301S+3 vs. P301S+3

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "P301S", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "P301S", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Astrocyte

```{r}
ctype <- "Astrocyte"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Astrocyte")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Astrocyte")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("P301S", ncol(gsva1_sub)), rep("NLGF_P301S", ncol(gsva2_sub))), levels = c("P301S", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_P301S-P301S, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Astrocyte NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Astrocyte NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Astrocyte")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Astrocyte")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("P301S", ncol(seurat1_sub)), rep("NLGF_P301S", ncol(seurat2_sub))), levels = c("P301S", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_P301S-P301S, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Astrocyte NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Astrocyte NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF P301S+3 vs. NLGF MAPT KI

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Astrocyte

```{r}
ctype <- "Astrocyte"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Astrocyte")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Astrocyte")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("NLGF_MAPTKI", ncol(gsva1_sub)), rep("NLGF_P301S", ncol(gsva2_sub))), levels = c("NLGF_MAPTKI", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_P301S-NLGF_MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Astrocyte NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Astrocyte NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Astrocyte")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Astrocyte")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("NLGF_MAPTKI", ncol(seurat1_sub)), rep("NLGF_P301S", ncol(seurat2_sub))), levels = c("NLGF_MAPTKI", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_P301S-NLGF_MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Astrocyte NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Astrocyte NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF MAPT KI vs. MAPT KI

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "MAPTKI", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "MAPTKI", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Cortical Glutamatergic

```{r}
ctype <- "Cortical Glutamatergic"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Cortical Glutamatergic")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Cortical Glutamatergic")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("MAPTKI", ncol(gsva1_sub)), rep("NLGF_MAPTKI", ncol(gsva2_sub))), levels = c("MAPTKI", "NLGF_MAPTKI")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_MAPTKI-MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Cortical Glutamatergic NLGF MAPT KI GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Cortical Glutamatergic NLGF MAPT KI GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Cortical Glutamatergic")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Cortical Glutamatergic")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("MAPTKI", ncol(seurat1_sub)), rep("NLGF_MAPTKI", ncol(seurat2_sub))), levels = c("MAPTKI", "NLGF_MAPTKI")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_MAPTKI-MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Cortical Glutamatergic NLGF MAPT KI DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Cortical Glutamatergic NLGF MAPT KI DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF S305N+3 vs. S305N+3

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "S305N", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_S305N", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "S305N", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_S305N", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Cortical Glutamatergic

```{r}
ctype <- "Cortical Glutamatergic"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Cortical Glutamatergic")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Cortical Glutamatergic")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("S305N", ncol(gsva1_sub)), rep("NLGF_S305N", ncol(gsva2_sub))), levels = c("S305N", "NLGF_S305N")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_S305N-S305N, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Cortical Glutamatergic NLGF S305N+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Cortical Glutamatergic NLGF S305N+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Cortical Glutamatergic")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Cortical Glutamatergic")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("S305N", ncol(seurat1_sub)), rep("NLGF_S305N", ncol(seurat2_sub))), levels = c("S305N", "NLGF_S305N")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_S305N-S305N, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Cortical Glutamatergic NLGF S305N+3 DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Cortical Glutamatergic NLGF S305N+3 DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF P301S+3 vs. P301S+3

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "P301S", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "P301S", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Cortical Glutamatergic

```{r}
ctype <- "Cortical Glutamatergic"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Cortical Glutamatergic")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Cortical Glutamatergic")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("P301S", ncol(gsva1_sub)), rep("NLGF_P301S", ncol(gsva2_sub))), levels = c("P301S", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_P301S-P301S, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Cortical Glutamatergic NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Cortical Glutamatergic NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Cortical Glutamatergic")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Cortical Glutamatergic")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("P301S", ncol(seurat1_sub)), rep("NLGF_P301S", ncol(seurat2_sub))), levels = c("P301S", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_P301S-P301S, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Cortical Glutamatergic NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Cortical Glutamatergic NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF P301S+3 vs. NLGF MAPT KI

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Cortical Glutamatergic

```{r}
ctype <- "Cortical Glutamatergic"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Cortical Glutamatergic")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Cortical Glutamatergic")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("NLGF_MAPTKI", ncol(gsva1_sub)), rep("NLGF_P301S", ncol(gsva2_sub))), levels = c("NLGF_MAPTKI", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_P301S-NLGF_MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Cortical Glutamatergic NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Cortical Glutamatergic NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Cortical Glutamatergic")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Cortical Glutamatergic")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("NLGF_MAPTKI", ncol(seurat1_sub)), rep("NLGF_P301S", ncol(seurat2_sub))), levels = c("NLGF_MAPTKI", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_P301S-NLGF_MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Cortical Glutamatergic NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Cortical Glutamatergic NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF MAPT KI vs. MAPT KI

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "MAPTKI", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "MAPTKI", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## GABAergic

```{r}
ctype <- "GABAergic"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "GABAergic")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "GABAergic")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("MAPTKI", ncol(gsva1_sub)), rep("NLGF_MAPTKI", ncol(gsva2_sub))), levels = c("MAPTKI", "NLGF_MAPTKI")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_MAPTKI-MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in GABAergic NLGF MAPT KI GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in GABAergic NLGF MAPT KI GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "GABAergic")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "GABAergic")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("MAPTKI", ncol(seurat1_sub)), rep("NLGF_MAPTKI", ncol(seurat2_sub))), levels = c("MAPTKI", "NLGF_MAPTKI")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_MAPTKI-MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in GABAergic NLGF MAPT KI DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in GABAergic NLGF MAPT KI DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

# mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
# mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
# set.seed(1)
# draw(
#   Heatmap(
#     mat,
#     colour,
#     cluster_rows = FALSE,
#     cluster_columns = FALSE,
#     show_column_names = FALSE,
#     column_split = condition,
#     column_title = ctype,
#     heatmap_legend_param = list(
#       title = "Scaled Normalised Counts", direction = "horizontal",
#       legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
#       labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
#     )
#   ),
#   heatmap_legend_side = "top",
#   annotation_legend_side = "top"
# )
```

# NLGF S305N+3 vs. S305N+3

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "S305N", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_S305N", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "S305N", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_S305N", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## GABAergic

```{r}
ctype <- "GABAergic"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "GABAergic")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "GABAergic")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("S305N", ncol(gsva1_sub)), rep("NLGF_S305N", ncol(gsva2_sub))), levels = c("S305N", "NLGF_S305N")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_S305N-S305N, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in GABAergic NLGF S305N+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in GABAergic NLGF S305N+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "GABAergic")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "GABAergic")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("S305N", ncol(seurat1_sub)), rep("NLGF_S305N", ncol(seurat2_sub))), levels = c("S305N", "NLGF_S305N")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_S305N-S305N, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in GABAergic NLGF S305N+3 DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in GABAergic NLGF S305N+3 DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF P301S+3 vs. P301S+3

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "P301S", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "P301S", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## GABAergic

```{r}
ctype <- "GABAergic"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "GABAergic")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "GABAergic")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("P301S", ncol(gsva1_sub)), rep("NLGF_P301S", ncol(gsva2_sub))), levels = c("P301S", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_P301S-P301S, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in GABAergic NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in GABAergic NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "GABAergic")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "GABAergic")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("P301S", ncol(seurat1_sub)), rep("NLGF_P301S", ncol(seurat2_sub))), levels = c("P301S", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_P301S-P301S, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in GABAergic NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in GABAergic NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF P301S+3 vs. NLGF MAPT KI

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## GABAergic

```{r}
ctype <- "GABAergic"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "GABAergic")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "GABAergic")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("NLGF_MAPTKI", ncol(gsva1_sub)), rep("NLGF_P301S", ncol(gsva2_sub))), levels = c("NLGF_MAPTKI", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_P301S-NLGF_MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in GABAergic NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in GABAergic NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "GABAergic")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "GABAergic")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("NLGF_MAPTKI", ncol(seurat1_sub)), rep("NLGF_P301S", ncol(seurat2_sub))), levels = c("NLGF_MAPTKI", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_P301S-NLGF_MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in GABAergic NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in GABAergic NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF MAPT KI vs. MAPT KI

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "MAPTKI", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "MAPTKI", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Hippocampal Glutamatergic

```{r}
ctype <- "Hippocampal Glutamatergic"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Hippocampal Glutamatergic")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Hippocampal Glutamatergic")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("MAPTKI", ncol(gsva1_sub)), rep("NLGF_MAPTKI", ncol(gsva2_sub))), levels = c("MAPTKI", "NLGF_MAPTKI")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_MAPTKI-MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Hippocampal Glutamatergic NLGF MAPT KI GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Hippocampal Glutamatergic NLGF MAPT KI GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Hippocampal Glutamatergic")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Hippocampal Glutamatergic")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("MAPTKI", ncol(seurat1_sub)), rep("NLGF_MAPTKI", ncol(seurat2_sub))), levels = c("MAPTKI", "NLGF_MAPTKI")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_MAPTKI-MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Hippocampal Glutamatergic NLGF MAPT KI DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Hippocampal Glutamatergic NLGF MAPT KI DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF S305N+3 vs. S305N+3

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "S305N", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_S305N", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "S305N", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_S305N", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Hippocampal Glutamatergic

```{r}
ctype <- "Hippocampal Glutamatergic"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Hippocampal Glutamatergic")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Hippocampal Glutamatergic")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("S305N", ncol(gsva1_sub)), rep("NLGF_S305N", ncol(gsva2_sub))), levels = c("S305N", "NLGF_S305N")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_S305N-S305N, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Hippocampal Glutamatergic NLGF S305N+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Hippocampal Glutamatergic NLGF S305N+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Hippocampal Glutamatergic")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Hippocampal Glutamatergic")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("S305N", ncol(seurat1_sub)), rep("NLGF_S305N", ncol(seurat2_sub))), levels = c("S305N", "NLGF_S305N")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_S305N-S305N, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Hippocampal Glutamatergic NLGF S305N+3 DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Hippocampal Glutamatergic NLGF S305N+3 DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF P301S+3 vs. P301S+3

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "P301S", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "P301S", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Hippocampal Glutamatergic

```{r}
ctype <- "Hippocampal Glutamatergic"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Hippocampal Glutamatergic")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Hippocampal Glutamatergic")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("P301S", ncol(gsva1_sub)), rep("NLGF_P301S", ncol(gsva2_sub))), levels = c("P301S", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_P301S-P301S, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Hippocampal Glutamatergic NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Hippocampal Glutamatergic NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Hippocampal Glutamatergic")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Hippocampal Glutamatergic")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("P301S", ncol(seurat1_sub)), rep("NLGF_P301S", ncol(seurat2_sub))), levels = c("P301S", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_P301S-P301S, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Hippocampal Glutamatergic NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Hippocampal Glutamatergic NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF P301S+3 vs. NLGF MAPT KI

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Hippocampal Glutamatergic

```{r}
ctype <- "Hippocampal Glutamatergic"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Hippocampal Glutamatergic")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Hippocampal Glutamatergic")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("NLGF_MAPTKI", ncol(gsva1_sub)), rep("NLGF_P301S", ncol(gsva2_sub))), levels = c("NLGF_MAPTKI", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_P301S-NLGF_MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Hippocampal Glutamatergic NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Hippocampal Glutamatergic NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Hippocampal Glutamatergic")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Hippocampal Glutamatergic")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("NLGF_MAPTKI", ncol(seurat1_sub)), rep("NLGF_P301S", ncol(seurat2_sub))), levels = c("NLGF_MAPTKI", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_P301S-NLGF_MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Hippocampal Glutamatergic NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Hippocampal Glutamatergic NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF MAPT KI vs. MAPT KI

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "MAPTKI", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "MAPTKI", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Microglia

```{r}
ctype <- "Microglia"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Microglia")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Microglia")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("MAPTKI", ncol(gsva1_sub)), rep("NLGF_MAPTKI", ncol(gsva2_sub))), levels = c("MAPTKI", "NLGF_MAPTKI")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_MAPTKI-MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Microglia NLGF MAPT KI GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Microglia NLGF MAPT KI GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Microglia")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Microglia")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("MAPTKI", ncol(seurat1_sub)), rep("NLGF_MAPTKI", ncol(seurat2_sub))), levels = c("MAPTKI", "NLGF_MAPTKI")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_MAPTKI-MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Microglia NLGF MAPT KI DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Microglia NLGF MAPT KI DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF S305N+3 vs. S305N+3

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "S305N", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_S305N", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "S305N", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_S305N", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Microglia

```{r}
ctype <- "Microglia"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Microglia")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Microglia")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("S305N", ncol(gsva1_sub)), rep("NLGF_S305N", ncol(gsva2_sub))), levels = c("S305N", "NLGF_S305N")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_S305N-S305N, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Microglia NLGF S305N+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Microglia NLGF S305N+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Microglia")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Microglia")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("S305N", ncol(seurat1_sub)), rep("NLGF_S305N", ncol(seurat2_sub))), levels = c("S305N", "NLGF_S305N")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_S305N-S305N, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Microglia NLGF S305N+3 DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Microglia NLGF S305N+3 DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF P301S+3 vs. P301S+3

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "P301S", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "P301S", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Microglia

```{r}
ctype <- "Microglia"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Microglia")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Microglia")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("P301S", ncol(gsva1_sub)), rep("NLGF_P301S", ncol(gsva2_sub))), levels = c("P301S", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_P301S-P301S, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Microglia NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Microglia NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Microglia")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Microglia")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("P301S", ncol(seurat1_sub)), rep("NLGF_P301S", ncol(seurat2_sub))), levels = c("P301S", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_P301S-P301S, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Microglia NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Microglia NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF P301S+3 vs. NLGF MAPT KI

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Microglia

```{r}
ctype <- "Microglia"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Microglia")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Microglia")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("NLGF_MAPTKI", ncol(gsva1_sub)), rep("NLGF_P301S", ncol(gsva2_sub))), levels = c("NLGF_MAPTKI", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_P301S-NLGF_MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Microglia NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Microglia NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Microglia")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Microglia")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("NLGF_MAPTKI", ncol(seurat1_sub)), rep("NLGF_P301S", ncol(seurat2_sub))), levels = c("NLGF_MAPTKI", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_P301S-NLGF_MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Microglia NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Microglia NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF MAPT KI vs. MAPT KI

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "MAPTKI", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "MAPTKI", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Oligodendrocyte

```{r}
ctype <- "Oligodendrocyte"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Oligodendrocyte")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Oligodendrocyte")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("MAPTKI", ncol(gsva1_sub)), rep("NLGF_MAPTKI", ncol(gsva2_sub))), levels = c("MAPTKI", "NLGF_MAPTKI")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_MAPTKI-MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Oligodendrocyte NLGF MAPT KI GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Oligodendrocyte NLGF MAPT KI GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Oligodendrocyte")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Oligodendrocyte")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("MAPTKI", ncol(seurat1_sub)), rep("NLGF_MAPTKI", ncol(seurat2_sub))), levels = c("MAPTKI", "NLGF_MAPTKI")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_MAPTKI-MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Oligodendrocyte NLGF MAPT KI DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Oligodendrocyte NLGF MAPT KI DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF S305N+3 vs. S305N+3

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "S305N", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_S305N", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "S305N", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_S305N", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Oligodendrocyte

```{r}
ctype <- "Oligodendrocyte"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Oligodendrocyte")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Oligodendrocyte")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("S305N", ncol(gsva1_sub)), rep("NLGF_S305N", ncol(gsva2_sub))), levels = c("S305N", "NLGF_S305N")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_S305N-S305N, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Oligodendrocyte NLGF S305N+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Oligodendrocyte NLGF S305N+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Oligodendrocyte")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Oligodendrocyte")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("S305N", ncol(seurat1_sub)), rep("NLGF_S305N", ncol(seurat2_sub))), levels = c("S305N", "NLGF_S305N")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_S305N-S305N, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Oligodendrocyte NLGF S305N+3 DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Oligodendrocyte NLGF S305N+3 DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF P301S+3 vs. P301S+3

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "P301S", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "P301S", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Oligodendrocyte

```{r}
ctype <- "Oligodendrocyte"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Oligodendrocyte")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Oligodendrocyte")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("P301S", ncol(gsva1_sub)), rep("NLGF_P301S", ncol(gsva2_sub))), levels = c("P301S", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_P301S-P301S, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Oligodendrocyte NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Oligodendrocyte NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Oligodendrocyte")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Oligodendrocyte")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("P301S", ncol(seurat1_sub)), rep("NLGF_P301S", ncol(seurat2_sub))), levels = c("P301S", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_P301S-P301S, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Oligodendrocyte NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Oligodendrocyte NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# NLGF P301S+3 vs. NLGF MAPT KI

## Prep

```{r}
gsva1 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "gsva.rds"))
gsva2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "gsva.rds"))

seurat1 <- readRDS(file.path("..", "cache", "NLGF_MAPTKI", "02", "02", "seurat_cleaned.rds"))
seurat2 <- readRDS(file.path("..", "cache", "NLGF_P301S", "02", "02", "seurat_cleaned.rds"))

data_mat <- GetAssayData(seurat1, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat1 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat1[[]])

data_mat <- GetAssayData(seurat2, layer = "data")
data_mat <- data_mat[match(gene_anno$ensembl_gene_id, rownames(data_mat)), ]
rownames(data_mat) <- gene_anno$external_gene_name
seurat2 <- CreateSeuratObject(counts = NULL, data = data_mat, assay = "SCT", meta.data = seurat2[[]])
```

## Oligodendrocyte

```{r}
ctype <- "Oligodendrocyte"
```

### GSE Analysis

```{r}
gsva1_sub <- gsva1[ , which(seurat1$annotations_short == "Oligodendrocyte")]
gsva2_sub <- gsva2[ , which(seurat2$annotations_short == "Oligodendrocyte")]
gsva1_sub <- ((2 * (gsva1_sub - min(gsva1_sub)) / (max(gsva1_sub) - min(gsva1_sub))) - 1)
gsva2_sub <- ((2 * (gsva2_sub - min(gsva2_sub)) / (max(gsva2_sub) - min(gsva2_sub))) - 1)
gsva <- cbind(gsva1_sub, gsva2_sub)

condition <- factor(
  c(rep("NLGF_MAPTKI", ncol(gsva1_sub)), rep("NLGF_P301S", ncol(gsva2_sub))), levels = c("NLGF_MAPTKI", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(gsva, design)
contrast_mat <- makeContrasts(NLGF_P301S-NLGF_MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
gsva_list <- list(all, up, down)
```

**Upregulated in Oligodendrocyte NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[2]])
datatable_download_exp(gsva_list[[2]][(ncol-3):ncol])
```

**Downregulated in Oligodendrocyte NLGF P301S+3 GSE**

```{r}
ncol <- ncol(gsva_list[[3]])
datatable_download_exp(gsva_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[2]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- gsva[rownames(gsva) %in% rownames(gsva_list[[3]])[1:25], ]
mat <- mat[match(rownames(gsva_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    row_names_max_width = max_text_width(rownames(mat)),
    heatmap_legend_param = list(
      title = "Scaled ssGSEA Score", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

### DGE Analysis

```{r}
seurat1_sub <- GetAssayData(seurat1[ , which(seurat1$annotations_short == "Oligodendrocyte")])
seurat2_sub <- GetAssayData(seurat2[ , which(seurat2$annotations_short == "Oligodendrocyte")])
seurat1_sub <- ((2 * (seurat1_sub - min(seurat1_sub)) / (max(seurat1_sub) - min(seurat1_sub))) - 1)
seurat2_sub <- ((2 * (seurat2_sub - min(seurat2_sub)) / (max(seurat2_sub) - min(seurat2_sub))) - 1)
seurat <- cbind(seurat1_sub, seurat2_sub)

condition <- factor(
  c(rep("NLGF_MAPTKI", ncol(seurat1_sub)), rep("NLGF_P301S", ncol(seurat2_sub))), levels = c("NLGF_MAPTKI", "NLGF_P301S")
)
design <- model.matrix(~ 0 + condition)
colnames(design) <- levels(condition)
fit <- lmFit(seurat, design)
contrast_mat <- makeContrasts(NLGF_P301S-NLGF_MAPTKI, levels = design)
cont_fit <- eBayes(contrasts.fit(fit, contrast_mat), trend = TRUE)

tests <- decideTests(cont_fit, p.value = 5e-2)
tests_up <- tests[tests[ , 1] == 1, ]
tests_down <- tests[tests[ , 1] == -1, ]
results <- topTable(cont_fit, number = Inf)

up <- results[rownames(results) %in% rownames(tests_up), ]
down <- results[rownames(results) %in% rownames(tests_down), ]
all <- rbind(up, down)
if (length(all) > 0) {
  all <- all[order(-all$B), ]
}
seurat_list <- list(all, up, down)
```

**Upregulated in Oligodendrocyte NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[2]])
datatable_download_exp(seurat_list[[2]][(ncol-3):ncol])
```

**Downregulated in Oligodendrocyte NLGF P301S+3 DGE**

```{r}
ncol <- ncol(seurat_list[[3]])
datatable_download_exp(seurat_list[[3]][(ncol-3):ncol])
```

```{r, fig.height = 5}
mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[2]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[2]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)

mat <- as.matrix(seurat[rownames(seurat) %in% rownames(seurat_list[[3]])[1:25], ])
mat <- mat[match(rownames(seurat_list[[3]])[1:25], rownames(mat)), ]
set.seed(1)
draw(
  Heatmap(
    mat,
    colour,
    cluster_rows = FALSE,
    cluster_columns = FALSE,
    show_column_names = FALSE,
    column_split = condition,
    column_title = ctype,
    heatmap_legend_param = list(
      title = "Scaled Normalised Counts", direction = "horizontal",
      legend_width = unit(2.5, "cm"), title_gp = gpar(fontsize = 7, fontface = "bold"), title_position = "topcenter",
      labels_gp = gpar(fontsize = 7), grid_height = unit(0.3, "cm")
    )
  ),
  heatmap_legend_side = "top",
  annotation_legend_side = "top"
)
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
