---
title: "MAPT KI vs. NLGF P301S+3 - Batch 02 - 01 GeneFunnel"
author:
  - name: "Emir Turkes"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_file = file.path("..", "results", "MAPTKI_NLGF_P301S_batch02_01genefunnel.html")
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 35px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [tau-mutant-omics-analysis](https://github.com/eturkes/tau-mutant-omics-analysis).*

The table of contents in the top left is clickable and can be used to quickly navigate the document.
To toggle the visibility of code, use the `CODE` toggles at the top right of chunks.
The toggle at the start of the document controls the visibility of all chunks.

```{r}
#    This file is part of tau-mutant-omics-analysis.
#    Copyright (C) 2023-2024  Emir Turkes, Naoto Watamura, UK DRI at UCL
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# Load required packages, suppressing startup messages.
# -----------------------------------------------------
library(conflicted)
conflicts_prefer(AnnotationDbi::select, edgeR::cpm, DT::JS, base::intersect, reshape2::melt, .quiet = TRUE)
packages <- c(
  "GSEABase", "Seurat", "ggplot2", "dplyr", "ggrepel", "scuttle", "GO.db", "tm", "pals", "ComplexHeatmap", "edgeR",
  "DT", "circlize", "purrr", "networkD3", "factoextra", "reshape2", "dendextend", "data.table"
)
invisible(suppressPackageStartupMessages(lapply(packages, FUN = library, character.only = TRUE)))
# -----------------------------------------------------

# Define global settings.
# -----------------------
knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# -----------------------

# Define functions.
# -----------------
source("utils.R")

`%notin%` <- Negate(`%in%`)
# -----------------

# Useful variables.
# -----------------
genotype <- "MAPTKI_NLGF_P301S"
batch <- 2
step <- 1

colour <- colorRamp2(c(0, 5, 10), c("#0077BB", "#FFFFFF", "#CC3311"))

cache_dir <- file.path("..", "cache", genotype, paste0("0", batch), paste0("0", step))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# -----------------

# Useful data.
# ------------
gene_anno <- readRDS(file.path("..", "cache", "MAPTKI", "02", "02", "gene_anno.rds"))
gene_sets <- readRDS(file.path("..", "cache", "filtered_gene_sets.rds"))
# ------------
```

# Merging

```{r}
rds <- file.path(cache_dir, "gse_merged.rds")
if (file.exists(rds)) {
  gse <- readRDS(rds)
} else {
  file_list <- list.files(file.path(cache_dir, "..", "..", ".."), recursive = TRUE)
  file_list <- file_list[grep("pseudobulk_gse.rds", x = file_list)]
  file_list <- c(file_list[1], file_list[3])
  files <- vector("list", length = length(file_list))
  for (i in seq_along(files)) {
    files[[i]] <- counts(readRDS(file.path(cache_dir, "..", "..", "..", file_list[i])))
    colnames(files[[i]]) <- paste(colnames(files[[i]]), sub("\\/.*", "", file_list[i]), sep = " ")
  }
  gse <- Reduce(cbind, x = files)
  saveRDS(gse, file = rds)
}

rds <- file.path(cache_dir, "counts_merged.rds")
if (file.exists(rds)) {
  counts <- readRDS(rds)
} else {
  file_list <- list.files(file.path(cache_dir, "..", "..", ".."), recursive = TRUE)
  file_list <- file_list[grep("pseudobulk_counts.rds", x = file_list)]
  file_list <- c(file_list[1], file_list[3])
  files <- vector("list", length = length(file_list))
  for (i in seq_along(files)) {
    files[[i]] <- counts(readRDS(file.path(cache_dir, "..", "..", "..", file_list[i])))
    colnames(files[[i]]) <- paste(colnames(files[[i]]), sub("\\/.*", "", file_list[i]), sep = " ")
  }
  counts <- Reduce(cbind, x = files)
  saveRDS(counts, file = rds)
}
```

# Gene Set Clustering

```{r}
gene_sets_def <- gene_sets
for (i in seq_along(gene_sets_def@.Data)) {
  suppressWarnings(gene_sets_def[[i]]@setName <- gene_sets_def[[i]]@shortDescription)
}

go_anno <- select(GO.db, names(gene_sets_def), columns = c("GOID", "DEFINITION"), keytype = "GOID")
gene_sets_def <- gene_sets
for (i in seq_along(gene_sets_def@.Data)) {
  suppressWarnings(gene_sets_def[[i]]@setName <- paste(gene_sets_def[[i]]@setName, go_anno$DEFINITION[i], sep = " "))
}

overlap <- computeGeneSetsOverlapMax(gene_sets_def, uniqGenes = unique(unlist(geneIds(gene_sets_def))))
dendro <- hclust(get_dist(overlap, method = "spearman"), method = "ward.D2")

rds <- file.path(cache_dir, "cluster_list.rds")
if (file.exists(rds)) {
  cluster_list <- readRDS(rds)
} else {
  cluster_list <- vector("list", length = 25)
  for (i in 2:25) {
    print(i)
    clustering <- cutree(dendro, k = i)
    clustering <- split(names(clustering), unname(clustering))

    dict_list <- vector("list", length = length(clustering))
    for (j in seq_along(dict_list)) {
      dict_list[[j]] <- suppressWarnings(word_cloud(clustering[[j]]))
      dict_list[[j]] <- strsplit(paste(unlist(dict_list[[j]]), collapse = " "), "\\s+")[[1]]
      dict_list[[j]] <- unique(dict_list[[j]])
    }
    dict <- table(unlist(dict_list))
    dict <- names(dict[dict == length(dict_list)])

    word_list <- vector("list", length = length(clustering))
    for (j in seq_along(word_list)) {

      for (k in seq_along(clustering[[j]])) {
        word_list[[j]][[k]] <- suppressWarnings(word_cloud(clustering[[j]][k], dict = dict))
        word_list[[j]][[k]] <- strsplit(paste(unlist(word_list[[j]][[k]]), collapse = " "), "\\s+")[[1]]
      }
      
      cluster_list[[i]][[j]] <- table(unlist(word_list[[j]]))
      cluster_list[[i]][[j]] <- cluster_list[[i]][[j]][order(cluster_list[[i]][[j]], decreasing = TRUE)]
    }
  }
  saveRDS(cluster_list, file = rds)
}

stats <- vector(length = length(cluster_list))
for (i in seq_along(stats)) {
  if (i != 1) {
    scores <- vector(length = length(cluster_list[[i]]))
    for (j in seq_along(scores)) {
      scores[j] <- max(cluster_list[[i]][[j]])
    }
    stats[i] <- mean(scores)
  }
}
optimal_k <- which.max(stats)

cluster_colours <- cols25(optimal_k)
clustering <- color_branches(dendro, k = optimal_k, col = cluster_colours)

cluster_labels <- cluster_list[[optimal_k]]
cluster_labels <- unlist(map(lapply(cluster_labels, FUN = names), .f = 1))
```

# All Cell-types

```{r, fig.height = 6, fig.width = 4}
gse_cpm <- log1p(cpm(gse))
counts_cpm <- log1p(cpm(counts))

groups <- factor(sub(".*\\s", "", colnames(gse_cpm)), levels = unique(sub(".*\\s", "", colnames(gse_cpm))))

top_anno <- HeatmapAnnotation(
  Genotype = anno_block(
    gp = gpar(fill = c("lightgray", "darkgray"), lwd = 0.5),
    labels = unique(groups),
    labels_gp = gpar(fontsize = 4),
    height = unit(2, "mm")
  )
)

left_anno <-  HeatmapAnnotation(
  cluster_labels = anno_block(
    gp = gpar(col = cluster_colours),
    labels = seq(optimal_k),
    labels_rot = 0,
    labels_gp = gpar(fontsize = 4),
    width = unit(1.5, "mm")
  ),
  which = "row"
)

set.seed(1)
draw(
  Heatmap(
    gse_cpm,
    col = colour,
    column_labels = sub("\\s+[^ ]+$", "", colnames(gse_cpm)),
    column_split = factor(groups, levels = unique(groups)),
    top_annotation = top_anno,
    left_annotation = left_anno,
    cluster_rows = clustering,
    row_split = optimal_k,
    rect_gp = gpar(col = "darkgray", lwd = 0.25),
    border_gp = gpar(lwd = 0.5),
    border = TRUE,
    row_gap = unit(1.75, "mm"),
    column_title = NULL,
    show_row_names = FALSE,
    column_names_side = "top",
    column_names_rot = 45,
    column_names_gp = gpar(fontsize = 3.5),
    cluster_columns = FALSE,
    row_title = cluster_labels,
    row_title_rot = 0,
    row_title_gp = gpar(fontsize = 5),
    heatmap_legend_param = list(
      title = "Log + 1", direction = "horizontal", title_position = "topcenter",
      legend_width = unit(2, "cm"), title_gp = gpar(fontsize = 4), labels_gp = gpar(fontsize = 3),
      grid_height = unit(0.1, "cm"), at = c(0, 2, 4, 6, 8, 10)
    )
  ),
  heatmap_legend_side = "top"
)
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
