---
title: "Otero-Garcia - Batch 01 - 02 GeneFunnel"
author:
  - name: "Emir Turkes"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_file = file.path("..", "results", "oterogarcia_batch01_02genefunnel.html")
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
.tocify-subheader .tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 35px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [tau-mutant-omics-analysis](https://github.com/eturkes/tau-mutant-omics-analysis).*

The table of contents in the top left is clickable and can be used to quickly navigate the document.
To toggle the visibility of code, use the `CODE` toggles at the top right of chunks.
The toggle at the start of the document controls the visibility of all chunks.

```{r}
#    This file is part of tau-mutant-omics-analysis.
#    Copyright (C) 2023-2024  Emir Turkes, Naoto Watamura, UK DRI at UCL
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# Load required packages, suppressing startup messages.
# -----------------------------------------------------
library(conflicted)
conflicts_prefer(AnnotationDbi::select, DT::JS, edgeR::cpm, .quiet = TRUE)
packages <- c(
  "GSEABase", "GSVA", "Seurat", "biomaRt", "parallelly", "dplyr", "scuttle", "GO.db", "tm", "pals", "ComplexHeatmap",
  "DT", "circlize", "edgeR", "purrr", "networkD3", "tm"
)
invisible(suppressPackageStartupMessages(lapply(packages, FUN = library, character.only = TRUE)))
# -----------------------------------------------------

# Define global settings.
# -----------------------
knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# -----------------------

# Define functions.
# -----------------
source("utils.R")

`%notin%` <- Negate(`%in%`)
# -----------------

# Useful variables.
# -----------------
sample <- "oterogarcia"
batch <- 1
step <- 2

cache_dir <- file.path("..", "cache", sample, paste0("0", batch), paste0("0", step))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# -----------------
```

# Prep

```{r}
rds <- file.path("..", "cache", "filtered_gene_sets_hsapiens.rds")
if (file.exists(rds)) {
  gene_sets <- readRDS(rds)
} else {

  gene_sets <- getGmt(
    file.path("..", "data", "gene-sets", "gprofiler_hsapiens.ENSG_2024.03.12", "hsapiens.GO.comb.ENSG.gmt")
  )
  keep <- filterGeneSets(gene_sets, min.sz = 2, max.sz = 1000)
  gene_sets <- gene_sets[names(gene_sets) %in% names(keep)]

  for (i in seq(length(gene_sets@.Data))) {
    go_id <- gene_sets[[i]]@setName
    suppressWarnings(gene_sets[[i]]@setName <- gene_sets[[i]]@shortDescription)
    suppressWarnings(gene_sets[[i]]@shortDescription <- go_id)
  }

  remove <- grep("regulation of", names(gene_sets))
  if (length(remove) > 0) {
    gene_sets <- gene_sets[-remove]
  }

  overlap <- computeGeneSetsOverlapMax(gene_sets, uniqGenes = unique(unlist(geneIds(gene_sets))))
  tmp <- rowSums(overlap)
  tmp <- tmp[order(tmp, decreasing = TRUE)]
  gene_sets_sorted <- gene_sets[match(names(tmp), names(gene_sets))]

  overlap <- computeGeneSetsOverlapMax(gene_sets_sorted, uniqGenes = unique(unlist(geneIds(gene_sets_sorted))))
  overlap[upper.tri(overlap)] <- 0
  diag(overlap) <- 0
  keep <- apply(overlap, MARGIN = 1, FUN = max)
  keep <- keep[keep <= 0.25]
  gene_sets <- gene_sets[names(gene_sets) %in% names(keep)]

  rm(overlap, gene_sets_sorted)
  saveRDS(gene_sets, rds)
}

rds <- file.path(cache_dir, "seurat_cleaned.rds")
rds2 <- file.path(cache_dir, "gene_anno.rds")
if (file.exists(rds) & file.exists(rds2)) {
  seurat <- readRDS(rds)
  gene_anno <- readRDS(rds2)
} else {

  seurat <- readRDS(file.path(cache_dir, "..", "01", "seurat_annotated.rds"))
  # DefaultAssay(seurat) <- "RNA"
  # seurat[["SCT"]] <- NULL

  mart <- useEnsembl("ensembl", "hsapiens_gene_ensembl")
  attributes <- c("external_gene_name", "ensembl_gene_id")
  gene_anno <- getBM(attributes, filters = "ensembl_gene_id", values = unique(unlist(geneIds(gene_sets))), mart = mart)
  add <- which(gene_anno$external_gene_name == "")
  for (idx in add) {
    gene_anno[idx, 1] <- gene_anno[idx, 2]
  }

  dup <- gene_anno[
    duplicated(gene_anno$external_gene_name) | duplicated(gene_anno$external_gene_name, fromLast = TRUE),
  ]
  if (nrow(dup) > 0) {
    for (i in seq(nrow(dup))) {
      for (j in seq(nrow(gene_anno))) {
        if (dup$ensembl_gene_id[i] == gene_anno$ensembl_gene_id[j]) {
          gene_anno$external_gene_name[j] <- paste(
            gene_anno$external_gene_name[j], gene_anno$ensembl_gene_id[j], sep = "-"
          )
        }
      }
    }
  }

  gene_anno_sub <- gene_anno[gene_anno$ensembl_gene_id %in% rownames(seurat), ]

  counts_mat <- GetAssayData(seurat, layer = "counts")
  counts_mat <- counts_mat[rownames(counts_mat) %in% gene_anno_sub$ensembl_gene_id, ]
  gene_anno_sub <- gene_anno_sub[match(rownames(counts_mat), gene_anno_sub$ensembl_gene_id), ]
  # rownames(counts_mat) <- gene_anno_sub$ensembl_gene_id

  data_mat <- GetAssayData(seurat)
  data_mat <- data_mat[rownames(data_mat) %in% gene_anno_sub$ensembl_gene_id, ]
  gene_anno_sub <- gene_anno_sub[match(rownames(data_mat), gene_anno_sub$ensembl_gene_id), ]
  # rownames(data_mat) <- gene_anno_sub$ensembl_gene_id

  add <- unique(gene_anno$ensembl_gene_id[which(gene_anno$ensembl_gene_id %notin% rownames(counts_mat))])
  counts_null_mat <- matrix(0, nrow = length(add), ncol = ncol(counts_mat))
  rownames(counts_null_mat) <- add
  colnames(counts_null_mat) <- colnames(counts_mat)
  data_null_mat <- counts_null_mat

  set.seed(1)
  col <- sample(seq(ncol(counts_null_mat)), size = length(add), replace = TRUE)
  for (i in seq_along(add)) {
    counts_null_mat[i, col[i]] <- 0
    data_null_mat[i, col[i]] <- log1p(0)
  }

  counts_mat <- rbind(counts_mat, counts_null_mat)
  data_mat <- rbind(data_mat, data_null_mat)
  seurat <- CreateSeuratObject(counts = counts_mat, data = data_mat, assay = "RNA", meta.data = seurat[[]])
  gene_anno <- gene_anno[match(rownames(seurat), gene_anno$ensembl_gene_id), ]

  rm(counts_mat, data_mat, counts_null_mat, data_null_mat, mart, dup, gene_anno_sub)
  gc()
  seurat <- SCTransform(seurat, vst.flavor = "v2", min_cells = 0, verbose = FALSE)

  mat <- GetAssayData(seurat, layer = "counts")
  mat[is.na(mat)] <- 0
  seurat <- SetAssayData(seurat, layer = "counts", new.data = mat)
  mat <- GetAssayData(seurat, layer = "data")
  mat[is.na(mat)] <- 0
  seurat <- SetAssayData(seurat, layer = "data", new.data = mat)

  saveRDS(gene_anno, rds2)
  saveRDS(seurat, rds)
}

# keep <- which(seurat$Cell.Types == "Ex01_CUX2-LAMP5 (L2-L3)")
# seurat <- seurat[ , keep]

remove <- which(seurat$SORT == "MAP2control")
seurat <- seurat[ , -remove]

# remove <- grep("Unknown", x = seurat$annotations_short)
# if (length(remove) > 0) {
#   seurat <- seurat[ , -remove]
# }
# 
# seurat$annotations_short <- factor(
#   seurat$annotations_short,
#   levels = c(
#       "Dentate Gyrus Glutamatergic", "Hippocampal Glutamatergic", "Cortical Glutamatergic", "GABAergic", "Endothelial",
#       "Astrocyte", "OPC", "Oligodendrocyte", "Microglia"
#   )
# )

seurat$Sample.ID <- factor(seurat$Sample.ID, levels = c(paste0("AD-MAP2-", rep(1:8)), paste0("AD-AT8-", rep(1:8))))

rds <- file.path(cache_dir, "pseudobulk_counts.rds")
if (file.exists(rds)) {
  pseudobulk_counts <- readRDS(rds)
} else {
  pseudobulk_counts <- as.SingleCellExperiment(seurat, assay = "SCT")
  pseudobulk_counts <- suppressWarnings(
    aggregateAcrossCells(
      pseudobulk_counts, colData(pseudobulk_counts)[ , "Sample.ID"],
      use_exprs_values = "counts", statistics = "sum"
    )
  )
  logcounts(pseudobulk_counts) <- log1p(counts(pseudobulk_counts))
  saveRDS(pseudobulk_counts, file = rds)
}

plot(density(counts(pseudobulk_counts)))
plot(density(logcounts(pseudobulk_counts)))

counts <- counts(pseudobulk_counts)
gene_ids <- geneIds(gene_sets)

rds <- file.path(cache_dir, "pseudobulk_gse.rds")
if (file.exists(rds)) {
  pseudobulk_gse <- readRDS(rds)
} else {

  mat <- matrix(NA, nrow = length(gene_sets) + 1, ncol = 1)
  for (j in seq(ncol(pseudobulk_counts))) {
    tmp <- matrix(NA, nrow = 1, ncol = 1)
    for (i in seq_along(gene_sets)) {
      idx <- which(rownames(pseudobulk_counts) %in% geneIds(gene_sets[[i]]))
      idx <- counts(pseudobulk_counts)[idx, j]
      sum <- sum(idx)
      var <- exp(sd(idx) / sum)
      if (is.na(var)) {
        var <- 1
      }
      length <- sqrt(length(idx))
      score <- sum / var / length
      tmp <- rbind(tmp, score)
    }
    mat <- cbind(mat, tmp)
  }
  mat <- mat[-1, -1]

  pseudobulk_gse <- pseudobulk_counts[1:nrow(mat), ]
  rownames(pseudobulk_gse) <- names(gene_sets)
  counts(pseudobulk_gse, withDimnames = FALSE) <- mat
  logcounts(pseudobulk_gse, withDimnames = FALSE) <- log1p(counts(pseudobulk_gse))

  saveRDS(pseudobulk_gse, file = rds)
}

plot(density(counts(pseudobulk_gse)))
plot(density(logcounts(pseudobulk_gse)))
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r}
sessionInfo()
```
